---
title: "CenSoc-Numident Sibships User Guide[^updated]"
output:
  pdf_document: null
  html_document:
    df_print: paged
urlcolor: blue
header-includes:
- \usepackage{hyperref}
- \usepackage{float}
- \restylefloat{table}
- \usepackage{caption}
---
[^updated]: Last updated: `r format(Sys.time(), '%d %B, %Y')`

```{r setup, include=FALSE}
#knitr::opts_chunk$set(include=FALSE)

```


```{r, include=FALSE, echo = FALSE}
library(tidyverse)
library(data.table)
library(kableExtra)
```

# Description

The purpose of the CenSoc-Numident sibling datasets is to allow researchers to identify siblings in CenSoc-Numident data. 

The CenSoc-Numident links Social Security Numident records (the Berkeley Unified Numident Mortality Database) with the 1940 Census. Siblings in in CenSoc-Numident are located by finding individuals with shared parent names as recorded in Social Security Numident records. No information from the 1940 Census is used to identify siblings. We use two methods to locate siblings from Numident records:

1) The **exact** match method identifies siblings only with exactly identical parent first and last names (after parents' names have undergone cleaning and standardization). This is the most match stringent method.

2) The **flexible** match method allows for parents' names to be slightly different, within a threshold using Jaro–Winkler string distance, in addition to exact matches. This allows siblings to be matched even in cases of minor misspellings, mistranscriptions, or spelling variations in parents' names (e.g., mother's maiden name recorded as "Brannum" and "Branum"). This increases the number of siblings found, but has higher potential to falsely match unrelated individuals. 

Sibling groups are identified among individuals in the CenSoc-Numident who died at age 65+ in the years 1988-2005. They may be of any gender composition and contain 2-8 siblings. An overview of the resulting datasets created by each method is presented below:

```{r, include=FALSE, echo = FALSE}
# Read numident data
numident_sibs_exact <- fread("/data/censoc/censoc_data_releases/siblings/siblings_v1/censoc_numident_siblings_v1/censoc_numident_siblings_exact_match.csv")
numident_sibs_flex <- fread("/data/censoc/censoc_data_releases/siblings/siblings_v1/censoc_numident_siblings_v1/censoc_numident_siblings_flexible_match.csv")
```

```{r, eval = FALSE, include=FALSE, echo = FALSE}
# get basic metrics to make table below

nrow(numident_sibs_exact) # total size
numident_sibs_exact %>% group_by(sib_group_id_exact) %>% filter(row_number() == 1) %>% nrow() # number of groups
numident_sibs_exact %>% group_by(sib_group_id_exact) %>% mutate(n_sibs=n()) %>% ungroup() %>% 
  group_by(n_sibs) %>% tally() # group sizes


nrow(numident_sibs_flex) # total size
numident_sibs_flex %>% group_by(sib_group_id_flexible) %>% filter(row_number() == 1) %>% nrow() # number of groups
numident_sibs_flex %>% group_by(sib_group_id_flexible) %>% mutate(n_sibs=n()) %>% ungroup() %>% 
  group_by(n_sibs) %>% tally() # group sizes

```


```{r, include=FALSE, echo = FALSE}
df <- tibble(
             "Match method" = c("Exact match", "Flexble match"),
             "Number of individuals" = c("908,472", "1,185,055"),
             "Number of sibling groups" = c("426,952", "551,641"),
             "Mean sibling group size" = c("2.13", "2.15")
             ) %>% 
  knitr::kable(format = "pipe")
```

```{r,echo =F}
df %>% kable_styling(full_width = F)
```


For a detailed description of sibling identification methodologies and characteristics of sibships, please see the paper: *Methods for Identifying Siblings in Administrative Mortality Data* available at [https://censoc.berkeley.edu/documentation/](https://censoc.berkeley.edu/documentation/).



# Usage

Each sibling identifier dataset consist of two columns: a unique individual identifier `HISTID` that can be used to merge the data with the full numident, and an identifier for each sibling group. The group identifier corresponds to the HISTID of one sibling.

\vspace{15pt}

```{r, include=TRUE, echo=TRUE}
# Show first 6 rows the exact match
head(numident_sibs_exact)
```

\vspace{15pt}

To get the subset of the CenSoc-Numident consisting of only identified siblings, a join on `HISTID` may be used, as in the example code below
```{r, include = TRUE, eval = FALSE, echo=TRUE}
# Read CenSoc-Numident 
numi <- fread("censoc_numident_v3.csv")
# Read CenSoc-Numident siblings 
numi_sib_id <- fread("censoc_numident_siblings_exact_match.csv")
# Attach numident variables to siblings, keeping only people in sibships
numi_sibs <- inner_join(numi, numi_sib_id, by = "HISTID")
```
