---
title: "3 Creat CenSoc Male-Only Sample"
output: html_notebook
---

### Set up  
- Create a new local project
    - New project from version control
    - Clone censoc from Github:  https://github.com/MJAlexander/censoc
- Update path names. 

This contains original code written by Monica Alexander (monica.alexander@utoronto.ca). 
    
### Load Packages

```{r}
library(censoc)
library(tidyverse)
library(data.table)
```

### State Match

```{r, eval = FALSE}
# keep a log

update_file <- "//home/ipums/casey-ipums/censoc/make_censoc/progress.txt"

matched_data_folder <- "//home/ipums/casey-ipums/censoc/make_censoc/states/"

# load in the socsec files

socsec_files = c("/home/ipums/josh-ipums/progs/ssdm/ssdm1",
                     "/home/ipums/josh-ipums/progs/ssdm/ssdm2",
                     "/home/ipums/josh-ipums/progs/ssdm/ssdm3")

socsec <- load_socsec_deaths(socsec_files)
sink(update_file)
cat("Socsec has been read in. \n")

# get the list of state census files

state_file_list <- list.files("/home/ipums/casey-ipums/IPUMS/1940/CSV/")
state_fips <- as.numeric(gsub("\\D", "", state_file_list))
state_fips <- state_fips[!is.na(state_fips)]
state_file_list <- paste0("/home/ipums/casey-ipums/IPUMS/1940/CSV/", state_file_list[grepl("state", state_file_list)])

cols_to_keep <- c("URBAN", "OWNERSHP", "RENT", "RACE", "MARST", "BPL", "HISPAN",
                  "EMPSTAT", "OCCSCORE", "INCWAGE","INCNONWG",
                  "EDUC",
                  "MBPL", "FBPL", "OCC")

# loop over each state file, do the match, and save the matched output

for(i in 1:length(state_file_list)){
  cat(paste("Reading in census file", i, "\n"))
  census <- try(load_census(census_file = state_file_list[i], cols_to_keep = cols_to_keep))
  if(class(census)=="try-error"){
    cat(paste("File", i, "is empty \n"))
  }
  else{
    cat(paste("Matching census file", i, "\n"))
    censoc <- create_censoc(census, socsec)
    cat(paste("Writing file", i, "\n"))
    write_csv(censoc, paste0(matched_data_folder, state_fips[i], "_matched.csv"))
  }
}

rm(socsec)

sink()

```

- Several warnings messages (1 per state): the condition has length > 1 and only the first element will be used

### National match
```{r, eval = FALSE}
# state censoc files
files  <- paste0("~/censoc/make_censoc/states//",list.files(path = "~/censoc/make_censoc/states//", pattern = '\\.csv') )

# national censoc
all_censoc_unique <- create_national_censoc(files)

# for BFDW2018
cols_to_keep <- c("URBAN", "OWNERSHP", "RENT", "RACE", "MARST", "BPL", "HISPAN",
                  "EMPSTAT", "OCCSCORE", "INCWAGE","INCNONWG",
                  "EDUC",
                  "MBPL", "FBPL", "OCC", "NAMELAST", "NAMEFRST", "lname.y", "fname.y", "mname")

bfdw <- all_censoc_unique[, c("SERIAL40", "SERIAL", "NUMPREC40", "PERNUM", "STATEFIP",  "census_age.x", "byear", "dyear", "bmonth", "dmonth", cols_to_keep)]
bfdw <- bfdw %>% rename(census_age = census_age.x)
write_csv(bfdw, path  = "~/censoc/make_censoc/censoc_bfdw.csv")

# save dataset with less variables
dr <- bfdw[, c("SERIAL", "PERNUM", "STATEFIP",  "census_age", "byear", "dyear", "bmonth", "dmonth")]
write_csv(dr, "~/censoc/make_censoc/national_matched_states_red.csv")

```


This is new code written by Casey Breen. It dedupes keys at the national level, instead of the state level. 

## 1940 Match 
```{r}
socsec_files = c("/home/ipums/josh-ipums/progs/ssdm/ssdm1",
                     "/home/ipums/josh-ipums/progs/ssdm/ssdm2",
                     "/home/ipums/josh-ipums/progs/ssdm/ssdm3")

socsec <- load_socsec_deaths(socsec_files)
census <- load_census(census_file = "/home/ipums/casey-ipums/IPUMS2019/1920/TSV/P.tsv")
censoc <- create_censoc(census, socsec)

```

## 1930 Match 
```{r}
## Load SocSec Files
socsec_files = c("/home/ipums/josh-ipums/progs/ssdm/ssdm1",
                     "/home/ipums/josh-ipums/progs/ssdm/ssdm2",
                     "/home/ipums/josh-ipums/progs/ssdm/ssdm3")

## Read in load socsec deaths function below (adapted for 1930 census)
socsec <- load_socsec_deaths_1920(socsec_files)
census <- load_census(census_file = "/home/ipums/casey-ipums/IPUMS2019/1920/TSV/P.tsv")
censoc <- create_censoc(census, socsec)

load_socsec_deaths_1920 <- function(socsec_files){

  socsec_list<- list()
  for(i in 1:length(socsec_files)){
    socsec_file <- socsec_files[i]
    cat(paste0("Reading in file ", i, ".\n"))
    tt <- read_fwf(socsec_file,
                          fwf_widths(c(1,9, 20, 4, 15, 15, 1, 8, 8),
                                     col_names = c("mode", "ssn", "lname",
                                                   "name_suffix", "fname", "mname",
                                                   "vorpcode", "dod", "dob")))
    assign(paste0("socsec",i) , as.data.table(tt))
    socsec_list[[i]] <-eval(parse(text = paste0("socsec",i)))
    rm(tt)
  }

  socsec <- rbindlist(socsec_list)
  rm(socsec_list, socsec1, socsec2, socsec3)

  cat("Cleaning variables. \n")
  ## A. clean the socsec data
  ## shorten names by removing blank space at end
  socsec[,"lname_orginal_socsec" := lname]
  socsec[,"fname_orginal_socsec" := fname]
  socsec[,lname := gsub(pattern = "\\s*$",
                        replacement = "", x = lname)]
  socsec[,fname := gsub(pattern = "\\s*$",
                        replacement = "", x = fname)]
  ## now get birth and death year
  socsec[,"byear" := as.numeric(substr(dob, 5, 9))]
  socsec[,"dyear" := as.numeric(substr(dod, 5, 9))]
  ## birth and death month
  socsec[,"bmonth" := as.numeric(substr(dob, 1, 2))]
  socsec[,"dmonth" := as.numeric(substr(dod, 1, 2))]
  ## birth and death dat
  socsec[,"bday" := as.numeric(substr(dob, 3, 4))]
  socsec[,"dday" := as.numeric(substr(dod, 3, 4))]
  ## now get census_age
  socsec[,"census_age" := ifelse(bmonth < 4,
                                 1920 - byear,
                                 1919 - byear)]

  ## filter to only include those with age info, that were born in 1940, and died after 1975.
  socsec <- socsec[socsec$census_age!=1920&socsec$census_age!=1919,]
  socsec <- socsec[socsec$census_age>=0,]
  socsec <- socsec[socsec$dyear>=1975,]

  cat("Creating keys. \n")
  # create key
  socsec[,"tmp_key" := paste(lname, fname, census_age, sep = "_")]
  socsec[,"clean_key" := clean_key(tmp_key),]
  socsec[,"n_clean_key" := .N, by = clean_key]

  socsec <- socsec[,!"tmp_key"]
  return(socsec)

}
```


