---
title: "CenSoc World War II Enlistment File Codebook[^updated]"
output:
  pdf_document: default
header-includes: \usepackage{hyperref}
---

[^updated]: Last updated: `r format(Sys.time(), '%d %B, %Y')`


| Page| Variable     | Label                                        |
|----:|:-------------|:---------------------------------------------|
|   \hyperlink{page.2}{2}  | \hyperlink{page.2}{serial\_number}         |Army Serial Number  |
|   \hyperlink{page.3}{3}  | \hyperlink{page.3}{byear}           |Year of Birth    |
|   \hyperlink{page.4}{4}  | \hyperlink{page.4}{fname}           |First Name                      |
|   \hyperlink{page.5}{5}  |\hyperlink{page.5}{mname}            |Middle Name                      |
|   \hyperlink{page.6}{6}  | \hyperlink{page.6}{lname}          |Last Name                  |
|   \hyperlink{page.7}{7}  |\hyperlink{page.7}{sex}             |Sex              |
|   \hyperlink{page.8}{8}  | \hyperlink{page.8}{date\_of\_enlistment}  |Date of Enlistment     |
|  \hyperlink{page.9}{9} |\hyperlink{page.9}{bpl}              |Place of Birth |
|  \hyperlink{page.10}{10} |\hyperlink{page.10}{residence\_state}           |State of Residence at Enlistment|
|  \hyperlink{page.11}{11} |\hyperlink{page.11}{residence\_county}           |County of Residence at Enlistment|
|  \hyperlink{page.12}{12} |\hyperlink{page.12}{place\_of\_enlistment}          |Place of Enlistment|
|  \hyperlink{page.13}{13} |\hyperlink{page.13}{education}                   |Education |
|  \hyperlink{page.14}{14} |\hyperlink{page.14}{grade\_code}                    |Army Grade (rank)|
|  \hyperlink{page.15}{15} |\hyperlink{page.15}{branch\_code}                     |Army Branch |
|  \hyperlink{page.16}{16} |\hyperlink{page.16}{term\_of\_enlistment}            |Term of Enlistment |
|  \hyperlink{page.17}{17} |\hyperlink{page.17}{race}    |Race  |
|  \hyperlink{page.18}{18} |\hyperlink{page.18}{citizenship}                 |Citizenship  |
|  \hyperlink{page.19}{19} |\hyperlink{page.19}{civilian\_occupation}            |Civilian Occupation |
|  \hyperlink{page.20}{20} |\hyperlink{page.20}{marital\_status}            |Marital Status at Enlistment |
|  \hyperlink{page.21}{21} |\hyperlink{page.21}{height}                |Height at Enlistment (inches) |
|  \hyperlink{page.22}{22} |\hyperlink{page.22}{weight\_or\_AGCT}                |Weight (pounds) or AGCT score|
|  \hyperlink{page.23}{23} |\hyperlink{page.23}{component\_enlistment}             |Army component |
|  \hyperlink{page.24}{24} |\hyperlink{page.24}{source}                          |Source of Army Personnel |

**Summary:** The CenSoc World War II Enlistment  File captures 9,200,232 of the 15 million enlistees of World War II. The dataset contains 22 variables on demographic information of the men and women enlisted in the United States Army and the Women’s Army Auxiliary Corps. 



```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
options(kableExtra.auto_format = FALSE, scipen=999)
```

```{r echo = F, results='hide', message=FALSE, warning=FALSE}
## Library Packages
library(tidyverse)
library(data.table)
library(stringi)
library(lubridate)
library(kableExtra)

data_path <- '/data/josh/CenSoc/censoc_data/enlistment-records-public/code/awenlistmentcode/finalenlistment.csv'
resource_file_path <- '/data/josh/CenSoc/censoc_data/enlistment-records-public/code/awenlistmentcode'
path_to_temp_data <- "/data/josh/CenSoc/censoc_data/enlistment-records-public/code/awenlistmentcode/temp3.csv"
out_path          <- "/data/josh/CenSoc/censoc_data/enlistment-records-public/code/awenlistmentcode"
codebook_path <- "/censoc_new/censocdev/codebase/04_create_enlistment_records_v2/10-enlistment-codebook_files"
## read in data file
#setwd(data_path)
#enlistment_dmf <- fread('finalenlistment.csv')


setwd(resource_file_path)
test_file_path <- '/data/josh/CenSoc/censoc_data/enlistment-records-public/code/awenlistmentcode'
enlistment_records_final <- fread('finalenlistment.csv')



```
```{r}
#read in temp2
data_combined <- read_csv(path_to_temp_data, 
                          col_types = cols(
                            empty_field = col_character(),
                            term_or_enlistment = col_character(),
                            longevity = col_character(),
                            education = col_character(),
                            marital_status = col_character(),
                            component = col_character(),
                            educ_spec = col_character(),
                            defer_date_mmyy = col_character(),
                            mname_clean = col_character(),
                            sex = col_integer(),
                            sex_r = col_integer(),
                            fname_std = col_character()
                            )
                          )

```



\newpage

**\huge serial\_number**
\normalsize
\vspace{12pt}

**Label**: Army Serial Number

**Description**: serial_number is a character variable reporting a person’s serial number, as recorded in the World War II Army Enlistment Records. This variable uniquely identifies each enlistee.


\newpage 

**\huge byear **
\normalsize
\vspace{12pt}

**Label**: Year of Birth 

**Description**: byear is a numeric variable reporting a person’s year of birth, as recorded in the World War II Army Enlistment Records.

```{r, include = F}
byear_plot <- enlistment_records_final %>%
    group_by(byear) %>%
    summarise(n = n()) %>%
    ggplot(aes(x = byear, y = n)) + 
    geom_line() + 
    geom_point() +
    theme_minimal(base_size = 15) + #replace with a different theme (theme_bw()) if the bbplot package isn't downloaded 
  ggtitle("Year of Birth") + 
  theme(legend.position="bottom") +
  xlab("title") + 
  labs(x = "Year", 
       y = "Count") + 
  scale_y_continuous(labels = scales::comma) +
  scale_x_continuous(breaks = scales::pretty_breaks(n=5)) 
```

\vspace{75pt}

```{r, echo = F}
byear_plot
```




\newpage 
**\huge fname **
\normalsize
\vspace{12pt}

**Label:** First Name

**Description:** fname is a character variable reporting a person’s first name, as recorded in the World War II Army Enlistment Records. This variable was cleaned from raw first names by removing titles (e.g., Dr.) and replacing nicknames with standard names (e.g., Billy to William. Non-alphabetical characters were also removed.





\newpage 

**\huge mname **
\normalsize
\vspace{12pt}

**Label**: Middle Name

**Description**: mname is a character variable reporting a person’s middle name, as recorded in the World War II Army Enlistment Records. This variable was cleaned from raw middle names by removing titles (e.g., Dr.) and replacing nicknames with standard names (e.g., Billy to William. Non-alphabetical characters were also removed.






\newpage 

**\huge lname **
\normalsize
\vspace{12pt}

**Label**: Last Name

**Description**: lname is a character variable reporting a person’s last name, as recorded in the World War II Army Enlistment Records. This variable was cleaned from raw last names by removing titles (e.g., Dr.) and replacing nicknames with standard names (e.g., Billy to William. Non-alphabetical characters were also removed.


\vspace{30pt}




\newpage
**\huge sex **
\normalsize
\vspace{12pt}

**Label**: Sex

**Description**: sex is a numeric variable that reports a person's sex, as recorded in the World War II Army Enlistment Records.    

```{r, include = F}
sex_tabulated <- knitr::kable(enlistment_records_final %>%
    group_by(sex) %>%
     tally() %>%
     mutate(freq = signif(n*100 / sum(n), 3)) %>%
     mutate(label = c("Men", "Women", "NA")) %>%
     select(sex, label, n, `freq %` = freq))
```

\vspace{75pt}
```{r, echo = F}
sex_tabulated
```

\newpage


**\huge date\_of\_enlistment **
\normalsize
\vspace{12pt}

**Label:** Date of Enlistment 

**Description**: date_of_enlistment is a date variable that reports a person's date of enlistment in dd-mm-yyyy format, as reported in the World War II Army Enlistment Records. 

\vspace{30pt}

```{r}
#min(data_combined$year_of_birth, na.rm=T) # "00"
#max(data_combined$year_of_birth, na.rm=T) # "99"

#data_combined %>%
  #group_by(year_of_birth) %>%
 # tally() %>%
#  print(n=101)

#split_date_of_enlistment_ddmmyy <- function(row) {
 # dd <- substr(row, 1, 2)
#  mm <- substr(row, 3, 4)
#  yy <- substr(row, 5, 6)
#  return(list(dd=dd, mm=mm, yy=yy))}
#res <- lapply(data_combined$date_of_enlistment_ddmmyy, split_date_of_enlistment_ddmmyy) # SLOW
#data_combined$date_of_enlistment_dd <- unlist(res)[attr(unlist(res), "names") == "dd"]
#data_combined$date_of_enlistment_mm <- unlist(res)[attr(unlist(res), "names") == "mm"]
#data_combined$date_of_enlistment_yy <- unlist(res)[attr(unlist(res), "names") == "yy"]
#rm(res)

#min(data_combined$date_of_enlistment_yy, na.rm=T) # "00"
#max(data_combined$date_of_enlistment_yy, na.rm=T) # "99"

#data_combined %>%
  #group_by(date_of_enlistment_yy) %>%
  #tally() %>%
 # print(n=100)

#nrow(subset(data_combined, date_of_enlistment_yy < "38" | date_of_enlistment_yy > "46")) # 41896

## create new date variable 
#data_combined <- data_combined %>% 
 # mutate(date_of_enlistment_yy = as.numeric(date_of_enlistment_yy) + 1900) %>% 
  #mutate(date_enlistment_recode = make_date(date_of_enlistment_yy, date_of_enlistment_mm, date_of_enlistment_dd))

## check dates look reasonable and are in yyyy-mm-dd format (match v1 release)


## visualize distribution (loooks reasonable - identical pattern to v1 release)
data_combined %>%
  filter(date_of_enlistment_yy %in% c(1938:1947)) %>% ## restrict to years around WWII 
  mutate(date_short = format(as.Date(date_enlistment_recode), "%Y-%m")) %>% 
    group_by(date_short) %>% 
    tally() %>% 
    filter(!is.na(date_short)) %>%
  ggplot(aes(x=date_short, y=n, group=1)) +
  geom_line() +
  geom_point() +
  scale_x_discrete(breaks = paste0(seq(1938, 1947, 1), "-01")) +
  theme_minimal(base_size = 15) +
  ggtitle("Enlistment Date (Month and Year Only)") +
  labs(x = "Date (yyyy-mm)", 
       y = "Count") + 
  scale_y_continuous(labels = scales::comma) +
  theme(axis.text.x = element_text(angle = 45))
```



\newpage

**\huge bpl **
\normalsize
\vspace{12pt}

**Label**: Place of Birth 

**Description**: bpl is a numeric variable that reports a person's place of birth, as reported in the World War II Army Enlistment Records.

For a complete list of codes, please see: https://usa.ipums.org/usa-action/variables/BPL#codes_section



\newpage 

**\huge residence\_state **
\normalsize
\vspace{12pt}

**Label**: State of residence

**Description**: residence_state is a character variable that reports a person's state of residence at time of enlistment, as reported in the World War II Army Enlistment Records. This variable is only available for enlistees residing in the US.

For a complete list of codes, please see: https://aad.archives.gov/aad/popup-codelist.jsp?cl_id=2057&dt=893&c_id=24998


\newpage 

**\huge residence\_county **
\normalsize
\vspace{12pt}

**Label**: County of residence

**Description**: residence_county is a character variable that reports a person's county of residence at time of enlistment, as reported in the World War II Army Enlistment Records. This variable is only available for enlistees residing in the US.

For a complete list of codes, please see: https://aad.archives.gov/aad/popup-codelist.jsp?cl_id=2057&dt=893&c_id=24998



\newpage 

**\huge place\_of\_enlistment **
\normalsize
\vspace{12pt}

**Label**: Place of enlistment

**Description**: place_of_enlistment_enlistment is a character variable that reports where a person was enlisted, as reported in the World War II Army Enlistment Records. 

For a complete list of codes, please see: https://aad.archives.gov/aad/popup-codelist.jsp?cl_id=4123&dt=893&c_id=24997




\newpage 

**\huge education **
\normalsize
\vspace{12pt}

**Label**: Education

**Description**: education  is a numeric variable that reports an individual's educational attainment at time of enlistment, as reported in the World War II Army Enlistment Records.
Codes also available at: https://usa.ipums.org/usa-action/variables/EDUC#codes_section

```{r, include = F}
education_tabulated <- knitr::kable(enlistment_records_final %>%
    group_by(education) %>%
     tally() %>%
     mutate(freq = signif(n*100 / sum(n), 3)) %>%
     mutate(label = c("NA/No Schooling", "Nursery school to grade 4" , "Grade 5, 6, 7, or 8", "Grade 9", "Grade 10", "Grade 11", "Grade 12", "1 year of college	", "2 years of college", "3 years of college" ,"NA")) %>%
     select(education, label, n, `freq %` = freq))
```

\vspace{75pt}
```{r, echo = F}
education_tabulated
```
\newpage 

**\huge grade\_code**
\normalsize
\vspace{12pt}

**Label**: Army grade 

**Description**: grade_code is a numeric variable that records a person's grade (rank), as reported in the World War II Army Enlistment Records. 
Code list also is available at: https://aad.archives.gov/aad/popup-codelist.jsp?cl_id=2072&dt=893&c_id=24999


```{r}
grades <- fread('cl_2072.csv')
grade_tabulated <- knitr::kable(enlistment_records_final %>%
  group_by(grade_code) %>% na.omit(grades) %>%
  tally() %>%
  left_join(grades %>% mutate(Code = as.integer(Code)), by = c('grade_code' = 'Code')) %>%
  mutate(freq = signif(n * 100 / sum(n), 3), `freq %` = paste0(freq, "%")) %>%
  select(grade_code, label = Meaning, n, `freq %`))

```


\vspace{30pt}

```{r}
grade_tabulated
```




\newpage 

**\huge branch\_code**
\normalsize
\vspace{12pt}

**Label**: Army branch 

**Description**: branch_code is a numeric variable that records the army branch in which the enlistee served, as reported in the World War II Army Enlistment Records. 

```{r}
branches <- fread('cl_2075.csv')
branch_tabulated <- knitr::kable(enlistment_records_final %>%
                          group_by(branch_code) %>% 
                          tally() %>% 
                          left_join(branches, by=c('branch_code' = 'Code')) %>%
                          mutate(freq = signif(n*100 / sum(n), 3)) %>% 
                          select('branch_code'= branch_code, 'label' = Meaning, n, `freq %` = freq),
                          format='latex', booktabs=T) %>% 
  kable_styling(position = "center") %>% 
  column_spec(column=2, width = '9cm')
```

\vspace{10 pt}

```{r}
# check type of codes vs. enlistment data -- may be mismatched
branch_tabulated
```


\newpage 
\newpage 

**\huge term\_of\_enlistment**
\normalsize
\vspace{12pt}

**Label**: Term of enlistment

**Description**: term_of_enlistment is a numeric variable that reports the length of a person's enlistment or their department of enlistment, as reported in the World War II Army Enlistment Records. 
Codes also available at: https://aad.archives.gov/aad/popup-codelist.jsp?cl_id=2076&dt=893&c_id=24979


```{r}
term <- fread('cl_2076.csv')
toe_tabulated <- knitr::kable(enlistment_records_final %>%
                          group_by(term_of_enlistment) %>% 
                          tally() %>% 
                          left_join(term, by=c('term_of_enlistment' = 'Code')) %>%
                          mutate(freq = signif(n*100 / sum(n), 3)) %>% 
                          select(term_of_enlistment, 'label' = Meaning, n, `freq %` = freq),
                          format='latex', booktabs=T) %>% 
  kable_styling(position = "center") %>% 
  column_spec(column=2, width = '7cm')
```

\vspace{30 pt}

```{r}
toe_tabulated
```








\newpage 

**\huge race**
\normalsize
\vspace{12pt}

**Label**: Race

**Description**: race is a numeric variable that reports an individual's race, as reported in the World War II Army Enlistment Records.

Codes also available at: https://usa.ipums.org/usa-action/variables/RACE#codes_section
```{r, include = F}
race_tabulated <- knitr::kable(enlistment_records_final %>%
    group_by(race) %>%
     tally() %>%
     mutate(freq = signif(n*100 / sum(n), 3)) %>%
     mutate(label = c("White", "Puerto Rican" , "Black", "Native American", "Chinese", "Japanese", "Filipino", "Native Hawaiian", "Other", "NA")) %>%
     select(race, label, n, `freq %` = freq))
```

\vspace{75pt}
```{r, echo = F}
race_tabulated
```


\newpage 

**\huge citizenship**
\normalsize
\vspace{12pt}

**Label**: Citizenship

**Description**: citizenship is a numeric variable that reports an individual's citizenship, as reported in the World War II Army Enlistment Records.


```{r, include = F}
cit_tabulated <- knitr::kable(enlistment_records_final %>%
    group_by(citizenship) %>%
     tally() %>%
     mutate(freq = signif(n*100 / sum(n), 3)) %>%
     mutate(label = c("Non-Citizen", "Citizen", "NA")) %>%
     select(citizenship, label, n, `freq %` = freq))
```

\vspace{75pt}
```{r, echo = F}
cit_tabulated
```

\newpage 

**\huge civilian\_occupation**
\normalsize
\vspace{12pt}

**Label**: civilian occupation

**Description**: civilian_occupation is a numeric variable that reports a person's occupation at time of enlistment, as reported in the World War II Army Enlistment Records. 

For complete codes, please see: https://aad.archives.gov/aad/popup-codelist.jsp?cl_id=3323&dt=893&c_id=24986




\newpage 

**\huge marital\_status**
\normalsize
\vspace{12pt}

**Label**: marital status

**Description**: marital_status is a numeric variable that reports a person's marital status at time of enlistment, as reported in the World War II Army Enlistment Records. 

Codes also available at: https://usa.ipums.org/usa-action/variables/MARST#codes_section


```{r, include = F}
marital_tabulated <- knitr::kable(enlistment_records_final %>%
    group_by(marital_status) %>%
     tally() %>%
     mutate(freq = signif(n*100 / sum(n), 3)) %>%
     mutate(label = c("Married", "Separated", "Divorced", "Widowed", "Single","NA")) %>%
     select(marital_status, label, n, `freq %` = freq))
```

\vspace{75pt}
```{r, echo = F}
marital_tabulated
```





\newpage
**\huge height**
\normalsize
\vspace{12pt}

**Label**: Height

**Description**: height is a numeric variable that reports a person's height in inches at time of enlistment, as reported in the World War II Army Enlistment Records. Height was filtered based on a reasonable range of numbers compliant with what is permissible for enlistment in the military, 60 - 109 inches. 

**Note**: Please be aware that instructions on how to use this field on enlistment punch cards varied over time. This field may encode military occupation or other information. Height may have been recorded through about 1943, but height data recorded after February 1943 are likely unreliable. See the [\textcolor{blue}{National Archives' technical documentation on the Amy Serial Number Electronic File}](https://catalog.archives.gov/OpaAPI/media/1263923/content/arcmedia/electronic-records/rg-064/asnf/100.1ND_NC.pdf?download=false) for further information.

```{r, include = F}
height_plot <- enlistment_records_final %>%
  ggplot(aes(x = height)) +
  geom_histogram(breaks = seq(0,100,1)) +
  theme_minimal(base_size = 15) +
  ggtitle("Height in inches") +
  scale_y_continuous(labels = scales::comma)
  
```

\vspace{75pt}

```{r, echo = F}
height_plot
```



\newpage

**\huge weight\_or\_AGCT\_enlistment**
\normalsize
\vspace{12pt}

**Label**: Weight (pounds) at enlistment or AGCT score

**Description**: weight_or_AGCT_enlistment is a numeric variable that reports a person's weight in pounds at time of enlistment or their Army General Classification Test (AGCT) score, as reported in the World War II Army Enlistment Records. 

**Note**: Please be aware that instructions on how to use this field on enlistment punch cards varied over time. This field was explicitly used for both weight and AGCT scores, but may also encode military occupation or other information. Weight may have been recorded through about 1943 at some enlistment sites, but the recording of AGCT score instead of weight at some enlistment sites likely began in March 1943. Refer to the [\textcolor{blue}{National Archives' technical documentation on the Amy Serial Number Electronic File}](https://catalog.archives.gov/OpaAPI/media/1263923/content/arcmedia/electronic-records/rg-064/asnf/100.1ND_NC.pdf?download=false) for further information.

```{r, include = F}
weight_plot <- enlistment_records_final %>%
  ggplot(aes(x = weight_or_AGCT)) +
  geom_histogram(breaks = seq(0,1000,10)) +
  theme_minimal(base_size = 15) +
  ggtitle("Weight or AGCT Score") +
  scale_y_continuous(labels = scales::comma)
  
```

\vspace{75pt}

```{r, echo = F}
weight_plot
```



\newpage
**\huge component\_enlistment**
\normalsize
\vspace{12pt}

**Label**: Army Component

**Description**:  component_enlistment is a numeric variable that records the component of the army in which an individual served, as recorded in the World War II Army Enlistment Records. 

Codes also available at: https://aad.archives.gov/aad/popup-codelist.jsp?cl_id=2078&dt=929&c_id=25014

```{r}
component_codes <- fread('cl_2078.csv')
component_tabulated <- knitr::kable(enlistment_records_final %>%
                          group_by(component) %>% 
                          tally() %>% 
                          left_join(component_codes, by=c('component' = 'Code')) %>%
                          mutate(freq = signif(n*100 / sum(n), 3)) %>% 
                          select(component, 'label' = Meaning, n, `freq %` = freq),
                          format='latex', booktabs=T) %>% 
  kable_styling(position = "center") %>% 
  column_spec(column=2, width = '7cm')
```

\vspace{75 pt}

```{r}
component_tabulated
```



\newpage

\newpage
**\huge source**
\normalsize
\vspace{12pt}

**Label**: Source of army personnel

**Description**: source is a numeric variable that reports the source of the enlisted person, as reported in the World War II Army Enlistment Records. This denotes if an enlistee was a civilian prior to enlistment, or if they were already enlisted in the National Guard or another faction of the military.
```{r}
source_codes <- fread('cl_2077.csv')
source_tabulated <- knitr::kable(enlistment_records_final %>%
                          group_by(source) %>% 
                          tally() %>% 
                          left_join(component_codes, by=c('source' = 'Code')) %>%
                          mutate(freq = signif(n*100 / sum(n), 3)) %>% 
                          select(source, 'label' = Meaning, n, `freq %` = freq),
                          format='latex', booktabs=T) %>% 
  kable_styling(position = "center") %>% 
  column_spec(column=2, width = '7cm')
```

\vspace{75 pt}

```{r}
source_tabulated
```

Codes also available at: https://aad.archives.gov/aad/popup-codelist.jsp?cl_id=2077&dt=893&c_id=24981

