---
title: "Berkeley Unified WWII Military Enlistment Database Codebook[^updated]"
output:
  pdf_document: default
header-includes: \usepackage{hyperref}
---

[^updated]: Last updated: `r format(Sys.time(), '%d %B, %Y')`


| Page| Variable     | Label                                        |
|----:|:-------------------|:---------------------------------------------|
|   \hyperlink{page.2}{2}  | \hyperlink{page.2}{serial\_number}         |Army Serial Number  |
|   \hyperlink{page.3}{3}  | \hyperlink{page.3}{byear}           |Year of Birth    |
|   \hyperlink{page.4}{4}  | \hyperlink{page.4}{fname}           |First Name                      |
|   \hyperlink{page.5}{5}  |\hyperlink{page.5}{mname}            |Middle Name                      |
|   \hyperlink{page.6}{6}  | \hyperlink{page.6}{lname}          |Last Name                  |
|   \hyperlink{page.7}{7}  |\hyperlink{page.7}{sex}             |Sex              |
|   \hyperlink{page.8}{8}  | \hyperlink{page.8}{date\_of\_enlistment}  |Date of Enlistment     |
|  \hyperlink{page.9}{9} |\hyperlink{page.9}{bpl}              |Place of Birth |
|  \hyperlink{page.10}{10} |\hyperlink{page.10}{residence\_state}           |State of Residence at Enlistment|
|  \hyperlink{page.11}{11} |\hyperlink{page.11}{residence\_county}           |County of Residence at Enlistment|
|  \hyperlink{page.12}{12} |\hyperlink{page.12}{place\_of\_enlistment}          |Place of Enlistment|
|  \hyperlink{page.13}{13} |\hyperlink{page.13}{education}                   |Education |
|  \hyperlink{page.14}{14} |\hyperlink{page.14}{grade\_code}                    |Army Grade (rank)|
|  \hyperlink{page.15}{15} |\hyperlink{page.15}{branch\_code}                     |Army Branch |
|  \hyperlink{page.17}{17} |\hyperlink{page.17}{term\_of\_enlistment}            |Term of Enlistment |
|  \hyperlink{page.18}{18} |\hyperlink{page.18}{race}    |Race  |
|  \hyperlink{page.19}{19} |\hyperlink{page.19}{citizenship}                 |Citizenship  |
|  \hyperlink{page.20}{20} |\hyperlink{page.20}{civilian\_occupation}            |Civilian Occupation |
|  \hyperlink{page.21}{21} |\hyperlink{page.21}{marital\_status}            |Marital Status at Enlistment |
|  \hyperlink{page.22}{22} |\hyperlink{page.22}{height}                |Height at Enlistment (inches) |
|  \hyperlink{page.23}{23} |\hyperlink{page.23}{weight\_before\_march\_1943}                |Weight (pounds)|
|  \hyperlink{page.24}{24} |\hyperlink{page.24}{weight\_or\_AGCT}                |Weight (pounds) or AGCT score|
|  \hyperlink{page.25}{25} |\hyperlink{page.25}{component\_enlistment}             |Army component |
|  \hyperlink{page.26}{26} |\hyperlink{page.26}{source}                          |Source of Army Personnel |

**Summary:** The Berkeley Unified WWII Military Enlistment Database captures 9,039,840 American men and women enlisted in the United States Army and the Women’s Army Auxiliary Corps circa 1938-1947. This dataset is a cleaned and harmonized version of the National Archives and Records Administration's [Electronic Army Serial Number Merged File, ca. 1938 - 1946 (2002)](https://catalog.archives.gov/id/1263923). It contains information on enlistee demographics, military position, and body measurements.



```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
options(kableExtra.auto_format = FALSE, scipen=999)
```

```{r echo = F, results='hide', message=FALSE, warning=FALSE}
## Library Packages
library(tidyverse)
library(data.table)
library(stringi)
library(lubridate)
library(kableExtra)
library(here)

data_path <- '/data/censoc/workspace/enlistment_records/finalenlistment.csv'
resource_file_path <- '/data/josh/CenSoc/censoc_data/enlistment-records-public/code/awenlistmentcode'
path_to_temp_data <- "/data/josh/CenSoc/censoc_data/enlistment-records-public/code/awenlistmentcode/temp3.csv"
#out_path          <- "/data/josh/CenSoc/censoc_data/enlistment-records-public/code/awenlistmentcode"
codebook_path <- here("04_create_enlistment_records/codebook")

setwd(resource_file_path)
#test_file_path <- '/data/josh/CenSoc/censoc_data/enlistment-records-public/code/awenlistmentcode'
enlistment_records_final <- fread(data_path)

```

\newpage

**\huge serial\_number**
\normalsize
\vspace{12pt}

**Label**: Army Serial Number

**Description**: serial_number is a character variable reporting a person’s army serial number, as recorded in the World War II Army Enlistment Records. These numbers are intended to uniquely identify each enlistee, however, due to scanning or transciption errors some serial numbers in this dataset are non-unique.


\newpage 

**\huge byear **
\normalsize
\vspace{12pt}

**Label**: Year of Birth 

**Description**: byear is a numeric variable reporting a person’s year of birth, as recorded in the World War II Army Enlistment Records.

```{r, include = F}
byear_plot <- enlistment_records_final %>%
    group_by(byear) %>%
    summarise(n = n()) %>%
    ggplot(aes(x = byear, y = n)) + 
    geom_line() + 
    geom_point() +
    theme_minimal(base_size = 15) + #replace with a different theme (theme_bw()) if the bbplot package isn't downloaded 
  ggtitle("Year of Birth") + 
  theme(legend.position="bottom") +
  xlab("title") + 
  labs(x = "Year", 
       y = "Count") + 
  scale_y_continuous(labels = scales::comma) +
  scale_x_continuous(breaks = scales::pretty_breaks(n=5)) 
```

\vspace{75pt}

```{r, echo = F}
byear_plot
```




\newpage 
**\huge fname **
\normalsize
\vspace{12pt}

**Label:** First Name

**Description:** fname is a character variable reporting a person’s first name, as recorded in the World War II Army Enlistment Records. This variable was cleaned from raw first names by removing non-alphabetical characters and titles (e.g., Dr.).





\newpage 

**\huge mname **
\normalsize
\vspace{12pt}

**Label**: Middle Name

**Description**: mname is a character variable reporting a person’s middle initial or name, as recorded in the World War II Army Enlistment Records. This variable was cleaned from raw names by removing non-alphabetic characters. Most middle names consist of a single character.

\newpage 

**\huge lname **
\normalsize
\vspace{12pt}

**Label**: Last Name

**Description**: lname is a character variable reporting a person’s last name, as recorded in the World War II Army Enlistment Records. This variable was cleaned and standardized to exclude non-alphabetic characters and whitespace (e.g., "O'Connor" is coded as "oconnor", and "Villa-Lobos" is coded as "villalobos").


\vspace{30pt}




\newpage
**\huge sex **
\normalsize
\vspace{12pt}

**Label**: Sex

**Description**: sex is a numeric variable that reports a person's sex, as inferred by army branch in World War II ensistment records.

**Note**: Sex was not listed on enlistee punchcards. If an individual's numeric branch code or alphabetic branch code in the raw enlistment corresponds to the Women's Army Corps, we declare that person to be female. All other persons are assigned male. There potential for error in this designation due to missing branch information or disagreements between army branch numeric codes, branch alphabetic codes (which we do not publish), and army component information. Thus, sex may not always correspond exactly to army branch and component information in the published data set.

```{r, include = F}
sex_tabulated <- knitr::kable(enlistment_records_final %>%
    group_by(sex) %>%
     tally() %>%
     mutate(freq = signif(n*100 / sum(n), 3)) %>%
     mutate(label = c("Men", "Women")) %>%
     select(sex, label, n, `freq %` = freq))
```

\vspace{75pt}
```{r, echo = F}
sex_tabulated
```

\newpage


**\huge date\_of\_enlistment **
\normalsize
\vspace{12pt}

**Label:** Date of Enlistment 

**Description**: date_of_enlistment is a date variable that reports a person's date of enlistment in dd-mm-yyyy format, as reported in the World War II Army Enlistment Records. 

\vspace{30pt}

```{r}
#min(data_combined$year_of_birth, na.rm=T) # "00"
#max(data_combined$year_of_birth, na.rm=T) # "99"

#data_combined %>%
  #group_by(year_of_birth) %>%
 # tally() %>%
#  print(n=101)

#split_date_of_enlistment_ddmmyy <- function(row) {
 # dd <- substr(row, 1, 2)
#  mm <- substr(row, 3, 4)
#  yy <- substr(row, 5, 6)
#  return(list(dd=dd, mm=mm, yy=yy))}
#res <- lapply(data_combined$date_of_enlistment_ddmmyy, split_date_of_enlistment_ddmmyy) # SLOW
#data_combined$date_of_enlistment_dd <- unlist(res)[attr(unlist(res), "names") == "dd"]
#data_combined$date_of_enlistment_mm <- unlist(res)[attr(unlist(res), "names") == "mm"]
#data_combined$date_of_enlistment_yy <- unlist(res)[attr(unlist(res), "names") == "yy"]
#rm(res)

#min(data_combined$date_of_enlistment_yy, na.rm=T) # "00"
#max(data_combined$date_of_enlistment_yy, na.rm=T) # "99"

#data_combined %>%
  #group_by(date_of_enlistment_yy) %>%
  #tally() %>%
 # print(n=100)

#nrow(subset(data_combined, date_of_enlistment_yy < "38" | date_of_enlistment_yy > "46")) # 41896

## create new date variable 
#data_combined <- data_combined %>% 
 # mutate(date_of_enlistment_yy = as.numeric(date_of_enlistment_yy) + 1900) %>% 
  #mutate(date_enlistment_recode = make_date(date_of_enlistment_yy, date_of_enlistment_mm, date_of_enlistment_dd))

## check dates look reasonable and are in yyyy-mm-dd format (match v1 release)


## visualize distribution (loooks reasonable - identical pattern to v1 release)
enlistment_records_final %>%
  filter(lubridate::year(date_of_enlistment) %in% c(1938:1947)) %>% ## restrict to years around WWII 
  mutate(date_short = format(as.Date(date_of_enlistment), "%Y-%m")) %>% 
    group_by(date_short) %>% 
    tally() %>% 
    filter(!is.na(date_short)) %>%
  ggplot(aes(x=date_short, y=n, group=1)) +
  geom_line() +
  geom_point() +
  scale_x_discrete(breaks = paste0(seq(1938, 1947, 1), "-01")) +
  theme_minimal(base_size = 15) +
  ggtitle("Enlistment Date (Month and Year Only)") +
  labs(x = "Date (yyyy-mm)", 
       y = "Count") + 
  scale_y_continuous(labels = scales::comma) +
  theme(axis.text.x = element_text(angle = 45))


```



\newpage

**\huge bpl **
\normalsize
\vspace{12pt}

**Label**: Place of Birth 

**Description**: bpl is a numeric variable that reports a person's place of birth, as reported in the World War II Army Enlistment Records.

For a complete list of codes, please see: https://usa.ipums.org/usa-action/variables/BPL#codes_section



\newpage 

**\huge residence\_state **
\normalsize
\vspace{12pt}

**Label**: State of residence

**Description**: residence_state is a character variable that reports a person's state of residence at time of enlistment, as reported in the World War II Army Enlistment Records. This variable is only available for enlistees residing in the US.

For a complete list of codes, please see: https://aad.archives.gov/aad/popup-codelist.jsp?cl_id=2046&dt=893&c_id=24996


\newpage 

**\huge residence\_county **
\normalsize
\vspace{12pt}

**Label**: County of residence

**Description**: residence_county is a character variable that reports a person's county of residence at time of enlistment, as reported in the World War II Army Enlistment Records. This variable is only available for enlistees residing in the US.

For a complete list of codes, please see: https://aad.archives.gov/aad/popup-codelist.jsp?cl_id=2057&dt=893&c_id=24998



\newpage 

**\huge place\_of\_enlistment **
\normalsize
\vspace{12pt}

**Label**: Place of enlistment

**Description**: place_of_enlistment_enlistment is a character variable that reports where a person was enlisted, as reported in the World War II Army Enlistment Records. 

For a complete list of codes, please see: https://aad.archives.gov/aad/popup-codelist.jsp?cl_id=4123&dt=893&c_id=24997




\newpage 

**\huge education **
\normalsize
\vspace{12pt}

**Label**: Education

**Description**: education  is a numeric variable that reports an individual's educational attainment at time of enlistment, as reported in the World War II Army Enlistment Records.
Codes also available at: https://usa.ipums.org/usa-action/variables/EDUC#codes_section

**Note**: Exact number of years of schooling is unknown for enlistees who did not complete at least one year of high school. Such enlistees were given originally given a code meaning "grammar school", though it is unknown if this corresponds to the modern equivalent of a 5th, 6th, or 8th grade education. Many enlistees with this code likely had little to no actual education. Before the attack on Pearl Harbor, literacy was required of army inductees, but this requirement relaxed as the war effort increased ([\textcolor{blue}{Brandt, 2004}](https://doi.org/10.2307/4140731)). By 1942, the Army accepted illiterate men if they were deemed trainable, though academic standards for admission increased again in 1944 and 1945.


```{r, include = F}
education_tabulated <- knitr::kable(enlistment_records_final %>%
    group_by(education) %>%
     tally() %>%
     mutate(freq = signif(n*100 / sum(n), 3)) %>%
     mutate(label = c( "Nursery school to grade 4" , "Grade 9", "Grade 10", "Grade 11", "Grade 12", "1 year of college	", "2 years of college", "3 years of college" ,"	4 years of college", "5+ years of college", "NA")) %>%
     select(education, label, n, `freq %` = freq))
```

\vspace{75pt}
```{r, echo = F}
education_tabulated
```
\newpage 

**\huge grade\_code**
\normalsize
\vspace{12pt}

**Label**: Army grade 

**Description**: grade_code is a numeric variable that records a person's grade (rank), as reported in the World War II Army Enlistment Records. 
Code list also is available at: https://aad.archives.gov/aad/popup-codelist.jsp?cl_id=2072&dt=893&c_id=24999


```{r, include = F}
grades <- fread('cl_2072.csv')
grades <- grades %>% mutate(Code = ifelse(Code %in% c("M", "N", "O"), NA_integer_, as.integer(Code))) %>% 
  drop_na(Code)
grade_tabulated <- knitr::kable(enlistment_records_final %>%
  group_by(grade_code) %>% 
  tally() %>%
  left_join(grades, by = c('grade_code' = 'Code')) %>%
  mutate(freq = signif(n * 100 / sum(n), 3)) %>%
  select(grade_code, label = Meaning, n, `freq %`=freq))

```


\vspace{30pt}

```{r, echo = F}
grade_tabulated
```




\newpage 

**\huge branch\_code**
\normalsize
\vspace{12pt}

**Label**: Army branch 

**Description**: branch_code is a numeric variable that records the army branch in which the enlistee served, as reported in the World War II Army Enlistment Records. 

```{r, include = F}
branches <- fread('cl_2075.csv')
branch_tabulated <- knitr::kable(enlistment_records_final %>%
                          group_by(branch_code) %>% 
                          tally() %>% 
                          left_join(branches, by=c('branch_code' = 'Code')) %>%
                          mutate(freq = signif(n*100 / sum(n), 3)) %>% 
                          select('branch_code'= branch_code, 'label' = Meaning, n, `freq %` = freq),
                          format='latex', booktabs=T) %>% 
  kable_styling(position = "center") %>% 
  column_spec(column=2, width = '9cm')
```

\vspace{75 pt}

```{r, echo = F}
# check type of codes vs. enlistment data -- may be mismatched
branch_tabulated
```


\newpage 

**\huge term\_of\_enlistment**
\normalsize
\vspace{12pt}

**Label**: Term of enlistment

**Description**: term_of_enlistment is a numeric variable that reports the length of a person's enlistment or their department of enlistment, as reported in the World War II Army Enlistment Records. 
Codes also available at: https://aad.archives.gov/aad/popup-codelist.jsp?cl_id=2076&dt=893&c_id=24979


```{r, include = F}
toe_tabulated <- knitr::kable(enlistment_records_final %>%
    group_by(term_of_enlistment) %>%
     tally() %>%
     mutate(freq = signif(n*100 / sum(n), 3)) %>%
     mutate(label = c("One year enlistment	", "Enlistment for Alaska	" , "Enlistment for Hawaiian Department	", " Enlistment for assignment to another corps area", "Enlistment for the duration of the War or other emergency, plus six months, subject to the discretion of the President or otherwise according to law	", "Enlistment for the Panama Canal Department	", "Enlistment for the Philippine Department	", "Enlistment for the Puerto Rican Department	 ", "Three year enlistment	", "NA")) %>%
     select(term_of_enlistment, label, n, `freq %` = freq)) %>% 
  column_spec(column=2, width = '6cm')
```

\vspace{30 pt}

```{r, echo = F}
toe_tabulated
```








\newpage 

**\huge race**
\normalsize
\vspace{12pt}

**Label**: Race

**Description**: race is a numeric variable that reports an individual's race, as reported in the World War II Army Enlistment Records.

Codes also available at: https://usa.ipums.org/usa-action/variables/RACE#codes_section
```{r, include = F}
race_tabulated <- knitr::kable(enlistment_records_final %>%
    group_by(race) %>%
     tally() %>%
     mutate(freq = signif(n*100 / sum(n), 3)) %>%
     mutate(label = c("White", "Puerto Rican" , "Black", "Native American", "Chinese", "Japanese", "Filipino", "Native Hawaiian", "Other", "NA")) %>%
     select(race, label, n, `freq %` = freq))
```

\vspace{75pt}
```{r, echo = F}
race_tabulated
```


\newpage 

**\huge citizenship**
\normalsize
\vspace{12pt}

**Label**: Citizenship

**Description**: citizenship is a numeric variable that reports an individual's citizenship, as reported in the World War II Army Enlistment Records.


```{r, include = F}
cit_tabulated <- knitr::kable(enlistment_records_final %>%
    group_by(citizenship) %>%
     tally() %>%
     mutate(freq = signif(n*100 / sum(n), 3)) %>%
     mutate(label = c("Non-Citizen", "Citizen", "NA")) %>%
     select(citizenship, label, n, `freq %` = freq))
```

\vspace{75pt}
```{r, echo = F}
cit_tabulated
```

\newpage 

**\huge civilian\_occupation**
\normalsize
\vspace{12pt}

**Label**: civilian occupation

**Description**: civilian_occupation is a numeric variable that reports a person's occupation at time of enlistment, as reported in the World War II Army Enlistment Records. 

For complete codes, please see: https://aad.archives.gov/aad/popup-codelist.jsp?cl_id=3323&dt=893&c_id=24986




\newpage 

**\huge marital\_status**
\normalsize
\vspace{12pt}

**Label**: marital status

**Description**: marital_status is a numeric variable that reports a person's marital status at time of enlistment, as reported in the World War II Army Enlistment Records. 

Codes also available at: https://usa.ipums.org/usa-action/variables/MARST#codes_section


```{r, include = F}
marital_tabulated <- knitr::kable(enlistment_records_final %>%
    group_by(marital_status) %>%
     tally() %>%
     mutate(freq = signif(n*100 / sum(n), 3)) %>%
     mutate(label = c("Married", "Separated", "Divorced", "Widowed", "Single","NA")) %>%
     select(marital_status, label, n, `freq %` = freq))
```

\vspace{75pt}
```{r, echo = F}
marital_tabulated
```





\newpage
**\huge height**
\normalsize
\vspace{12pt}

**Label**: Height

**Description**: height is a numeric variable that reports a person's height in inches at time of enlistment, as reported in the World War II Army Enlistment Records. Height was filtered based on the range of permissible heights for enlistment in the military, 60 - 78 inches ([\textcolor{blue}{Foster et al, 1967}](https://apps.dtic.mil/sti/pdfs/ADA291761.pdf)). 

**Note**: Please be aware that instructions on how to use this field on enlistment punch cards varied over time. This field may encode military occupation or other information depending on time and site of enlistment. We have removed height data for men who enlisted in March of 1943 or later, as the distribution of this variable changes significantly at that point. However, it is possible that other data remains in this field. See the [\textcolor{blue}{National Archives' technical documentation on the Amy Serial Number Electronic File}](https://catalog.archives.gov/id/1263923) for further information.

```{r, include = F}
height_plot <- enlistment_records_final %>%
  ggplot(aes(x = height)) +
  geom_bar(stat = "count") +
  #geom_histogram(breaks = seq(60,78,1)) +
  theme_minimal(base_size = 15) +
  ggtitle("Height in inches") +
  scale_y_continuous(labels = scales::comma)
  
```

\vspace{75pt}

```{r, echo = F}
height_plot
```

\newpage
**\huge weight_before_march_1943**
\normalsize
\vspace{12pt}


**Description**: weight_before_march_1943 is a numeric variable that reports a person's weight in pounds at time of enlistment as reported in the World War II Army Enlistment Records. This variable is only available for enlistees who enlisted before the March of 1943. Weight is filtered to a minimum of 105 pounds, the minimum permissible weight for a 60-inch-tall enlistee ([\textcolor{blue}{Foster et al, 1967}](https://apps.dtic.mil/sti/pdfs/ADA291761.pdf)). While there was no clear maximum body weight for army enlistees, we have removed values of over 500.

**Note**: On enlistment punchcards, the field used to record weight was also used for Army General Classification Test (AGCT) scores, and may also encode military occupation or other information. Weight may have been recorded through 1943 at some enlistment sites, but the recording of AGCT score instead of weight at many enlistment sites likely began in March 1943. Because of this ambiguity, we only make this variable available for enlistees who enlisted before 1943. 
Refer to the [\textcolor{blue}{National Archives' technical documentation on the Amy Serial Number Electronic File}](https://catalog.archives.gov/id/1263923) for further information. For enlistees who enlisted starting in March 1943, we have created an additional variable called **weight_or_AGCT**.

```{r, include = F}
weight_plot <- enlistment_records_final %>%
  ggplot(aes(x = weight_before_march_1943)) +
  geom_histogram(breaks = seq(100,500,5)) +
  theme_minimal(base_size = 15) +
  ggtitle("Weight") +
  scale_y_continuous(labels = scales::comma)
  
```

\vspace{75pt}

```{r, echo = F}
weight_plot
```





\newpage

**\huge weight\_or\_AGCT**
\normalsize
\vspace{12pt}

**Label**: Weight (pounds) at enlistment or AGCT score

**Description**: weight_or_AGCT is a numeric variable that reports a person's weight in pounds at time of enlistment or their Army General Classification Test (AGCT) score, as reported in the World War II Army Enlistment Records. This variable is only available for enlistees who enlisted in January 1943 or later.  We have removed zeros and values above 500, which presumably exceed the maximum realistic body weights and AGCT scores, but otherwise leave the data in this field unfiltered and unsorted.

**Note**: On enlistment punchcards, the field used to record weight was also used for Army General Classification Test (AGCT) scores, and may also encode military occupation or other information. Weight may have been recorded through 1943 at some enlistment sites, but the recording of AGCT score instead of weight at many enlistment sites likely began in March 1943. Because of this ambiguity, we created this variable for persons enlisted in March 1943 or later, which may contain weight or AGCT or other information.  
Refer to the [\textcolor{blue}{National Archives' technical documentation on the Amy Serial Number Electronic File}](https://catalog.archives.gov/OpaAPI/media/1263923/content/arcmedia/electronic-records/rg-064/asnf/100.1ND_NC.pdf?download=false) for further information. For enlistees prior to January 1943, we used the information recorded in this field to create the **weight_before_march_1943** variable.

```{r, include = F}
weight_agct_plot <- enlistment_records_final %>%
  ggplot(aes(x = weight_or_AGCT)) +
  geom_histogram(breaks = seq(0,500,5)) +
  theme_minimal(base_size = 15) +
  ggtitle("Weight or AGCT Score") +
  scale_y_continuous(labels = scales::comma)
  
```

\vspace{75pt}

```{r, echo = F}
weight_agct_plot
```


\newpage

**\huge component**
\normalsize
\vspace{12pt}

**Label**: Army Component

**Description**:  component is a numeric variable that records the component of the army in which an individual served, as recorded in the World War II Army Enlistment Records. 

Codes also available at: https://aad.archives.gov/aad/popup-codelist.jsp?cl_id=2078&dt=929&c_id=25014
\vspace{75 pt}

```{r}
component_tabulated <- knitr::kable(enlistment_records_final %>%
    group_by(component) %>%
     tally() %>%
     mutate(freq = signif(n*100 / sum(n), 3)) %>%
     mutate(label = c("Regular Army", "Regular Army Reserve","Reserves", "National Guard ", "Philippine Scouts", "Army of the United States", "Selectees", "Women’s Army Corps","NA")) %>%
     select(component, label, n, `freq %` = freq))

```

```{r}
component_tabulated
```



\newpage
**\huge source**
\normalsize
\vspace{12pt}

**Label**: Source of army personnel

**Description**: source is a numeric variable that reports the source of the enlisted person, as reported in the World War II Army Enlistment Records. This denotes if an enlistee was a civilian prior to enlistment, or if they were already enlisted in the National Guard or another faction of the military.
Codes also available at: https://aad.archives.gov/aad/popup-codelist.jsp?cl_id=2077&dt=893&c_id=24981

\vspace{75 pt}
```{r}
source_tabulated <- knitr::kable(enlistment_records_final %>%
    group_by(source) %>%
     tally() %>%
     mutate(freq = signif(n*100 / sum(n), 3)) %>%
     mutate(label = c("Civil Life", "Enlisted Man, Regular Army, within 3 months of Discharge or former WAAC Auxiliary","National Guard in Federal Service, within 3 months of Discharge", "Enlisted Man, Regular Army, after 3 months of Discharge", "National Guard in Federal Service, after 3 months of Discharge ", "Enlisted Reserve or Medical Administrative Corps (MAC) Officer", "Enlisted Man, Philippine Scout or recall to AD of an enlisted man who had been transferred to the ERC", "National Guard	", "Returned to Military Control, Surrendered", "Returned to Military Control Apprehended", "Selective Service within 3 months of Discharge", "Selective Service after 3 months of Discharge","NA")) %>%
     select(source, label, n, `freq %` = freq))

```


```{r, echo = F}
source_tabulated
```


