---
title: "Read in Original Numident Files"
author: Casey Breen
---

Background: In 2013, the Social Security Administration transferred a set of Numident records to the
National Archives (NARA). In 2019, we obtained the NARA Numident records, along with
their accompanying documentation. The NARA Numident records are a subset of the records
in the complete Numident. The NARA Numident records contain three types of entries:
applications, claims, and deaths. NARA delivered each set of entries seperately as a set 20
fixed-width .txt files (3 x 20 = 60 files in total).

For each of the three entry types, the below code reads in the 20 fixed-width .txt files using column
position specified in the record layout documents. The code appends the 20 files into a single
file. The result is three files, one for the applications, claims, and deaths. 

Column positions and variable names were obtained from the NARA documentation website: https://aad.archives.gov/aad/popup-tech-info.jsp?s=5057

The below functions read in all variables with the colun position and variables names specified in the NARA record layout documents.The PDF files were first converted into a spreadsheet in order facilitate writing the code to read in the fixed-width files: 
/bcourses/documentation/numident_documentation/numident_file_layouts_to_create_r_code.csv

Variable names are consistent with the original "field names" with two exceptions: 

1) All dashes were replaced with underscores (e.g. "-" were replaced with "_")
2) The seven variables titled "FILLER" were renamed ("Filler_1, Filler_2, ... , Filler_7)".


```{r}
# library(devtools) 
# install_github("caseybreen/censocdev")
```

Library Packages
```{r}
library(here)
library(tidyverse)
library(data.table)

source(here("R/read_death_fwf.R"))
source(here("R/read_app_fwf.R"))
source(here("R/read_claim_fwf.R"))
```


```{r}
## read in fixed-width death records
deaths <- read_death_fwf(path = "/global/scratch/p2p3/pl1_demography/censoc/input_data/nara_numident/death_records/records/")

## write out death records
fwrite(x = deaths, file = "/global/scratch/p2p3/pl1_demography/censoc_internal/data/numident/1_numident_files_with_original_varnames/deaths_original.csv")

## read in fixed-width appliation records
apps <- read_app_fwf(path = "/global/scratch/p2p3/pl1_demography/censoc/input_data/nara_numident/application_records/records/")

## write out application records
fwrite(apps, "/global/scratch/p2p3/pl1_demography/censoc_internal/data/numident/1_numident_files_with_original_varnames/applications_original.csv")

## read in claims fixed-width application records 
claims <- read_claim_fwf(path = "/global/scratch/p2p3/pl1_demography/censoc/input_data/nara_numident/claim_records/records/")

## write out application records
fwrite(claims, "/global/scratch/p2p3/pl1_demography/censoc_internal/data/numident/1_numident_files_with_original_varnames/claims_original.csv")
```


