---
title: "05_clean_names_bunmd"
output: html_document
date: "2023-10-05"
---

#This Rmd is intended to clean the names of the bunmd file, resulting in the creation of cleaned name variables for an individual, their mother, and their father. 
#cleaned name variables: fname, middle, lname, father_fname, father_middle, father_lname, mother_fname, mother_middle, mother_lname
#uncleaned name variables: fname_raw, middle_raw, lname_raw, father_fname_raw, father_middle_raw, father_lname_raw, mother_fname_raw, mother_middle_raw, mother_lname_raw

```{r}
library(censocdev)
library(ipumsr)
```


```{r, echo = FALSE}
#read in data
 bunmd <- fread("/censoc/data/numident/4_berkeley_unified_mortality_database/bunmd.csv")
sampled_data <- bunmd[sample(nrow(bunmd), 1000), ]
fwrite(sampled_data, "/censoc/data/numident/4_berkeley_unified_mortality_database/bunmd_sampled.csv")

bunmd %>% group_by(sex) %>% count()
```
# Read titles and nicknames
```{r}
tan_path          <-  "/data/josh/CenSoc/censoc_data/enlistment-records-public/data/codes/titles-and-nicknames"
titles <- fread(paste(tan_path, "titles.csv", sep="/"))
male_nicknames <- fread(paste(tan_path, "male_nicknames.csv", sep="/"))
female_nicknames <- fread(paste(tan_path, "female_nicknames.csv", sep="/"))
nicknames <- data.table(rbind(male_nicknames, female_nicknames))
old_nicknames <- fread(paste(tan_path, "old_nicknames.csv", sep="/"))
```

```{r}
#look at data 
head(bunmd)
colnames(bunmd)
bunmd %>% count(lname)
```
# Read in functions for name cleaning
```{r}
#clean fname ABE Method https://github.com/uguryi/abeR/blob/master/R/clean-names.R

# Define remove_titles
#' @export

# Define remove_titles
#' @export
remove_titles <- function(row) {
  fname_clean <- row
  for (title in titles$title) {
    pattern <- paste0(" ", title, " ")
    fname_clean <- stringi::stri_replace_all_regex(fname_clean, pattern, "")
  }
  return(fname_clean)
}


clean_names <- function(in_path, file_name, lname_col, fname_col, sex_col, other_cols, out_path, middle) {

  # Read data
  path_to_data <- paste(in_path, file_name, sep="/")
  data <- readr::read_csv(path_to_data)

  # Copy raw names before processing
  data$lname_clean <- data[[lname_col]]
  data$fname_clean <- data[[fname_col]]

  # Concatenate middle name to first name if middle name exists
  if (middle != "") {
    concat_middle_to_first <- function(row) {
      if (is.na(row["fname_clean"])) {first <- NA}
      else if (is.na(row[[middle]])) {first <- row["fname_clean"]}
      else {first <- paste(row["fname_clean"], row[[middle]])}
      return(first)
    }
    data$fname_clean <- apply(data, 1, concat_middle_to_first) # SLOW
    data$mname_raw <- data[[middle]]
    data[[middle]] <- NULL
  }
  # NOTES
  # 1. This statement is here to make sure that there aren't any inconsistencies
  #    in relation to middle initials. For example, in the case of CenSoc, BUNMD already
  #    has middle initials but the way the ABE algorithm processes first names and
  #    initials are sometimes different from the way BUNMD stores first names and initials.
  #    ABE turns "A J" into "A J" as first name and "J" as initial, while in BUNMD "A J"
  #    would be first name "A" and initial "J".

  # Convert unicode to ascii to remove accents
  data$lname_clean <- iconv(data$lname_clean, to="ASCII//TRANSLIT")
  data$fname_clean <- iconv(data$fname_clean, to="ASCII//TRANSLIT")

  # Convert to lower case
  data$lname_clean <- tolower(data$lname_clean)
  data$fname_clean <- tolower(data$fname_clean)

  # Trim whitespace
  data$lname_clean <- stringr::str_trim(data$lname_clean)
  data$fname_clean <- stringr::str_trim(data$fname_clean)

  # Replace . with " "
  data$lname_clean <- gsub("\\.", " ", data$lname_clean)
  data$fname_clean <- gsub("\\.", " ", data$fname_clean)

  # Keep only letters and spaces (accents don't work)
  data$lname_clean <- gsub("[^a-z ]", "", data$lname_clean)
  data$fname_clean <- gsub("[^a-z ]", "", data$fname_clean)

  # Get rid of titles
  #data$fname_clean <- paste0(" ", data$fname_clean)
  #data$fname_clean <- remove_titles(data$fname_clean)

  # Fix two-word names (e.g., mc donnell --> mcdonnell)
  data$lname_clean <- paste0(" ", data$lname_clean)
  data$lname_clean <- gsub(" st ",    " st",    data$lname_clean)
  data$lname_clean <- gsub(" ste ",   " ste",   data$lname_clean)
  data$lname_clean <- gsub(" saint ", " saint", data$lname_clean)
  data$lname_clean <- gsub(" virg ",  " virg",  data$lname_clean)
  data$lname_clean <- gsub(" mac ",   " mac",   data$lname_clean)
  data$lname_clean <- gsub(" mc ",    " mc",    data$lname_clean)
  data$lname_clean <- stringr::str_trim(data$lname_clean)
  data$fname_clean <- gsub(" st ",    " st",    data$fname_clean)
  data$fname_clean <- gsub(" ste ",   " ste",   data$fname_clean)
  data$fname_clean <- gsub(" saint ", " saint", data$fname_clean)
  data$fname_clean <- gsub(" virg ",  " virg",  data$fname_clean)
  data$fname_clean <- gsub(" mac ",   " mac",   data$fname_clean)
  data$fname_clean <- gsub(" mc ",    " mc",    data$fname_clean)
  data$fname_clean <- stringr::str_trim(data$fname_clean)
  # NOTES
  # 1. Many more words can be added here, such as:
  #    la, van, du, de, del, o, pal, le, de la

   # Split first name (i.e., deal with middle names and initials)
  res <- lapply(data$fname_clean, split_first_name) # SLOW (~60s with ~3M obs)
  data$first  <- unlist(res)[attr(unlist(res),"names") == "first"]
  data$mname <- unlist(res)[attr(unlist(res),"names") == "middle"]
  rm(res)
  
  # Merge nicknames
  data_final <- merge(data, nicknames, by.x = c("first", sex_col),
                      by.y = c("nickname", "sex"), all.x = TRUE) # SLOW (~30s with ~3M obs)
  data_final$fname_final <- apply(data_final, 1, merge_names_no_middle) # SLOW (~45s with ~3M obs)

  # Keep only the first multicharacter part of lname_clean
  res <- lapply(data_final$lname_clean, split_first_name) # SLOW (~60s with ~3M obs)
  data_final$lname_final <- unlist(res)[attr(unlist(res),"names") == "first"]
  rm(res)
   # Reorder columns, drop unnecessary ones (fname_clean, lname_clean, first, name)
  if (middle == "") {
    cols_to_keep <- c("fname_final", "mname", "lname_final", sex_col, fname_col, lname_col, other_cols)
    data_final <- data_final[,cols_to_keep]
    colnames(data_final) <- c("fname", "mname", "lname", "sex", "fname_raw", "lname_raw", other_cols)
  } else {
    cols_to_keep <- c("fname_final", "mname", "lname_final", sex_col, fname_col, "mname_raw", lname_col, other_cols)
    data_final <- data_final[,cols_to_keep]
    colnames(data_final) <- c("fname", "mname", "lname", "sex", "fname_raw", "mname_raw", "lname_raw", other_cols)
  }
 # Do a final trim
  data_final$lname <- stringr::str_trim(data_final$lname)
  data_final$fname <- stringr::str_trim(data_final$fname)
  data_final$mname <- stringr::str_trim(data_final$mname)

  # Save dataset
  dir.create(out_path, showWarnings = FALSE)
  path_to_processed_data <- paste(out_path, gsub(".csv", "_cleaned.csv", file_name), sep="/")
  readr::write_csv(data_final, path_to_processed_data)

  # Clean up after yourself
  rm(list = ls())
}


```

#Apply code to BUNMD fname lname mname
```{r}
in_path <-"/censoc/data/numident/4_berkeley_unified_mortality_database"
lname_col <- "lname"
fname_col <- "fname"
sex_col <- "sex"
out_path      <- "/censoc/data/numident/4_berkeley_unified_mortality_database"
other_cols <- c("ssn","zip_residence","byear","dyear","socstate","bmonth","dmonth","bday","dday",
                "number_apps","race_first","race_last","race_last_cyear","race_first_cyear",
                "race_last_cmonth","race_first_cmonth","race_change","number_claims","bpl",
                "father_fname","father_mname","father_lname","mother_fname","mother_mname",
                "mother_lname","age_first_application","death_age","weight","ccweight",
                "bpl_string","socstate_string")
middle <- "mname"
file_name <- "bunmd.csv"
clean_names(in_path, file_name, lname_col, fname_col,
            sex_col, other_cols, out_path, middle)

```
#Check cleaned dataset 
```{r}
bunmd_cleaned <- fread("/censoc/data/numident/4_berkeley_unified_mortality_database/bunmd_cleaned.csv")
```



#REPEAT W FATHER NAMES - create function clean_names_father
```{r}
#clean fname ABE Method https://github.com/uguryi/abeR/blob/master/R/clean-names.R

clean_names_father <- function(in_path, file_name, lname_col, fname_col, sex_col, other_cols, out_path, middle) {

  # Read data
  path_to_data <- paste(in_path, file_name, sep="/")
  data <- readr::read_csv(path_to_data)

  # Copy raw names before processing
  data$lname_clean <- data[[lname_col]]
  data$fname_clean <- data[[fname_col]]

  # Concatenate middle name to first name if middle name exists
  if (middle != "") {
    concat_middle_to_first <- function(row) {
      if (is.na(row["fname_clean"])) {first <- NA}
      else if (is.na(row[[middle]])) {first <- row["fname_clean"]}
      else {first <- paste(row["fname_clean"], row[[middle]])}
      return(first)
    }
    data$fname_clean <- apply(data, 1, concat_middle_to_first) # SLOW
    data$mname_raw <- data[[middle]]
    data[[middle]] <- NULL
  }
  # NOTES
  # 1. This statement is here to make sure that there aren't any inconsistencies
  #    in relation to middle initials. For example, in the case of CenSoc, BUNMD already
  #    has middle initials but the way the ABE algorithm processes first names and
  #    initials are sometimes different from the way BUNMD stores first names and initials.
  #    ABE turns "A J" into "A J" as first name and "J" as initial, while in BUNMD "A J"
  #    would be first name "A" and initial "J".

  # Convert unicode to ascii to remove accents
  data$lname_clean <- iconv(data$lname_clean, to="ASCII//TRANSLIT")
  data$fname_clean <- iconv(data$fname_clean, to="ASCII//TRANSLIT")

  # Convert to lower case
  data$lname_clean <- tolower(data$lname_clean)
  data$fname_clean <- tolower(data$fname_clean)

  # Trim whitespace
  data$lname_clean <- stringr::str_trim(data$lname_clean)
  data$fname_clean <- stringr::str_trim(data$fname_clean)

  # Replace . with " "
  data$lname_clean <- gsub("\\.", " ", data$lname_clean)
  data$fname_clean <- gsub("\\.", " ", data$fname_clean)

  # Keep only letters and spaces (accents don't work)
  data$lname_clean <- gsub("[^a-z ]", "", data$lname_clean)
  data$fname_clean <- gsub("[^a-z ]", "", data$fname_clean)

  # Get rid of titles
  #data$fname_clean <- paste0(" ", data$fname_clean)
  #data$fname_clean <- remove_titles(data$fname_clean)

  # Fix two-word names (e.g., mc donnell --> mcdonnell)
  data$lname_clean <- paste0(" ", data$lname_clean)
  data$lname_clean <- gsub(" st ",    " st",    data$lname_clean)
  data$lname_clean <- gsub(" ste ",   " ste",   data$lname_clean)
  data$lname_clean <- gsub(" saint ", " saint", data$lname_clean)
  data$lname_clean <- gsub(" virg ",  " virg",  data$lname_clean)
  data$lname_clean <- gsub(" mac ",   " mac",   data$lname_clean)
  data$lname_clean <- gsub(" mc ",    " mc",    data$lname_clean)
  data$lname_clean <- stringr::str_trim(data$lname_clean)
  data$fname_clean <- gsub(" st ",    " st",    data$fname_clean)
  data$fname_clean <- gsub(" ste ",   " ste",   data$fname_clean)
  data$fname_clean <- gsub(" saint ", " saint", data$fname_clean)
  data$fname_clean <- gsub(" virg ",  " virg",  data$fname_clean)
  data$fname_clean <- gsub(" mac ",   " mac",   data$fname_clean)
  data$fname_clean <- gsub(" mc ",    " mc",    data$fname_clean)
  data$fname_clean <- stringr::str_trim(data$fname_clean)
  # NOTES
  # 1. Many more words can be added here, such as:
  #    la, van, du, de, del, o, pal, le, de la

   # Split first name (i.e., deal with middle names and initials)
  res <- lapply(data$fname_clean, split_first_name) # SLOW (~60s with ~3M obs)
  data$first  <- unlist(res)[attr(unlist(res),"names") == "first"]
  data$mname <- unlist(res)[attr(unlist(res),"names") == "middle"]
  rm(res)
  
  # Merge nicknames
  data_final <- merge(data, nicknames, by.x = c("first", sex_col),
                      by.y = c("nickname", "sex"), all.x = TRUE) # SLOW (~30s with ~3M obs)
  data_final$fname_final <- apply(data_final, 1, merge_names_no_middle) # SLOW (~45s with ~3M obs)

  # Keep only the first multicharacter part of lname_clean
  res <- lapply(data_final$lname_clean, split_first_name) # SLOW (~60s with ~3M obs)
  data_final$lname_final <- unlist(res)[attr(unlist(res),"names") == "first"]
  rm(res)
   # Reorder columns, drop unnecessary ones (fname_clean, lname_clean, first, name)
  if (middle == "") {
    cols_to_keep <- c("fname_final", "mname", "lname_final", sex_col, fname_col, lname_col, other_cols)
    data_final <- data_final[,cols_to_keep]
    colnames(data_final) <- c("fname", "mname", "lname", "sex", "fname_raw", "lname_raw", other_cols)
  } else {
    cols_to_keep <- c("fname_final", "mname", "lname_final", sex_col, fname_col, "mname_raw", lname_col, other_cols)
    data_final <- data_final[,cols_to_keep]
   colnames(data_final) <- c("father_fname", "father_mname", "father_lname", "father_sex", "father_fname_raw", "father_mname_raw", "father_lname_raw", other_cols)
  }
 # Do a final trim
  data_final$lname <- stringr::str_trim(data_final$lname)
  data_final$fname <- stringr::str_trim(data_final$fname)
  data_final$mname <- stringr::str_trim(data_final$mname)

  # Save dataset
  dir.create(out_path, showWarnings = FALSE)
  path_to_processed_data <- paste(out_path, gsub(".csv", "_cleaned.csv", file_name), sep="/")
  readr::write_csv(data_final, path_to_processed_data)

  # Clean up after yourself
  rm(list = ls())
}

```

#apply function to clean father names
```{r}
#add sex column for father 
bunmd_cleaned$father_sex <- 1
#write out new dataset
fwrite(bunmd_cleaned, "/censoc/data/numident/4_berkeley_unified_mortality_database/bunmd_f.csv")


in_path_f <-"/censoc/data/numident/4_berkeley_unified_mortality_database"
lname_col_f <- "father_lname"
fname_col_f <- "father_fname"
sex_col_f <- "father_sex" 
out_path_f      <- "/censoc/data/numident/4_berkeley_unified_mortality_database"
other_cols_f <- c("fname","mname","lname","sex","fname_raw","mname_raw","lname_raw", "ssn","zip_residence","byear","dyear","socstate","bmonth","dmonth","bday","dday",
                "number_apps","race_first","race_last","race_last_cyear","race_first_cyear",
                "race_last_cmonth","race_first_cmonth","race_change","number_claims","bpl",
                "mother_fname","mother_mname",
                "mother_lname","age_first_application","death_age","weight","ccweight",
                "bpl_string","socstate_string")
middle_f <- "father_mname"
file_name_f <- "bunmd_f.csv"
clean_names_father(in_path_f, file_name_f, lname_col_f, fname_col_f,
            sex_col_f, other_cols_f, out_path_f, middle_f)
```

#Check cleaned dataset 
```{r}
bunmd_cleaned_f <- fread("/censoc/data/numident/4_berkeley_unified_mortality_database/bunmd_f_cleaned.csv")
```

#repeat for mothers- create clean_names_mother function
```{r}
#clean fname ABE Method https://github.com/uguryi/abeR/blob/master/R/clean-names.R

clean_names_mother <- function(in_path, file_name, lname_col, fname_col, sex_col, other_cols, out_path, middle) {

  # Read data
  path_to_data <- paste(in_path, file_name, sep="/")
  data <- readr::read_csv(path_to_data)

  # Copy raw names before processing
  data$lname_clean <- data[[lname_col]]
  data$fname_clean <- data[[fname_col]]

  # Concatenate middle name to first name if middle name exists
  if (middle != "") {
    concat_middle_to_first <- function(row) {
      if (is.na(row["fname_clean"])) {first <- NA}
      else if (is.na(row[[middle]])) {first <- row["fname_clean"]}
      else {first <- paste(row["fname_clean"], row[[middle]])}
      return(first)
    }
    data$fname_clean <- apply(data, 1, concat_middle_to_first) # SLOW
    data$mname_raw <- data[[middle]]
    data[[middle]] <- NULL
  }
  # NOTES
  # 1. This statement is here to make sure that there aren't any inconsistencies
  #    in relation to middle initials. For example, in the case of CenSoc, BUNMD already
  #    has middle initials but the way the ABE algorithm processes first names and
  #    initials are sometimes different from the way BUNMD stores first names and initials.
  #    ABE turns "A J" into "A J" as first name and "J" as initial, while in BUNMD "A J"
  #    would be first name "A" and initial "J".

  # Convert unicode to ascii to remove accents
  data$lname_clean <- iconv(data$lname_clean, to="ASCII//TRANSLIT")
  data$fname_clean <- iconv(data$fname_clean, to="ASCII//TRANSLIT")

  # Convert to lower case
  data$lname_clean <- tolower(data$lname_clean)
  data$fname_clean <- tolower(data$fname_clean)

  # Trim whitespace
  data$lname_clean <- stringr::str_trim(data$lname_clean)
  data$fname_clean <- stringr::str_trim(data$fname_clean)

  # Replace . with " "
  data$lname_clean <- gsub("\\.", " ", data$lname_clean)
  data$fname_clean <- gsub("\\.", " ", data$fname_clean)

  # Keep only letters and spaces (accents don't work)
  data$lname_clean <- gsub("[^a-z ]", "", data$lname_clean)
  data$fname_clean <- gsub("[^a-z ]", "", data$fname_clean)

  # Get rid of titles
  #data$fname_clean <- paste0(" ", data$fname_clean)
  #data$fname_clean <- remove_titles(data$fname_clean)

  # Fix two-word names (e.g., mc donnell --> mcdonnell)
  data$lname_clean <- paste0(" ", data$lname_clean)
  data$lname_clean <- gsub(" st ",    " st",    data$lname_clean)
  data$lname_clean <- gsub(" ste ",   " ste",   data$lname_clean)
  data$lname_clean <- gsub(" saint ", " saint", data$lname_clean)
  data$lname_clean <- gsub(" virg ",  " virg",  data$lname_clean)
  data$lname_clean <- gsub(" mac ",   " mac",   data$lname_clean)
  data$lname_clean <- gsub(" mc ",    " mc",    data$lname_clean)
  data$lname_clean <- stringr::str_trim(data$lname_clean)
  data$fname_clean <- gsub(" st ",    " st",    data$fname_clean)
  data$fname_clean <- gsub(" ste ",   " ste",   data$fname_clean)
  data$fname_clean <- gsub(" saint ", " saint", data$fname_clean)
  data$fname_clean <- gsub(" virg ",  " virg",  data$fname_clean)
  data$fname_clean <- gsub(" mac ",   " mac",   data$fname_clean)
  data$fname_clean <- gsub(" mc ",    " mc",    data$fname_clean)
  data$fname_clean <- stringr::str_trim(data$fname_clean)
  # NOTES
  # 1. Many more words can be added here, such as:
  #    la, van, du, de, del, o, pal, le, de la

   # Split first name (i.e., deal with middle names and initials)
  res <- lapply(data$fname_clean, split_first_name) # SLOW (~60s with ~3M obs)
  data$first  <- unlist(res)[attr(unlist(res),"names") == "first"]
  data$mname <- unlist(res)[attr(unlist(res),"names") == "middle"]
  rm(res)
  
  # Merge nicknames
  data_final <- merge(data, nicknames, by.x = c("first", sex_col),
                      by.y = c("nickname", "sex"), all.x = TRUE) # SLOW (~30s with ~3M obs)
  data_final$fname_final <- apply(data_final, 1, merge_names_no_middle) # SLOW (~45s with ~3M obs)

  # Keep only the first multicharacter part of lname_clean
  res <- lapply(data_final$lname_clean, split_first_name) # SLOW (~60s with ~3M obs)
  data_final$lname_final <- unlist(res)[attr(unlist(res),"names") == "first"]
  rm(res)
   # Reorder columns, drop unnecessary ones (fname_clean, lname_clean, first, name)
  if (middle == "") {
    cols_to_keep <- c("fname_final", "mname", "lname_final", sex_col, fname_col, lname_col, other_cols)
    data_final <- data_final[,cols_to_keep]
    colnames(data_final) <- c("fname", "mname", "lname", "sex", "fname_raw", "lname_raw", other_cols)
  } else {
    cols_to_keep <- c("fname_final", "mname", "lname_final", sex_col, fname_col, "mname_raw", lname_col, other_cols)
    data_final <- data_final[,cols_to_keep]
   colnames(data_final) <- c("mother_fname", "mother_mname", "mother_lname", "mother_sex", "mother_fname_raw", "mother_mname_raw", "mother_lname_raw", other_cols)
  }
 # Do a final trim
  data_final$lname <- stringr::str_trim(data_final$lname)
  data_final$fname <- stringr::str_trim(data_final$fname)
  data_final$mname <- stringr::str_trim(data_final$mname)

  # Save dataset
  dir.create(out_path, showWarnings = FALSE)
  path_to_processed_data <- paste(out_path, gsub(".csv", "_cleaned.csv", file_name), sep="/")
  readr::write_csv(data_final, path_to_processed_data)

  # Clean up after yourself
  rm(list = ls())
}
```

#apply function to mothers
```{r}
#add sex column for mother 
bunmd_cleaned_f$mother_sex <- 2
#write out new dataset
fwrite(bunmd_cleaned_f, "/censoc/data/numident/4_berkeley_unified_mortality_database/bunmd_sampled_f.csv")


in_path_f <-"/censoc/data/numident/4_berkeley_unified_mortality_database"
lname_col_f <- "mother_lname"
fname_col_f <- "mother_fname"
sex_col_f <- "mother_sex" 
out_path_f      <- "/censoc/data/numident/4_berkeley_unified_mortality_database"
other_cols_f <- c("fname","mname","lname","sex","fname_raw","mname_raw","lname_raw","father_fname",
                  "father_mname","father_lname","father_sex","father_fname_raw",
                  "father_mname_raw","father_lname_raw",
                  "ssn","zip_residence",
                  "byear","dyear","socstate","bmonth","dmonth","bday","dday",
                "number_apps","race_first","race_last","race_last_cyear","race_first_cyear",
                "race_last_cmonth","race_first_cmonth","race_change","number_claims","bpl",
                "age_first_application","death_age","weight","ccweight",
                "bpl_string","socstate_string")
middle_f <- "mother_mname"
file_name_f <- "bunmd_sampled_f.csv"
clean_names_mother(in_path_f, file_name_f, lname_col_f, fname_col_f,
            sex_col_f, other_cols_f, out_path_f, middle_f)
```

#Check cleaned dataset 
```{r}
bunmd_cleaned_m <- fread("/censoc/data/numident/4_berkeley_unified_mortality_database/bunmd_sampled_f_cleaned.csv")
```

#reorder columns in clean dataset and export final file
```{r}
bunmd_names_cleaned <- bunmd_cleaned_m[,c("fname","mname","lname","fname_raw","mname_raw","lname_raw","sex","ssn","zip_residence","byear","dyear","socstate","bmonth","dmonth","bday","dday","number_apps","race_first","race_last","race_last_cyear","race_first_cyear","race_last_cmonth","race_first_cmonth","race_change" , "number_claims","bpl", "age_first_application","death_age","weight","ccweight","bpl_string","socstate_string",
                   "father_fname","father_mname","father_lname","father_fname_raw","father_mname_raw","father_lname_raw","father_sex","mother_fname","mother_mname","mother_lname","mother_fname_raw","mother_mname_raw","mother_lname_raw","mother_sex")]
fwrite(bunmd_names_cleaned, "/censoc/data/numident/4_berkeley_unified_mortality_database/bunmd_names_cleaned.csv")

```

