---
title: "2-process-name-enlistment"
output: html_document
---
#Adapted from code written by Ugur Yildirim

# Library packages, source functions, set paths
```{r}
library(dplyr)     # %>%, group_by, distinct, tally

library(readr)     # read_csv, write_csv
library(stringr)   # str_trim, str_split
library(tidyverse) # add_row
library(data.table)
packageurl <- "https://cran.r-project.org/src/contrib/Archive/gender/gender_0.5.4.tar.gz"
install.packages(packageurl, repos=NULL, type="source") #gender package not available on this 
#r software so must use different method to install gender & gender data packages
library(gender)
library(genderdata)

```
#Set paths to temp data, outpath, and tan path to titles

```{r}
path_to_temp_data <- "/data/josh/CenSoc/censoc_data/enlistment-records-public/code/awenlistmentcode/temp1.csv"
outpath <- "/data/josh/CenSoc/censoc_data/enlistment-records-public/code/awenlistmentcode"
tan_path   <- "/data/josh/CenSoc/censoc_data/enlistment-records-public/data/codes/titles-and-nicknames"
```


# Read in temp1
```{r}
data_combined <- read_csv(path_to_temp_data, 
                          col_types = cols(
                            empty_field = col_character(),
                            term_or_enlistment = col_character(),
                            longevity = col_character(),
                            education = col_character(),
                            marital_status = col_character(),
                            component = col_character(),
                            educ_spec = col_character(),
                            defer_date_mmyy = col_character()
                            )
                          )
```

# Process name

# Keep original name as is for future reference
```{r}
data_combined$name_clean <- data_combined$name

# Do some preliminary cleaning in line with abeclean
data_combined$name_clean <- iconv(data_combined$name_clean, to="ASCII//TRANSLIT")
data_combined$name_clean <- tolower(data_combined$name_clean)
data_combined$name_clean <- str_trim(data_combined$name_clean)
data_combined$name_clean <- gsub("\\.", " ", data_combined$name_clean)
data_combined$name_clean <- gsub("[^a-z ]", "", data_combined$name_clean)

# Prefixes in abeclean
data_combined$name_clean <- gsub("^st ",    "st",    data_combined$name_clean)
data_combined$name_clean <- gsub("^ste ",   "ste",   data_combined$name_clean)
data_combined$name_clean <- gsub("^saint ", "saint", data_combined$name_clean)
data_combined$name_clean <- gsub("^virg ",  "virg",  data_combined$name_clean)
data_combined$name_clean <- gsub("^mac ",   "mac",   data_combined$name_clean)
data_combined$name_clean <- gsub("^mc ",    "mc",    data_combined$name_clean)

# Additional prefixes
data_combined$name_clean <- gsub("^la ",    "la",   data_combined$name_clean)
data_combined$name_clean <- gsub("^van ",   "van",  data_combined$name_clean)
data_combined$name_clean <- gsub("^du ",    "du",   data_combined$name_clean)
data_combined$name_clean <- gsub("^de ",    "de",   data_combined$name_clean)
data_combined$name_clean <- gsub("^del ",   "del",  data_combined$name_clean)
data_combined$name_clean <- gsub("^o ",     "o",    data_combined$name_clean)
data_combined$name_clean <- gsub("^le ",    "le",   data_combined$name_clean)
data_combined$name_clean <- gsub("^dela ",  "dela", data_combined$name_clean)
data_combined$name_clean <- gsub("^de la ", "dela", data_combined$name_clean)
data_combined$name_clean <- gsub("^di ",    "di",   data_combined$name_clean)
data_combined$name_clean <- gsub("^von ",   "von",  data_combined$name_clean)
data_combined$name_clean <- gsub("^pal ",   "pal",  data_combined$name_clean)
```


# Split name into fname and lname
``` {r}
split_name <- function(row) {
  name_split <- str_split(row, " ", 2)[[1]]
  lname <- name_split[1]
  fname <- name_split[2]
  return(list(lname=lname, fname=fname))
}

res <- lapply(data_combined$name_clean, split_name) # SLOW (~5 minutes with ~9 million observations)
data_combined$lname_clean <- unlist(res)[attr(unlist(res), "names") == "lname"]
data_combined$fname_clean <- unlist(res)[attr(unlist(res), "names") == "fname"]
rm(res)

data_combined$lname_clean <- str_trim(data_combined$lname_clean)
data_combined$fname_clean <- str_trim(data_combined$fname_clean)
```

#Define functions for name cleaning
``` {r}
# Define remove_titles
#' @export
remove_titles <- function(row) {
  fname_clean <- row
  for (title in titles$title) {
    pattern <- paste0(" ", title, " ")
    fname_clean <- stringi::stri_replace_all_regex(fname_clean, pattern, "")
  }
  return(fname_clean)
}

# Define split_first_name
#' @export
split_first_name <- function(row) {
  fname_split <- strsplit(row, " ")[[1]]
  fname_split <- fname_split[fname_split != ""]
  if (is.na(fname_split[1])) {
    first <- NA
    middle <- NA
  } else if (is.na(fname_split[2])) {
    first <- fname_split[1]
    middle <- NA
  } else if (stringr::str_length(fname_split[1]) > 1) {
    if (length(unique(fname_split)) != length(fname_split)) {
      fname_split <- rle(fname_split)$values
    }
    first <- fname_split[1]
    middle <- fname_split[2]
  } else if (stringr::str_length(fname_split[2]) > 1) {
    if (length(unique(fname_split)) != length(fname_split)) {
      fname_split <- rle(fname_split)$values
    }
    first <- fname_split[2]
    middle <- fname_split[3]
  } else {
    first <- row
    middle <- fname_split[2]
  }
  return(list(first=first, middle=middle))
}

```

#ABECLEAN Method for name cleaning
```{r}
# Titles in abeclean
titles <- read_csv(paste(tan_path, "titles.csv", sep="/"))

# Additional titles
titles <- titles %>% add_row(title = "trd")
titles <- titles %>% add_row(title = "ii")
titles <- titles %>% add_row(title = "iii")

# Remove titles from fname
#data_combined$fname_clean <- paste0(" ", data_combined$fname_clean) # this is how abeclean does it
data_combined$fname_clean <- paste0(" ", data_combined$fname_clean, " ")
data_combined$fname_clean <- remove_titles(data_combined$fname_clean) # SLOW (~11 minutes with ~9 million observations)
data_combined$fname_clean <- str_trim(data_combined$fname_clean)

# Split fname_clean into fname_clean and mname_clean
res <- lapply(data_combined$fname_clean, split_first_name) # SLOW (~6 minutes with ~9 million obs)
data_combined$fname_clean <- unlist(res)[attr(unlist(res),"names") == "first"]
data_combined$mname_clean <- unlist(res)[attr(unlist(res),"names") == "middle"]
rm(res)
```
#Sex determination
```{r}
# Determine sex from fname_clean
sex_temp <- gender(data_combined$fname_clean)
sex_temp <- sex_temp %>% group_by(name) %>% distinct()

sex_temp %>% group_by(gender) %>% tally() # 1 either
sex_temp$gender <- ifelse(sex_temp$gender == "either", NA, sex_temp$gender) 
sex_temp %>% group_by(gender) %>% tally() # 1 NA

sex_temp$sex <- ifelse(is.na(sex_temp$gender), NA, ifelse(sex_temp$gender == "male", 1, 2))
sex_temp %>% group_by(sex) %>% tally()
  
sex_temp <- sex_temp[,c("name", "sex")]
colnames(sex_temp) <- c("fname_clean", "sex")

data_combined <- merge(data_combined, sex_temp, by = "fname_clean", all.x = T)

nrow(subset(data_combined, sex == 1)) / 
  (nrow(subset(data_combined, sex == 1)) + nrow(subset(data_combined, sex == 2))) # 0.9625251

data_combined$sex_NA <- ifelse(is.na(data_combined$sex), 1, 0)

sum(data_combined$sex_NA) # 447140

# Use branch_alpha and branch_code to get a more precise sex value
# branch_alpha == "WAC" --> Women's Army Corps
# branch_code  == "18"  --> Women's Army Corps

data_combined$branch_alpha <- str_trim(data_combined$branch_alpha)
data_combined$branch_code  <- str_trim(data_combined$branch_code)

nrow(subset(data_combined, branch_alpha == "WAC" | branch_code == "18")) # 133056
nrow(subset(data_combined, sex == 2))                                    # 322011

get_gender_from_branch <- function(row) {
  if (is.na(row["branch_alpha"]) & is.na(row["branch_code"])) {return(row["sex"])}
  else if (is.na(row["branch_alpha"])) {if (row["branch_code"]  == "18")  {return(2)} else {return(1)}}
  else if (is.na(row["branch_code"]))  {if (row["branch_alpha"] == "WAC") {return(2)} else {return(1)}}
  else {if (row["branch_alpha"] == "WAC" | row["branch_code"] == "18") {return(2)} else {return(1)}}
}

data_combined$sex_r <- apply(data_combined, 1, get_gender_from_branch) # SLOW

nrow(subset(data_combined, sex_r == 2)) # 133056

nrow(subset(data_combined, sex_r == 1)) / 
  (nrow(subset(data_combined, sex_r == 1)) + nrow(subset(data_combined, sex_r == 2))) # 0.9852812

data_combined$sex_r_NA <- ifelse(is.na(data_combined$sex_r), 1, 0) # 0 NA's!
```

# Standardize names
```{r}
male_nicknames   <- read_csv(paste(tan_path, "male_nicknames.csv",   sep="/"))
female_nicknames <- read_csv(paste(tan_path, "female_nicknames.csv", sep="/"))
nicknames        <- as.data.frame(rbind(male_nicknames, female_nicknames))

colnames(nicknames) <- c("sex", "fname_std", "nickname")

data_combined <- merge(data_combined, nicknames, by.x = c("fname_clean", "sex_r"),
                       by.y = c("nickname", "sex"), all.x = TRUE)

merge_names <- function(row) {
  if (is.na(row["fname_std"])) {first <- row["fname_clean"]}
  else {first <- row["fname_std"]}
  return(first)
}

data_combined$fname_clean_std <- apply(data_combined, 1, merge_names) # SLOW

```

# Save processed dataset
``` {r}
write_csv(data_combined, paste0(outpath, "/temp2.csv"))
rm(list = ls())
```