---
title: "2 censoc all gender sample"
output: html_notebook
---

This notebook will read in the Berkeley Unified Mortality Database (BUNMD) and the 1940 Census (this can be switched). The code constructs linking keys, dedupes the keys, and then performs a direct match on the linking keys.  

First, we will read in the numdeath matched with numapp file. 

```{r}
library(censocdev)
```

```{r}
bunmd <- fread("/censoc/data/numident/4_berkeley_unified_mortality_database/bunmd.csv", colClasses = list(character= 'ssn')) 
```

Read in the census file. 
```{r, eval = F}
## Load census with HISTID
census.1940 <- load_census_numident_match(census_file = "/home/ipums/casey-ipums/IPUMS2019/1940/TSV/P.tsv")
```

Merge Numident and Census Files on linking keys: 1940
```{r, eval = F}
setDT(bunmd)
setDT(census.1940)

## censoc numident
censoc.numident <- census_bunmd_merge(bunmd = bunmd, census = census.1940, census_year = 1940)

## censoc numident
censoc.numident <- clean_wcensoc(censoc.numident, census_year = 1940) 

## Restrict to matches only in high coverage years
censoc.numident.select <- censoc.numident %>% 
  filter(dyear %in% c(1988:2005)) %>% 
  select(-weight, -ccweight)

## calculate age at death
censoc.numident.select <- calculate_age_at_death(censoc.numident.select)

## calculate weights 
censoc.numident.select <- create_weights_censoc_numident(censoc.numident.select)

## write out full version matched with censoc 
fwrite(censoc.numident.select, "/censoc/data/censoc_linked_with_census/1940/censoc_numident_v1.csv")
```

Select the variables for the public release — this will likely change 
```{r}
censoc.numident.select <- censoc.numident.select %>% 
  select(histid = HISTID, byear, bmonth, dyear, dmonth, death_age, sex, race_first, race_first_cyear, race_first_cmonth,race_last, race_last_cyear, race_last_cmonth, bpl, zip_residence, socstate, age_first_application, weight)

fwrite(censoc.numident.select, "/censoc/data/censoc_files_for_website/censoc_numident_v1.csv")
```





