---
title: "Make CenSoc-Numident"
author: Casey Breen
---

Summary: This notebook will link records in the Berkeley Unified Mortality Database (BUNMD) to the 1940 Census. The BUNMD is a cleaned and harmonized version of the NARA Numident Records. The code constructs linking keys, dedupes the keys, and then performs a direct match on the linking keys.  

If the `censocdev` package hasn't already been installed, install from GitHub: 

```{r}
# library(devtools) 
# install_github("caseybreen/censocdev")
```

Library packages
```{r}
library(censocdev)
library(data.table)
library(tidyverse)
```

Read in the Berkeley Unified Numident Mortality Database. This is a harmonized and cleaned version on the NARA Numident files. 
```{r}
bunmd <- fread("/censoc/data/numident/4_berkeley_unified_mortality_database/bunmd.csv", colClasses = list(character= 'ssn')) %>% 
  select(-weight, -ccweight)
```

Read in the 1940 Census file. 
```{r, eval = F}
## Load census with HISTID
census.1940 <- load_census_numident_match(census_file = "/home/ipums/casey-ipums/IPUMS2019/1940/TSV/P.tsv")
```

Merge Numident and Census Files on linking keys.

TODO put in more info about merge 

- MARST codes: https://usa.ipums.org/usa-action/variables/MARST#codes_section

```{r, eval = F}
## Set BUNMD and census as data.
setDT(bunmd)
setDT(census.1940)

## censoc bunmd merge
censoc.numident <- create_censoc_numident(bunmd = bunmd, census = census.1940, census_year = 1940)

## censoc numident
censoc.numident <- clean_wcensoc(censoc.numident, census_year = 1940) 

## Restrict to matches only in high coverage years
censoc.numident <- censoc.numident %>% 
  filter(dyear %in% c(1988:2005)) 

## calculate age at death
censoc.numident <- calculate_age_at_death(censoc.numident)

## calculate weights 
censoc.numident <- create_weights_censoc_numident(censoc.numident, cohorts = c(1895:1939), death_ages = c(65:100))
```

Select the variables for the public release. We don't include bday and dday variables per our agreement with IPUMS.
```{r}
## select numident 
censoc.numident.website <- censoc.numident %>% 
  select(HISTID, byear, bmonth, dyear, dmonth, death_age, sex, race_first, race_first_cyear,
         race_first_cmonth,race_last, race_last_cyear, race_last_cmonth, bpl, zip_residence,
         socstate, age_first_application, weight)

## write out public version  
fwrite(censoc.numident.website, "/censoc/data/censoc_files_for_website/censoc_numident_v1.csv")
```

Add person and household variables
```{r}
## Drop census
census <- fread("/home/ipums/casey-ipums/IPUMS2019/1940/TSV/P.tsv")

## Add person variables
## drop vars so we won't have doubles
censoc.numident.all.vars <- censoc.numident %>% 
  select(-AGE, -SEX, -NAMELAST, -NAMEFRST, -BPL, -RACE, -MARST) %>% 
  inner_join(census)

## add household variables
household_census <- fread("/home/ipums/casey-ipums/IPUMS2019/1940/TSV/H.tsv")
setDT(censoc.numident.all.vars)
censoc.numident.all.vars <- add_household_variables(censoc = censoc.numident.all.vars, household_census = household_census)

## write out file
fwrite(censoc.numident.all.vars, "/censoc/data/censoc_linked_with_census/1940/censoc_numident_all_vars_v1.csv")
```


