---
title: "Make CenSoc DMF"
author: Casey Breen (caseybreen@berkeley.edu)
---

Summary: This vignette goes through the steps of create the CenSoc-DMF file from input source data (Social Security Death Master File and 1940 Census). We match men on their first name, last name, and age at census. 

A copy of the `censocdev` package must be cloned from GitHub: 

```{r}
# library(devtools) 
# install_github("caseybreen/censocdev")
```

Library Packages
```{r}
library(censocdev)
library(tidyverse)
library(data.table)
```

Read in Social Security Death Master file, 1940 census file, and create censoc file. 

Note: n_clean_keys variable in the "census" file refers to the total number of unique keys for both men *and* women. 
```{r}
## Read in DMF (SSDM) files
dmf_files = c("/home/ipums/josh-ipums/progs/ssdm/ssdm1",
                     "/home/ipums/josh-ipums/progs/ssdm/ssdm2",
                     "/home/ipums/josh-ipums/progs/ssdm/ssdm3")

dmf <- load_dmf_deaths(dmf_files)

## Read in 1940 Census files
census <- load_census_dmf_match(census_file = "/home/ipums/casey-ipums/IPUMS2019/1940/TSV/P.tsv")

## Create CenSoc File
censoc.dmf <- create_censoc_dmf(census, dmf)
```

- Calculate age at death variables 
- Calculate weights
- Remove matches where first name was only one character 
```{r}
## Calculate age at death
censoc.dmf <- calculate_age_at_death(censoc.dmf)

## Calculate weights
censoc.dmf <- create_weights_censoc_dmf(censoc.dmf, cohorts = c(1895:1939), death_ages = c(65:100))

## Calculate DMF
censoc.dmf <- censoc.dmf %>% 
  filter(!nchar(fname) <= 1)
```

Select variables for public release
```{r}
censoc.dmf.web <- censoc.dmf %>% 
  select(HISTID, byear, bmonth, dyear, dmonth, death_age, weight)
```

Write out CenSoc with for website. This is the file we will publicly release. As per our agreement with IPUMS, we cannot release day of death or day of birth. 
```{r}
# fwrite(censoc.dmf.web, "/censoc/data/censoc_files_for_website/censoc_dmf_v1.csv")
```

Optional: add all person variables and household variables from 1940 census
```{r}
census <- fread("/home/ipums/casey-ipums/IPUMS2019/1940/TSV/P.tsv")

censoc.dmf.all.vars <- censoc.dmf %>% 
  select(-AGE, -SEX, -NAMELAST, -NAMEFRST) %>% 
  inner_join(census)

household_census <- fread("/home/ipums/casey-ipums/IPUMS2019/1940/TSV/H.tsv")
setDT(censoc.dmf.all.vars)
censoc.dmf.all.vars <- add_household_variables(censoc = censoc.dmf.all.vars, household_census = household_census)
```

Write out CenSoc file with all variables
```{r}
# fwrite(censoc.dmf.all.vars, "/censoc/data/censoc_linked_with_census/1940/censoc_dmf_all_vars_v1.csv")
```
