---
title: "BUNMD Supplementary Geography Variables File"
date: "2023-02-14"
output:
  pdf_document: default
  html_document: default
---

To create this file, we will need to read in a truncated version of the BUNMD and ZIP code of death variables files along with place of birth variables that were created using Seth Sanders' Crosswalk.

To construct the “place of birth” variables, we used a crosswalk originally developed for this paper: 

Black, Dan A., Seth G. Sanders, Evan J. Taylor, and Lowell J. Taylor. 2015. “The Impact of the Great Migration on Mortality of African Americans: Evidence from the Deep South.” American Economic Review 105(2):477–503. doi: 10.1257/aer.20120642.

```{r}
library(tidyverse)
library(data.table)
library(dplyr)
```

file paths
```{r}
# zip features data (ZIP to state, county, region, etc)
zip_features_path <- "/data/censoc/workspace/geography_variables/Zip_Features.csv"   

# string to GNIS feature by ssn, using Black et al.
bunmd_birthplace_crosswalk_path <- "/data/censoc/workspace/geography_variables/bunmd_Birthplace.csv"  

# raw unclean birthplace strings
unclean_birthplace_stings_path <- "/data/censoc/workspace/geography_variables/bunmd_city_string.csv" 

# full BUNMD V2.0
bunmd_path <- "/data/censoc/censoc_data_releases/bunmd/bunmd_v2/bunmd_v2.csv" 

# path to published Numident 2.1
numident_path <- "/data/censoc/censoc_data_releases/censoc_numident/censoc_numident_v2.1/censoc_numident_v2.1.csv"

# bunmd to numident ID variable crosswalk
ssn_crosswalk_path <- "/data/censoc/crosswalks/numident_ssn_histid_v2_1_crosswalk.csv"

# state FIPS codes
state_fips_path <- "/data/censoc/workspace/geography_variables/state_fips_codes.csv"

# path to write files
geo_files_out_path <- "/data/censoc/workspace/geography_variables/"
```



Here, I read in the data and select only for the variables of interest.
```{r}
Zip_Features <- read.csv(zip_features_path,
                         colClasses = c(zip = "character",FIPS = "character")) 

Zip_Features <- subset(Zip_Features, select=-c(RUC1974, area_name, conterminous, acceptable_cities,district_name,RUC1983,timezone,area_codes,OASDI_benefits_2005,OASDI_total_2005,type,elevation,state_full,delv_address,OASDI_average_2005,OASDI_disabled_2005,OASDI_benefits_2005,OASDI_retired_2005,geocode))

birthplaces_df <- fread(bunmd_birthplace_crosswalk_path,
                        select=c("ssn","bpl_city","feature_id","county_name", "Region", "county_numeric", "bpl_string"))

bunmd_trunc <- fread(bunmd_path,
                     select=c("ssn", "zip_residence","bpl_string"))
```


Next, I rename the ZIP code death place variables so that it'll be clear that these refer to places of death in the final file.
```{r}
Zip_Features <- Zip_Features %>% 
  rename(
    death_fips = names(Zip_Features)[1],
    death_zip = names(Zip_Features)[2],
    death_state = names(Zip_Features)[3],
    death_city = names(Zip_Features)[4],
    death_county = names(Zip_Features)[5],
    death_country = names(Zip_Features)[6],
    death_region = names(Zip_Features)[7],
    death_ruc1993 = names(Zip_Features)[8]
  )
head(Zip_Features)
```


I truncate the BUNMD zip_residence variable so that it only has 5 digits. This is an important step since the places of death file only contains 5 digit ZIP codes, and gives us a variable that we can match on. 
```{r}
bunmd_trunc <- bunmd_trunc %>% 
  mutate(
    zip = str_trunc(as.character(zip_residence), width = 5, side = "right", ellipsis = "")
  ) %>% 
  select(-2) %>% 
  rename(death_zip = names(.)[3])

head(bunmd_trunc)
```


```{r}
west_states <- c("Arizona", "Colorado", "Idaho", "Montana", "Nevada", "New Mexico", "Utah", "Wyoming", "Alaska", "California", "Hawaii", "Oregon", "Washington")
midwest_states <- c("Illinois", "Indiana", "Michigan", "Ohio", "Wisconsin", "Iowa", "Kansas", "Minnesota", "Missouri", "Nebraska", "North Dakota", "South Dakota")
northeast_states <- c("Connecticut", "Maine", "Massachusetts", "New Hampshire", "Rhode Island", "Vermont", "New Jersey", "New York", "Pennsylvania")
south_states <- c("Delaware", "Florida", "Georgia", "Maryland", "North Carolina", "South Carolina", "Virginia", "West Virginia", "Alabama", "Kentucky", "Mississippi", "Tennessee", "Arkansas", "Louisiana", "Oklahoma", "Texas")

bunmd_trunc$birth_region <- ifelse(bunmd_trunc$bpl_string %in% west_states, "west",
                              ifelse(bunmd_trunc$bpl_string %in% midwest_states, "midwest",
                                     ifelse(bunmd_trunc$bpl_string %in% northeast_states, "northeast",
                                            ifelse(bunmd_trunc$bpl_string %in% south_states, "south", NA))))
```


Once we have the matching ZIP code variables, we merge the information from the ZIP code file to the BUNMD file. This results in a new file that contains a person's SSN, ZIP code of death, and additional information sourced from the ZIP code file, such as the corresponding city, county, and state.
```{r}
bunmd_trunc <- merge(bunmd_trunc, Zip_Features, by = "death_zip", all.x = TRUE,sort = FALSE)
rm(Zip_Features)
head(bunmd_trunc)
```

```{r}
bunmd_trunc$death_state <- ifelse(bunmd_trunc$death_county == "DISTRICT OF COLUMBIA", "DC", bunmd_trunc$death_state)
```


Next, we rename birth place variables.
```{r}
birthplaces_df <- birthplaces_df %>% 
  rename(
    birth_city = names(birthplaces_df)[2],
    birth_gnis_code = names(birthplaces_df)[3],
    birth_county = names(birthplaces_df)[4],
    birth_region = names(birthplaces_df)[5],
    birth_fips = names(birthplaces_df)[6])

head(birthplaces_df)
```

merge county and state numeric codes to get birth fips
```{r}
state_fips <- fread(state_fips_path, select = c("stname", "st"))
names(state_fips) <- c("bpl_string", "state_fips_code")
state_fips <- state_fips %>%
  mutate(state_fips_code = str_pad(as.character(state_fips_code), width = 2, side = "left", pad = "0"))

birthplaces_df <- birthplaces_df %>%
  mutate(birth_fips = str_pad(as.character(birth_fips), width = 3, side = "left", pad = "0"))

birthplaces_df <- birthplaces_df %>% left_join(state_fips, by = "bpl_string")

birthplaces_df <- birthplaces_df %>% mutate(birth_fips = paste0(state_fips_code, birth_fips))

# drop if not a proper 5 digit FIPS code
birthplaces_df <- birthplaces_df %>% mutate(birth_fips = ifelse(nchar(birth_fips) < 5, "", birth_fips))
birthplaces_df$state_fips_code <-  NULL
```


Next, we merge birth place variables to the BUNMD file. These variables include the city, county, and state where the person was born. Now, we have a new file that contains information on people's death and birthplaces, in addition to their SSN.
```{r}
birthplaces_df <- merge(bunmd_trunc, birthplaces_df, by = "ssn", all.x = TRUE,sort = FALSE)
rm(bunmd_trunc)
head(birthplaces_df)
```

Let's see how well of a job the crosswalk did and count how many birth states in the BUNMD match with the birth states in the birthplaces file.
```{r}
birthplaces_df$birth_place_matched <- ifelse(birthplaces_df$bpl_string.x == birthplaces_df$bpl_string.y & !is.na(birthplaces_df$bpl_string.x),0,
                                             ifelse(birthplaces_df$bpl_string.x != birthplaces_df$bpl_string.y & !is.na(birthplaces_df$bpl_string.x),1,NA))

print(paste("There are",sum(birthplaces_df$birth_place_matched,na.rm=T),"mismatched birth states in the birth places data frame"))
```


```{r}
birthplaces_df <- birthplaces_df %>%
  select(-birth_region.y, -bpl_string.y, -bpl_string.x, -birth_place_matched) %>% 
  rename(birth_region = names(.)[3])
```

In the chunks of code below, we standardize the data set and make everything lowercase. We also make sure that any differences in wording are standardized. 
```{r}
birthplaces_df$birth_city <- tolower(birthplaces_df$birth_city)
birthplaces_df$birth_county <- tolower(birthplaces_df$birth_county)
birthplaces_df$death_state <- tolower(birthplaces_df$death_state)
birthplaces_df$death_city <- tolower(birthplaces_df$death_city)
birthplaces_df$death_county <- tolower(birthplaces_df$death_county)

head(birthplaces_df)
```

```{r}
birthplaces_df <- birthplaces_df %>% 
  mutate(
    death_region = ifelse(death_region == "North_East", "northeast", death_region)
  )

birthplaces_df$death_region <- tolower(birthplaces_df$death_region)

birthplaces_df <- birthplaces_df %>%
  mutate(death_region = str_trim(death_region)) %>%
  mutate(death_region = ifelse(death_region == "midewest", "midwest", death_region))

table(birthplaces_df$death_region)
```

```{r}
birthplaces_df <- birthplaces_df %>% 
  mutate(
    death_county = str_remove_all(death_county, "county"),
    death_county = str_trim(death_county)
  )
head(birthplaces_df$death_county, 10) #we remove the word "county" 
```


```{r}
birthplaces_df$death_state <- gsub("\\b\\s+", "", birthplaces_df$death_state)

birthplaces_df <- birthplaces_df %>% 
  mutate(
    death_state = case_when(
      death_state == "" ~ NA_character_,
      death_state == "arizona14/       " ~ "arizona",
      death_state == "newhampshire" ~ "new hampshire",
      death_state == "southcarolina" ~ "south carolina",
      death_state == "newjersey" ~ "new jersey",
      death_state == "newmexico" ~ "new mexico",
      death_state == "northdakota" ~ "north dakota",
      death_state == "northcarolina" ~ "north carolina",
      death_state == "southdakota" ~ "south dakota",
      death_state == "westvirginia" ~ "west virginia",
      death_state == "rhodeisland" ~ "rhode island",
      death_state == "newyork" ~ "new york",
      TRUE ~ death_state
    )
  )
unique(birthplaces_df$death_state)
```

We then make sure that the FIPS and ZIP code variables are coded as string variables so that the leading zeroes do not get truncated.  
```{r}
birthplaces_df$death_zip <- as.character(birthplaces_df$death_zip)
birthplaces_df$death_fips <- as.character(birthplaces_df$death_fips)

# change NA's to blank strings for these variables (death_zip already okay)
# will prevent them from being read as integers
birthplaces_df[is.na(death_fips), death_fips := ""]
birthplaces_df[is.na(birth_fips), birth_fips := ""]
```

We also include the "uncleaned" birthplace string for researchers who might want to work with them
```{r}
birth_city_uncleaned <- fread(unclean_birthplace_stings_path)

head(birth_city_uncleaned)
```


```{r}
birthplaces_df <- merge(birthplaces_df, birth_city_uncleaned, by = "ssn", all.x = TRUE,sort = FALSE)

birthplaces_df <- select(birthplaces_df, -c(V1,bpl_city_oflo))

birthplaces_df <- birthplaces_df %>% 
  rename(birth_city_uncleaned = names(birthplaces_df)[15])
    
birthplaces_df$birth_city_uncleaned <- tolower(birthplaces_df$birth_city_uncleaned)

birthplaces_df <- birthplaces_df %>%
  relocate(names(birthplaces_df)[15], .before = names(birthplaces_df)[11])

head(birthplaces_df)
rm(birth_city_uncleaned)
```

Then, we remove all people who have neither a birth or a death city. 
```{r}
birthplaces_df <- birthplaces_df %>% 
  mutate(
    deletes = paste(birth_region, death_city)
  ) %>% 
  filter(!deletes %in% "NA NA") %>% 
  select(-deletes)
head(birthplaces_df)
```
rearrange a bit
```{r}
birthplaces_df <- birthplaces_df %>%
  relocate(birth_region, .before = birth_city_uncleaned) %>% 
  relocate(death_fips, .after = death_county)
```



Next, we save the file as a csv. This file is usable with the BUNMD
```{r}
write.csv(birthplaces_df,
          paste0(geo_files_out_path, "BUNMD_Geography_File.csv"),
          row.names = FALSE)
```

Next, let's read in our HISTID crosswalk and merge it onto the Geography File so we can use it to merge onto the Numident.
```{r}
HISTID <- fread(ssn_crosswalk_path)
birthplaces_df <- fread(paste0(geo_files_out_path, "BUNMD_Geography_File.csv"))
```


```{r}
numident_birthplaces_df <- left_join(HISTID, birthplaces_df, by = "ssn")

numident_birthplaces_df <- numident_birthplaces_df %>%
  select(-ssn)

write.csv(numident_birthplaces_df,
          paste0(geo_files_out_path, "Numident_Geography_File.csv"),
          row.names = FALSE)
```


# create conservative numident version for publication
```{r}
censoc_numident <- fread(numident_path, select = c("HISTID", "link_abe_exact_conservative"))

# limit to conservative links
censoc_numident <- censoc_numident[link_abe_exact_conservative == 1]

numident_birthplaces_df_conservative <- numident_birthplaces_df[histid %in% censoc_numident$HISTID]

# capitalize histid variable
numident_birthplaces_df_conservative <- numident_birthplaces_df_conservative %>% 
  rename(HISTID = histid)

# write
write.csv(numident_birthplaces_df_conservative,
          paste0(geo_files_out_path, "numident_geography_file_v2.1_conservative.csv"),
          row.names = FALSE)
```


