---
title: "Numident Supplementary Geography File Codebook[^updated]"
output:
  pdf_document:
header-includes:
- \usepackage{hyperref}
- \usepackage{float}
- \restylefloat{table}
- \usepackage{caption}
---
[^updated]: Last updated: `r format(Sys.time(), '%d %B, %Y')`

\captionsetup[table]{labelformat=empty}

| Page                        | Variable                                     | Label                                        |
|-----------------------------|:---------------------------------------------|:---------------------------------------------|
| \hyperlink{page.3}{3}       | \hyperlink{page.3}{HISTID}                   |Historical Unique Identifier                  |
| \hyperlink{page.4}{4}       | \hyperlink{page.4}{birth\_gnis\_code}        |Place of birth, GNIS code                     |
| \hyperlink{page.5}{5}       | \hyperlink{page.5}{birth\_city}              |Place of birth, city                          | 
| \hyperlink{page.6}{6}       | \hyperlink{page.6}{birth\_county}            |Place of birth, county                        |
| \hyperlink{page.7}{7}       | \hyperlink{page.7}{birth\_fips}              |Place of birth, FIPS county code              |
| \hyperlink{page.8}{8}       | \hyperlink{page.8}{birth\_region}            |Place of birth, census region                 |
| \hyperlink{page.9}{9}       | \hyperlink{page.9}{death\_zip}               |Place of death, 5-digit ZIP code              |
| \hyperlink{page.10}{10}     | \hyperlink{page.10}{death\_city}              |Place of death, city                          |
| \hyperlink{page.11}{11}     | \hyperlink{page.11}{death\_county}           |Place of death, county                        |
| \hyperlink{page.12}{12}     | \hyperlink{page.12}{death\_fips}             |Place of death, FIPS county codes             |
| \hyperlink{page.13}{13}     | \hyperlink{page.13}{death\_state}            |Place of death, state                         |
| \hyperlink{page.14}{14}     | \hyperlink{page.14}{death\_region}           |Place of death, census region                 |
| \hyperlink{page.15}{15}     | \hyperlink{page.15}{death\_country}          |Place of death, country                       |
| \hyperlink{page.16}{16}     | \hyperlink{page.16nrow()}{death\_ruc1993}          |place of death, county rural-urban continuum  |

\vspace{10pt}

\begin{center}
\underline{\textbf{Summary and Methodology}}
\end{center}
\vspace{10pt}


The Numident Supplementary Geography File (N = 6,971,468) provides a set of supplementary geography variables reporting place of birth and death for individuals in the Numident. This file can be linked onto the Numident at the individual-level using HISTID. Individuals records may contain only place of birth variables, only place of death variables, or both place of birth and place of death variables.

**Place of death variables:** To construct the place of death variables, we use the ZIP code of last residence available in Social Security Numident records. It is likely that this reflects the ZIP code where an individual last lived. We map these ZIP codes onto city, county, state, census region, country, and rural-urban continuum codes using a database from the United States Postal Service (USPS) [\textcolor{blue}{(link)}](https://postalpro.usps.com/ZIP_Locale_Detail). For ZIP codes that have been decommissioned, we use a secondary database from UnitedStatesZipCodes.org. Approximately 6.3 million records have a ZIP code mapped to state, county, and city information, though some fields may be missing for certain records. Place of death variables are primarily available for individuals who died in the 50 U.S. states and the District of Columbia, but some are also present for deaths in Puerto Rico and a small number of ZIP codes indicating military bases abroad.


**Place of birth variables:** To construct the place of birth variables, we use 12-character city/county of birth string from the Socal Security Numident records. These strings are uncleaned and contain misspellings and other inconsistencies. We mapped these uncleaned strings onto Geographic Names Information System (GNIS) codes using the crosswalk developed for the paper:

|     Black, Dan A., Seth G. Sanders, Evan J. Taylor, and Lowell J. Taylor. 2015. “The Impact of the Great 
|     Migration on Mortality of African Americans: Evidence from  the Deep South.”*American Economic Review.*
|    105(2):477–503. doi: 10.1257/aer.20120642.

To construct other place of birth variable, we mapped the GNIS codes onto city, county, state, and census regions using a database from the U.S. Board on Geographic Names [\textcolor{blue}{(link)}](https://www.usgs.gov/u.s.-board-on-geographic-names/download-gnis-data). Please cite Black et al. (2015) if you are using any of the birthplace geography variables in the file. Birthplace variables are available for nearly 6.5 million records in this dataset. Place of birth variables are only available for locations within the 50 United States and the District of Columbia. Note that state/country of birth is already included in the CenSoc-Numident (see the variables **bpl** and **bpl_string**) and thus not present in this dataset.


\newpage 


```{r echo = F, results='hide', message=FALSE, warning=FALSE}
library(dplyr)
library(data.table)
library(tidyverse)
library(knitr)
library(forcats)
library(cowplot)
library(readxl)
library(kableExtra)

Numident_Geography_File <- fread("/data/censoc/workspace/geography_variables/numident_geography_file_v2.1_conservative.csv",
                              colClasses = c("death_fips" = "character", "death_zip" = "character",
                                             "birth_fips" = "character"))


numident <- fread("/data/censoc/censoc_data_releases/censoc_numident/censoc_numident_v2.1/censoc_numident_v2.1.csv")
numident <- numident[link_abe_exact_conservative == 1]

numident_merged <- merge(numident, Numident_Geography_File, by = "HISTID", all.x = TRUE)
rm(numident)

#replace empty strings with NA to make match rates easier to compute
numident_merged[numident_merged == ""] <- NA # SLOW

numident_merged$death_place_match <-ifelse(!is.na(numident_merged$death_county),1,0)

nrow(numident_merged)
```

\newpage 

**\huge HISTID**
\normalsize
\vspace{12pt}

**Label**: Historical Unique Identifier

**Description**: HISTID is a unique individual-level identifier from IPUMS-USA 1940 Census data. This variable uniquely identifies all records in the dataset and can be used to merge onto the CenSoc-Numident mortality dataset.




\newpage 

**\huge birth_gnis_code **

\normalsize
\vspace{12pt}

**Label:** Place of birth, GNIS code

**Description**:birth_gnis_code is a numeric variable that reports a person's Geographic Names Information System (GNIS) code for their place of birth. Each GNIS code maps onto physical locations located in the U.S., including longitude and latitude coordinates, physical feature names, counties, and states. For more information on GNIS codes, please see [\textcolor{blue}{https://www.usgs.gov/faqs/what-geographic-names-information-system-gnis}](https://www.usgs.gov/faqs/what-geographic-names-information-system-gnis).

Missing values occur if birth city is absent from Numident records, or if the original birth city string was not matched to a GNIS code.

\vspace{40 pt}

```{r echo = FALSE, message = FALSE, warning = FALSE, fig.align = "center"}
unclean_birthplace_stings_path <- "/data/censoc/workspace/geography_variables/bunmd_city_string.csv" 
birth_city_raw_strings <- fread(unclean_birthplace_stings_path, select = c("ssn", "bpl_city"))
ssn_crosswalk_path <- "/data/censoc/crosswalks/numident_ssn_histid_v2_1_crosswalk.csv"
ssn_to_histid <- fread(ssn_crosswalk_path)
histid_with_raw_city_strings <- left_join(ssn_to_histid, birth_city_raw_strings, by = "ssn") %>% 
  select(-c(ssn)) %>% rename(HISTID = histid)
numident_merged <- left_join(numident_merged, histid_with_raw_city_strings, by = "HISTID")

valid_birth_percent <- 100 - (mean(is.na(numident_merged$birth_gnis_code)) * 100)
birth_percent <- 100 - (mean(is.na(numident_merged$bpl_city)) * 100)
death_percent <- (mean(!is.na(numident_merged$zip_residence)) * 100)
valid_death_percent <- 100 - (mean(is.na(numident_merged$death_county)) * 100)

birth_props <- c(birth_percent,valid_birth_percent)
death_props <- c(death_percent,valid_death_percent)

birth_props_plot <- barplot(birth_props, names.arg = c("Have birth city information", "Birth city matched to GNIS code"),
        xlab = "Birth City Information Availibility", ylab = "Percentage of CenSoc-Numident Records",
        main = "Birth City Match Percentage",
        col = c("lightblue", "grey"),
        ylim = c(0, 100)) # Set the y-axis range to 0-100

# Calculate the bar centers
bar_centers <- birth_props_plot

#value_labels <- c("98%", "91%")
value_labels <- paste0(round(birth_props,1), "%")

text(x = bar_centers, y = birth_props, labels = value_labels, col = "black", cex = 1, pos = 1)
```





\newpage 

**\huge birth_city **
\normalsize
\vspace{12pt}

**Label**: Place of birth, city

**Description**: birth_city reports the 12-character city/county of birth string from the Numident Application records. This variable is generally uncleaned and unprocessed, and contains spelling errors and naming inconsistencies. 

Missing values occur when birth city is not available in Numident records.







\newpage 

**\huge birth_county **
\normalsize
\vspace{12pt}

**Label**: Place of birth, county

**Description**: birth_county is a character variable reporting a person's county of birth, as determined by GNIS place of birth. 

Missing values occur if birth city is absent from Numident records, or if the original birth city string was not matched to a GNIS code.

**Note**: In 2022, the state of Connecticut began replacing its eight counties with nine county-equivalent planning regions. This dataset contains the eight "legacy" county regions of Connecticut.





\newpage

**\huge birth_fips **
\normalsize
\vspace{12pt}

**Label**: Place of birth, county FIPS code

**Description**: birth_fips is a character variable reporting a person's county FIPS code of birth. FIPS codes are 5 digits, of which the first two are the FIPS code of the state to which the county belongs. For a list of modern FIPS codes, see: [\textcolor{blue}{https://en.wikipedia.org/wiki/List\_of\_United\_States\_FIPS\_codes\_by\_county}](https://en.wikipedia.org/wiki/List\_of\_United\_States\_FIPS\_codes\_by\_county). 

Missing values occur if birth city is absent from Numident records, or if the original birth city string was not matched to a GNIS code or FIPS code.

**Notes**: FIPS codes appear numeric but can contain leading zeros. County-level FIPS codes are always 5 digits. In 2022, the state of Connecticut began replacing its eight counties with nine county-equivalent planning regions. This dataset contain FIPS codes for the the eight "legacy" county regions of Connecticut. FIPS codes occasionally change due to county boundary changes, name changes, etc., and thus codes contained in this dataset may not be completely consistent with external county-level data.





\newpage 

**\huge birth_region **
\normalsize
\vspace{12pt}

**Label**: Place of birth, census region

**Description**:  birth_region is a character variable reporting a person's census region of birth. For more information on census regions, please see [\textcolor{blue}{https://www.census.gov/programs-surveys/economic-census/guidance-geographies/levels.html}](https://www.census.gov/programs-surveys/economic-census/guidance-geographies/levels.html). 

NA values are due to an inability to match uncleaned birth city values with the GNIS crosswalk, birth city being outside of a census region (in a different country for example), or a birth city not being entered in social security records.

\vspace{40 pt}

```{r echo = F, message=FALSE, warning=FALSE, fig.pos="H", results='asis'}

birthregion_table <- table(Numident_Geography_File$birth_region, useNA = "ifany")

birthregion_table <- as.data.frame(birthregion_table)

birthregion_table$Freq_perc <- round((birthregion_table$Freq/sum(birthregion_table$Freq))*100, digits = 1)
birthregion_table$label <- c("NA", "Midwest", "Northeast", "South", "West")

br_tab <- kable(birthregion_table %>% select(Var1, label, Freq, Freq_perc), 
      caption = "Births by Census Region",
      col.names = c("birth_region", "label", "n","freq %"),
      align = c("l", "r","l"),
      booktabs = TRUE,
      format = "latex") %>% 
  kableExtra::kable_styling(latex_options = "HOLD_position")

br_tab
```

\newpage 

**\huge death_zip **
\normalsize
\vspace{12pt}

**Label**: Place of death, ZIP code

**Description**: death_zip is a character variable that represents the 5-digit ZIP code of the last residence, as obtained from Numident Death records

Missing values may occur if ZIP code was entered incorrectly in Numident records, if ZIP code is absent from Numident records, or if place of death occured outside USPS identifiable ZIP codes (e.g., most locations abroad). However, Non-identifiable codes such as "XX768" and "00000" may remain in this field.


**Notes**:  ZIP codes appear numeric, but can contain leading zeros. They should always be 5 digits.

We note that ZIP codes are primarily used for USPS mail delivery purposes, not as geographic units. For a more detailed description of ZIP codes, please see: [\textcolor{blue}{https://faq.usps.com/s/article/ZIP-Code-The-Basics}](https://faq.usps.com/s/article/ZIP-Code-The-Basics). 

\vspace{75 pt}
```{r echo = F, message=FALSE, warning=FALSE, fig.pos="H", results='asis'}
# Your existing barplot code
death_props_plot <- barplot(death_props, names.arg = c("Have a death ZIP code", "ZIP Code matched to county"),
        xlab = "Match Percentage", ylab = "Percentage of CenSoc-Numident Records",
        main = "Death ZIP Match Percentage",
        col = c("lightblue", "grey"),
        ylim = c(0, 100)) # Set the y-axis range to 0-100

# Calculate the bar centers
bar_centers <- death_props_plot

# Define the value labels
#value_labels <- c("90%", "89%")
value_labels <- paste0(round(death_props,1), '%')

text(x = bar_centers, y = birth_props, labels = value_labels, col = "black", cex = 1, pos = 1, offset = 1.5)
```


\newpage 


**\huge death_city **
\normalsize
\vspace{12pt}

**Label**: Place of death, city

**Description**: death_city is a character variable reporting a person's city of death, as determined from the ZIP code of their residence at time of death. For ZIP codes that cross city lines, we report the city where the primary post office for that ZIP code is located. 

Missing values may occur if ZIP code was entered incorrectly in Numident records, if ZIP code is absent from Numident records, or if place of death could not be matched to an USPS identifiable ZIP code.






\newpage 


**\huge death_county **
\normalsize
\vspace{12pt}

**Label**: Place of death, county

**Description**: death_county is a character variable reporting a person's county of death, as determined by the ZIP code of their residence at time of death. 

NA values are due to ZIP code not being entered correctly in social security records or a ZIP code that could not be matched against USPS records.

**Note**: In 2022, the state of Connecticut began replacing its eight counties with nine county-equivalent planning regions. This dataset contains the eight "legacy" county regions for Connecticut.

\vspace{40 pt}

```{r echo = F, results='hide', message=FALSE, warning=FALSE, results='asis'}
numident_men <- filter(numident_merged, sex == 1)
numident_women <- filter(numident_merged, sex == 2)

death_year_match_men <- prop.table(table(numident_men$byear, numident_men$death_place_match), margin = 1)

death_year_match_men <- as.data.frame.matrix(death_year_match_men)

death_year_match_men$byear <- rownames(death_year_match_men)

death_year_match_men <- reshape2::melt(death_year_match_men, id.vars = "byear", variable.name = "match", value.name = "proportion")

death_year_match_men$byear_numeric <- as.numeric(as.character(death_year_match_men$byear))

death_year_match_women <- prop.table(table(numident_women$byear, numident_women$death_place_match), margin = 1)

death_year_match_women <- as.data.frame.matrix(death_year_match_women)

death_year_match_women$byear <- rownames(death_year_match_women)

death_year_match_women <- reshape2::melt(death_year_match_women, id.vars = "byear", variable.name = "match", value.name = "proportion")

death_year_match_women$byear_numeric <- as.numeric(as.character(death_year_match_women$byear))

death_year_match_men$gender_category <- "Men"
death_year_match_women$gender_category <- "Women"

birth_year_match_combined_gender <- rbind(death_year_match_men, death_year_match_women)

# Create the combined plot
combined_plot <- birth_year_match_combined_gender %>%
  filter(byear >= 1890 & byear <= 1940) %>% 
  filter(match == 1) %>% 
  ggplot(aes(x = byear_numeric, y = proportion, color = gender_category)) +
  geom_point(size = 1, alpha = 0.5) + # Keep only the points
  labs(title = "Matched Death ZIP Proportions by Gender and Cohort", x = "Birth Year", y = "Proportion") +
  theme_cowplot() +
  scale_color_manual(values = c("blue", "red"), name = "Gender")

# Print the combined plot
print(combined_plot)
```




\newpage 

**\huge death_county_fips **
\normalsize
\vspace{12pt}

**Label**: Place of death, county FIPS code

**Description**: death_fips is a character variable reporting a person's county FIPS code of birth. FIPS codes are 5 digits, of which the first two are the FIPS code of the state to which the county belongs. For a list of modern FIPS codes, see: [\textcolor{blue}{https://en.wikipedia.org/wiki/List\_of\_United\_States\_FIPS\_codes\_by\_county}](https://en.wikipedia.org/wiki/List\_of\_United\_States\_FIPS\_codes\_by\_county). 

Missing values may occur if ZIP code was entered incorrectly in Numident records, if ZIP code is absent from Numident records, if place of death could not be matched to an identifiable USPS identifiable ZIP code, or if county could not be matched to a FIPS code. County FIPS codes are not available for deaths occurring in Puerto Rico.


**Notes**: FIPS codes appear numeric but can contain leading zeros. County-level FIPS codes are are always 5 digits. In 2022, the state of Connecticut began replacing its eight counties with nine county-equivalent planning regions. This dataset contain FIPS codes for the the eight "legacy" county regions of Connecticut. FIPS codes occasionally change due to county boundary changes, name changes, etc., and thus codes contained in this dataset may not be completely consistent with external county-level data.


```{r echo = T, message=FALSE, warning=FALSE, fig.pos="H", results='asis', include = F}
#Hypothetical education data
education <- read_excel("/data/censoc/workspace/geography_variables/Education.xlsx") 

#we rename FIPS to death_fips so that it can match
colnames(education)[1] <- "death_fips" 

#recode Miami-Dade's FIPS to Dade County's FIPS
education <- education %>%
  mutate(death_fips = ifelse(death_fips == "12086", "12025", death_fips)) %>% 
  mutate(death_fips = as.character(death_fips))

#merge onto Numident 
numident_merged <- merge(numident_merged, education, by = "death_fips", all.x = TRUE) 
```



\newpage 


**\huge death_state **
\normalsize
\vspace{12pt}

**Label**: Place of death, state

**Description**: death_state is a character variable reporting a person's state of death, as determined by the ZIP code of their residence at time of death. Includes the 50 U.S states, the District of Columbia, and Puerto Rico.

Missing values may occur if ZIP code was entered incorrectly in Numident records, if ZIP code is absent from Numident records, or if place of death could not be matched to an identifiable USPS identifiable ZIP code in the United States.

\vspace{30 pt}

```{r, echo = F}
state_key <- data.frame(death_state = tolower(state.abb), label = state.name) %>% 
  rbind(c("pr", "Puerto Rico")) %>% 
  rbind(c("dc", "Washington DC"))
  
  
#death_state_tab 
tab <-  Numident_Geography_File %>% group_by(death_state) %>%
  tally() %>% 
  left_join(state_key, by = "death_state") %>% 
  mutate(`freq %` = round(n*100 / sum(n), 1)) %>% 
  select(death_state, label, n, `freq %`)
rows <- seq_len(nrow(tab) %/% 2)
death_state_tab  <- knitr::kable(list(tab[rows, 1:4],
                                      matrix(numeric(), nrow=0, ncol=1),
                                      tab[-rows, 1:4]),
                                 caption = "Deaths by State",
                                 format = "latex",
                                 booktabs = TRUE)  %>%
  kableExtra::kable_styling(latex_options = "HOLD_position")

death_state_tab

```







\newpage 

**\huge death_region **
\normalsize
\vspace{12pt}

**Label**: Place of death, census region

**Description**: death_region is a character variable reporting a person's census region of death, as sourced from the ZIP code of their residence at time of death. This variable is available for persons dying withing the 50 United States and the District of Columbia.

Missing values may occur if ZIP code was entered incorrectly in Numident records, if ZIP code is absent from Numident records,if place of death could not be matched to an identifiable USPS identifiable ZIP code in the United States.

\vspace{70 pt}


```{r echo = F, results='asis'}
deathregion_table <- table(Numident_Geography_File$death_region, useNA = "ifany")

deathregion_table <- as.data.frame(deathregion_table)

deathregion_table$Freq_perc <- round((deathregion_table$Freq/sum(deathregion_table$Freq))*100, digits = 1)
deathregion_table$label <- c("NA", "Midwest", "Northeast", "South", "West")

region_tab <- knitr::kable(deathregion_table %>% select(Var1, label, Freq, Freq_perc), 
      caption = "Deaths by Census Region",
      col.names = c("birth_region", "label", "n","freq %"),
      align = c("l", "r","l"),
      format = "latex",
      booktabs = TRUE) %>% 
  kable_styling(latex_options = "hold_position")

region_tab
```

\newpage 

**\huge death_country **
\normalsize
\vspace{12pt}

**Label**: Place of death, country

**Description**: death_country is a character variable reporting a person's country of death, as determined by the ZIP code of their residence at time of death. While the vast majority of ZIP codes are within the U.S., countries that host U.S. military bases on their territories may also have ZIP codes associated with them. Codes follow [\textcolor{blue}{ISO 3166-1 alpha-2 country codes}](https://www.iso.org/obp/ui/#search/code/).

Missing values may occur if ZIP code being entered incorrectly in Numident records, if ZIP code is absent from Numident records,if place of death could not be matched to an identifiable USPS identifiable ZIP code. 

\newpage 

**\huge death_ruc1993 **
\normalsize
\vspace{12pt}

**Label**: County of death, rural-urban continuum code

**Description**: death_ruc1993 report the 1993 Rural-Urban continuum code for a person's county (or county equivalent) of death, as sourced from the United States Department of Agriculture (USDA). The Urban-Rural Continuum Codes are a classification system developed to categorize U.S. counties based on their degree of urbanization and adjacency to metropolitan areas. The system consists of ten codes, ranging from 0 to 9, where lower codes represent more urbanized counties and higher codes represent more rural areas. 

Missing values may occur if ZIP code being entered incorrectly in Numident records, if ZIP code is absent from Numident records, or if ZIP code could not be matched to a county.

For more information, see: \textcolor{blue}{<https://www.ers.usda.gov/data-products/rural-urban-continuum-codes.aspx>}

\vspace{40 pt}

```{r echo = F, results='hide', message=FALSE, warning=FALSE, results='asis'}
ruc1993_table <- table(Numident_Geography_File$death_ruc1993, useNA = "always")

ruc1993_table <- as.data.frame(ruc1993_table)

ruc1993_table <- ruc1993_table %>% 
  mutate(Rural_Urban_Label = factor(Var1, levels = c("0", "1", "2", "3", "4", "5", "6", "7", "8", "9", "NA"),
                                    labels = c("Central counties of metro areas of 1 million population or more", "Fringe counties of metro areas of 1 million population or more", "Counties in metro areas of 250,000 to 1 million population", "Counties in metro areas of fewer than 250,000 population", "Urban population of 20,000 or more, adjacent to a metro area", "Urban population of 20,000 or more, not adjacent to a metro area","Urban population of 2,500 to 19,999, adjacent to a metro area","Urban population of 2,500 to 19,999, not adjacent to a metro area","Rural or fewer than 2,500 urban population, adjacent to a metro area","Rural or fewer than 2,500 urban population, not adjacent to a metro area", "NA")))

ruc1993_table$Freq_perc <- round((ruc1993_table$Freq/sum(ruc1993_table$Freq))*100,digits = 1)

ruc1993_table <- ruc1993_table[, c(1, 3, 2, 4)]
```

```{r echo = F, results='asis'} 
ructab <- knitr::kable(ruc1993_table, 
      caption = "Rural-Urban County of Death",
      col.names = c("death_ruc1993", "label", "n","freq %"),
      align = c("l", "r", "l","l"),
      format = "latex",
      booktabs = TRUE) %>% 
  kable_styling(latex_options = "hold_position")
  

ructab
```
