---
title: CenSoc-Numident Codebook[^updated]
output:
  pdf_document: default
header-includes:
- \usepackage{hyperref}
- \usepackage{caption}
---
[^updated]: Last updated: `r format(Sys.time(), '%d %B, %Y')`

\captionsetup[table]{labelformat=empty}

| Page| Variable     | Label                                        |
|----:|:-----------------------|:---------------------------------------------|
|   \hyperlink{page.2}{2}  | \hyperlink{page.2}{HISTID}        |Historical unique identifier                  |
|   \hyperlink{page.3}{3}  | \hyperlink{page.3}{byear}         |Year of birth                                 |
|   \hyperlink{page.4}{4}  | \hyperlink{page.4}{bmonth}        |Month of birth                                |
|   \hyperlink{page.5}{5}  | \hyperlink{page.5}{dyear}         |Year of death                                 |
|   \hyperlink{page.6}{6}  | \hyperlink{page.6}{dmonth}        |Month of death                                |
|   \hyperlink{page.7}{7}  | \hyperlink{page.7}{death\_age}    |Age at death (years)                          |
|  \hyperlink{page.8}{8}   |\hyperlink{page.8}{sex}              |Sex                                        |
|  \hyperlink{page.9}{9}   |\hyperlink{page.9}{race\_first}       |Race on First Application                  |
|  \hyperlink{page.10}{10} |\hyperlink{page.10}{race\_first\_cyear} |First Race: Application Year               |
|  \hyperlink{page.11}{11} |\hyperlink{page.11}{race\_first\_cmonth}|First Race: Application Month              |
|  \hyperlink{page.12}{12} |\hyperlink{page.12}{race\_last}        |Race on Last Application                   |
|  \hyperlink{page.13}{13} |\hyperlink{page.13}{race\_last\_cyear}  |Last Race: Application Year                |
|  \hyperlink{page.14}{14} |\hyperlink{page.14}{race\_last\_cmonth} |Last Race: Application Month               |
|  \hyperlink{page.15}{15} |\hyperlink{page.15}{bpl}              |Place of Birth                             |
|  \hyperlink{page.17}{17} |\hyperlink{page.17}{zip\_residence}    |ZIP Code of Residence at Time of Death     |
|  \hyperlink{page.18}{18} |\hyperlink{page.18}{socstate}         |State where Social Security Number Issued  |
|  \hyperlink{page.19}{19} |\hyperlink{page.19}{age\_first\_application}    |Age at First Social Security Application   |
|   \hyperlink{page.20}{20}  | \hyperlink{page.20}{link\_abe\_exact\_conservative} |Flag for conservative ABE match |
|   \hyperlink{page.21}{21}  | \hyperlink{page.21}{weight}        |CenSoc weight                                 |
|   \hyperlink{page.22}{22}  | \hyperlink{page.22}{weight\_conservative} |CenSoc weight (Conservative Sample)  |
|   \hyperlink{page.23}{23}  | \hyperlink{page.23}{Additional IPUMS variables}| Additional variables include `pernum`, `perwt`, `age`, `mbpl`, `fbpl`, `educd`, `empstatd`, `hispan`, `inconwg`, `marst`, `nativity`, `occ`, `occscore`, `ownership`, `race`, `rent`, `serial`, `statefip`, `urban`|  


\vspace{100pt}

**Summary:** The CenSoc-Numident Demo dataset (N = 85,865) was constructed by (i) linking the CenSoc-Numident dataset to the IPUMS 1940 1% census sample and (ii) selecting a set of 20 mortality covariates from the 1940 census. The smaller size of the file — approximately 1% of the records in the full CenSoc-Numident dataset — makes it easier to work with. However, the demo file isn’t conducive to high-resolution mortality research, and we recommend working with the complete CenSoc-Numident file for any final analysis.  


\newpage 

**\huge HISTID**
\normalsize
\vspace{12pt}

**Label**: Historical Unique Identifier

**Description**: HISTID is a unique individual-level identifier. It can be used to merge the CenSoc-Numident file with the 1940 Full-Count Census from IPUMS. 

\newpage 

**\huge byear **
\normalsize
\vspace{12pt}

**Label**: Birth Year

**Description**: byear reports a person's year of birth, as recorded in the Numident death records.  

```{r echo = F, results='hide', message=FALSE, warning=FALSE}
## Library Packages
library(tidyverse)
library(data.table)
library(kableExtra)

## read in censoc_numident_v2.1 data file
# censoc_numident_v2.1 <- read_csv("/data/josh/CenSoc/data_release_v2.1/numident_v2.1/censoc_numident_v2.1.csv")
censoc_numident_v2.1<-read_csv("~/censoc_numident_demo_v2.1.csv") # Numident file I created in codebook demo

censoc_numident_v2.1 %>% 
  count(bpl, bpl_string)
```

```{r, include = F}
byear_plot <- censoc_numident_v2.1 %>%
    group_by(byear) %>%
    summarise(n = n()) %>%
    ggplot(aes(x = byear, y = n)) + 
    geom_line() + 
    geom_point() +
    theme_minimal(base_size = 15) +
  ggtitle("Year of Birth") + 
  theme(legend.position="bottom") +
  xlab("title") + 
  labs(x = "Year", 
       y = "Count") + 
  # scale_y_continuous(labels = scales::comma, limits = c(0, 5000)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n=5)) 
```

\vspace{75pt}

```{r, echo = F}
byear_plot
```


\newpage 
**\huge bmonth **
\normalsize
\vspace{12pt}

**Label:** Birth Month

**Description:** bmonth reports a person's month of birth, as recorded in the Numident death records.  

```{r, include = F}
## run in the console and copy and paste into documentation
bmonth_tabulated <- censoc_numident_v2.1 %>% 
    group_by(bmonth) %>%
    tally() %>%
    mutate(freq = signif(n*100 / sum(n), 2)) %>%
    mutate(label = c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")) %>%
    select(bmonth, label, n, `freq %` = freq) %>% 
  knitr::kable(format = "pipe")
```

\vspace{30pt}

```{r, echo = F}
bmonth_tabulated
```

\newpage 

**\huge dyear **
\normalsize
\vspace{12pt}

**Label**: Death Year

**Description**: dyear reports a person's year of death, as recorded in the Numident death records.     

```{r, include = F, message = F}
dyear_plot <- censoc_numident_v2.1 %>%
    group_by(dyear) %>%
    summarise(n = n()) %>%
    ggplot(aes(x = dyear, y = n)) + 
    geom_line() + 
    geom_point() +
    theme_minimal(base_size = 15) +  
  ggtitle("Year of Death") + 
  theme(legend.position="bottom") +
  xlab("title") + 
  labs(x = "Year", 
       y = "Count") + 
  scale_y_continuous(labels = scales::comma) + 
 # scale_y_continuous(labels = scales::comma, limits = c(0, 7000)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n=5)) 
#The minimum y is 3000 deaths a year when the scale is removed
```

\vspace{75pt}

```{r, echo = F}
dyear_plot
```

\newpage 

**\huge dmonth **
\normalsize
\vspace{12pt}

**Label**: Death Month

**Description**: dmonth reports a person's month of death, as recorded in the Numident death records.  

```{r, include = F}
dmonth_tabulated <- censoc_numident_v2.1 %>%
    group_by(dmonth) %>%
    tally() %>%
    mutate(freq = signif(n*100 / sum(n), 2)) %>%
    mutate(label = c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")) %>%
    select(dmonth, label, n, `freq %` = freq) %>% 
  knitr::kable(format = "pipe")
```

\vspace{30pt}

```{r, echo = F}
dmonth_tabulated
```

\newpage 

**\huge death\_age **
\normalsize
\vspace{12pt}

**Label**: Age at Death (Years)

**Description**: death_age reports a person's age at death in years, calculated using the birth and death information recorded in the Numident death records.    

```{r, include = F}
death_age_plot <- censoc_numident_v2.1 %>%
    group_by(death_age) %>%
    summarise(n = n()) %>%
    ggplot(aes(x = death_age, y = n)) + 
    geom_line() + 
    geom_point() +
    theme_minimal(base_size = 15) + 
  ggtitle("Age at Death") + 
  theme(legend.position="bottom") +
  xlab("title") + 
  labs(x = "Age at Death", 
       y = "Count") + 
  scale_y_continuous(labels = scales::comma) + 
  scale_x_continuous(breaks = scales::pretty_breaks(n=5)) 
```

\vspace{75pt}

```{r, echo = F}
death_age_plot
```

\newpage


**\huge sex **
\normalsize
\vspace{12pt}

**Label:** Sex

**Description**: sex reports a person's sex, as recorded in the Numident death, application, or claim records. 

```{r, include = F}
sex_tabulated <- censoc_numident_v2.1 %>%
    group_by(sex) %>%
     tally() %>%
     mutate(freq = signif(n*100 / sum(n), 3)) %>%
     mutate(label = c("Men", "Women")) %>%
     select(sex, label, n, `freq %` = freq) %>% 
  knitr::kable(format = "pipe")
```

\vspace{30pt}

```{r, echo = F}
sex_tabulated
```


\newpage

\newpage 

**\huge race\_first **
\normalsize
\vspace{12pt}

**Label**: race first

**Description**: race_first reports a person's race, as recorded on their first application entry.

**Note**: Before 1980, the race schema in the Social Security application form contained three categories: White, Black, and Other. In 1980, the SSA added three categories: (1) Asian, Asian American, or Pacific Islander, (2) Hispanic, and (3) North American Indian or Alaskan Native. The Other category was also removed. 

```{r, include = F}
race_first_tabulated <- censoc_numident_v2.1 %>%
    group_by(race_first) %>%
    tally() %>%
    mutate(freq = signif(n*100 / sum(n), 3)) %>%
    mutate(label = c("White", "Black", "Other", "Asian", "Hispanic", "North American Native", "Missing")) %>%
    select(race_first, label, n, `freq %` = freq) %>% 
  knitr::kable(format = "pipe")
```

\vspace{30pt}

```{r, echo = F}
race_first_tabulated
```

\newpage

**\huge race_first_cyear **
\normalsize
\vspace{12pt}

**Label:** First Race: Application Year

**Description**: race_first_cyear is a numeric variable reporting the year of the application on which a person reported their first race.

\vspace{12pt}

\newpage 

**\huge race_first_cmonth **
\normalsize
\vspace{12pt}

**Label:** First Race: Application Month

**Description**: race_first_cmonth is a numeric variable reporting the month of the application on which a person reported their first race. 

\newpage 


**\huge race\_last **
\normalsize
\vspace{12pt}

**Label**: race last

**Description**: race_last reports a person's race, as recorded on their most recent application entry.

**Note**: Before 1980, the race schema in the Social Security application form contained three categories: White, Black, and Other. In 1980, the SSA added three categories: (1) Asian, Asian American, or Pacific Islander, (2) Hispanic, and (3) North American Indian or Alaskan Native. They also removed the Other category.

```{r, include = F}
race_last_tabulated <- censoc_numident_v2.1 %>%
    group_by(race_last) %>%
    tally() %>%
    mutate(freq = signif(n*100 / sum(n), 3)) %>%
    mutate(label = c("White", "Black", "Other", "Asian", "Hispanic", "North American Native", "Missing")) %>%
    select(race_last, label, n, `freq %` = freq) %>% 
  knitr::kable(format = "pipe")
```

\vspace{30pt}

```{r, echo = F}
race_last_tabulated
```

\newpage

**\huge race_last_cyear **
\normalsize
\vspace{12pt}

**Label:** First Race: Application Year

**Description**: race_last_cyear reports the year of the application on which a person reported their last race. 

\newpage 

**\huge race_last_cmonth **
\normalsize
\vspace{12pt}

**Label:** Last Race: Application Month

**Description**: race_last_cmonth is a numeric variable reporting the month of the application on which a person reported their last race.

\newpage

**\huge bpl **
\normalsize
\vspace{12pt}

**Label**: Birthplace

**Description**: bpl is a numeric variable reporting a person's place of birth, as recorded in the Numident application or claims records. The accompanying bpl_string variable reports the person's place of birth as a character string. The coding schema matches the **detailed** IPUMS-USA Birthplace coding schema. 

For a complete list of IPUMS Birthplace codes, please see: https://usa.ipums.org/usa-action/variables/BPL

```{r echo = F, message=FALSE, warning=FALSE, fig.pos="H"}
bpl_tabulation <- censoc_numident_v2.1 %>%
    filter(bpl < 10000 | is.na(bpl)) %>%
    group_by(bpl, bpl_string) %>%
    tally() %>%
    ungroup() %>%
    mutate(freq = round(n*100 / sum(n), 2)) %>%
    select(bpl, bpl_string, n, `freq %` = freq)

rows <- seq_len(nrow(bpl_tabulation) %/% 2)

bpltable<-knitr::kable(list(bpl_tabulation[rows,1:4],
           matrix(numeric(), nrow=0, ncol=1),
           bpl_tabulation[-rows, 1:4]),
      caption = "BPL Tabulation (Native born only)",
      label = "tables", format = "latex", booktabs = TRUE)  %>%
      # add_header_above(c("bpl",  "bpl_string", "n",  "freq")) %>% 
  kableExtra::kable_styling(latex_options = c("HOLD_position"))

# knitr::kable(list(socstate_tabulation[rows,1:4],
#            matrix(numeric(), nrow=0, ncol=1),
#            socstate_tabulation[-rows, 1:4]),
#       caption = "Tabulation of socstate (Native born only)",
#       label = "tables", format = "latex", booktabs = TRUE)  %>%
#   kableExtra::kable_styling(latex_options = "HOLD_position")
# %>%  kableExtra::kable_styling(latex_options = c("HOLD_position")) #doesn't go past one page
```
\newpage  

```{r echo = F, message=FALSE, warning=FALSE, fig.pos="H"}
bpltable
```


\newpage 


**\huge zip\_residence **

\normalsize

\vspace{12pt}

**Label**: ZIP Code of Residence at Time of Death

**Description**: zip_residence is a string variable (9-characters) reporting a person's ZIP Code of residence at time of death, as recorded in the Numident death records.

\newpage 
**\huge socstate**

\normalsize

\vspace{12pt}

**Label**: State where Social Security Number Issued

**Description**: The state in which a person's social security card was issued. Determined by first three (3) digits of Social Security number, as recorded in Numident death records. The accompanying socstate_string variable reports the state in which a person's social security card was issued as a character string. The coding schema matches the detailed IPUMS-USA Birthplace coding schema.

\vspace{30pt}

```{r echo = F, message=FALSE, warning=FALSE, fig.pos="H"}
socstate_tabulation <- censoc_numident_v2.1 %>%
    filter(socstate < 10000 | is.na(socstate)) %>%
    group_by(socstate, socstate_string) %>%
    tally() %>%
    ungroup() %>%
    mutate(freq = round(n*100 / sum(n), 2)) %>%
    select(socstate, socstate_string, n, `freq %` = freq)

rows <- seq_len(nrow(socstate_tabulation) %/% 2)

knitr::kable(list(socstate_tabulation[rows,1:4],
           matrix(numeric(), nrow=0, ncol=1),
           socstate_tabulation[-rows, 1:4]),
      caption = "Tabulation of socstate (Native born only)",
      label = "tables", format = "latex", booktabs = TRUE)  %>%
  kableExtra::kable_styling(latex_options = "HOLD_position")
```

\newpage 

**\huge age\_first\_app **

\normalsize

\vspace{12pt}

**Label**: Age at First Social Security Application 

**Description**: age_first_application reports the age at which a person submitted their first Social Security Application. 

```{r, include = F}
age_first_app_plot <- censoc_numident_v2.1 %>%
  group_by(age_first_application) %>%
  filter(age_first_application %in% c(0:110)) %>% 
  summarise(n = n()) %>%
  ggplot(aes(x = age_first_application, y = n)) + 
  geom_point() +
  geom_line() + 
  theme_minimal(15) + 
  theme(legend.position="bottom") +
  labs(title = "Age of First Application",
       x = "Age of First Application", 
       y = "Count") + 
  scale_y_continuous(labels = scales::comma) + 
  scale_x_continuous(breaks = scales::pretty_breaks(n=5)) 
```

\vspace{75pt}

```{r, echo = F}
age_first_app_plot
```


\newpage

**\huge link\_abe\_exact\_conservative**

\normalsize

\vspace{12pt}

**Label**: Flag for conservative ABE match

**Description**: A flag variable reporting whether a match was established with the ABE conservative match with exact names.  

**Note**: All matches were established with the standard ABE match. A subset of these records were also matched with the conservative ABE match.

```{r, include = F}
link_abe_exact_conservative_tabulated <- censoc_numident_v2.1 %>%
  group_by(link_abe_exact_conservative) %>%
  tally() %>%
  mutate(freq = signif(n*100 / sum(n), 3)) %>%
  arrange(desc(link_abe_exact_conservative)) %>% 
  mutate(label = c("Match Established with Conservative ABE Algorithm", "Not Established")) %>%
  select(link_abe_exact_conservative, label, n, `freq %` = freq) %>% 
  knitr::kable(format = "pipe")
```

\vspace{50pt}

```{r, echo = F}
link_abe_exact_conservative_tabulated
```

\newpage

**\huge weight**
\normalsize

**Label**: Sample Weight [^1]  

[^1]: The IPUMS-USA 1940 1% sample also includes a weight (`perweight`) to account for the 1940 sampling procedure (thus no weights for the 100% complete count 1940 census). For analysis, we recommend using both sets of weights. A final weight can be constructed by multiplying the two weights together.  

\vspace{12pt}

**Description**: A post-stratification person-weight to Human Mortality Database (HMD) totals for persons (1) born between 1895-1939 (2) dying between 1988-2005 (3) dying between ages 65-100. Please see the CenSoc Methods Protocol for more details on weighting procedure.

```{r echo = F, results='hide', message=FALSE, warning=FALSE}
weights_tabulated <- censoc_numident_v2.1 %>%
  filter(!is.na(weight)) %>% 
  summarize('Min Weight' = round(min(weight),2), 'Max Weight' = round(max(weight), 2)) %>%
  mutate(id = 1:n()) %>% 
  pivot_longer(-id, names_to = "Label", values_to = "Value") %>% 
  select(Value, Label) %>% 
  add_row(Label = "No Weight Assigned", Value = NA) %>% 
  knitr::kable(format = "markdown")

weights <- censoc_numident_v2.1 %>% 
  filter(!is.na(weight)) %>% 
  group_by(death_age, dyear) %>% 
  summarize(weight = mean(weight))

## plot mortality sex ratio Lexis surface
weights_lexis <- weights %>% 
  ggplot() +
  geom_raster(aes(x = dyear, y = death_age,
                  fill = weight)) +
  ## Lexis grid
  geom_hline(yintercept = seq(65, 100, 10),
             alpha = 0.2, lty = "dotted") +
  geom_vline(xintercept = seq(1985, 2005, 10),
             alpha = 0.2, lty = "dotted") +
  geom_abline(intercept = seq(-100, 100, 10)-1910,
              alpha = 0.2, lty = "dotted") +
  scale_fill_viridis_c(option = "magma") +
  scale_x_continuous("Year", expand = c(0.02, 0),
                     breaks = seq(1988, 2005, 5)) +
  scale_y_continuous("Age", expand = c(0, 0),
                     breaks = seq(65, 100, 10)) +
  guides(fill = guide_legend(reverse = TRUE)) +
  # coord
  coord_equal() +
  # theme
  theme_void() +
  theme(
    axis.text = element_text(colour = "black"),
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 10, angle = 45, hjust = .5), 
    plot.title = element_text(size = 10, vjust = 2),
    legend.text = element_text(size = 10), 
    axis.title=element_text(size = 10,face="bold")
  ) + 
  labs(X = "Year",
       Y = "Age",
       title = "Average weight by age at death and year of death") 
```

\vspace{50pt}
```{r echo = F, message=FALSE, warning=FALSE}
weights_tabulated
```

\vspace{50pt}
```{r echo = F, message=FALSE, warning=FALSE, fig.align="center"}
weights_lexis
```

\newpage 

**\huge weight_conservative**
\normalsize

**Label**: Sample Weights (Conservative Sample)

\vspace{12pt}

**Description**: A post-stratification person-weight to Human Mortality Database (HMD) totals for persons (1) born between 1895-1939 (2) dying between 1988-2005 (3) dying between ages 65-100. Please see the CenSoc Methods Protocol for more details on weighting procedure.

```{r echo = F, message=FALSE, warning=FALSE}

weights_conservative_tabulated <- censoc_numident_v2.1 %>%
  filter(!is.na(weight_conservative)) %>% 
  summarize('Min Weight' = round(min(weight_conservative),2), 'Max Weight' = round(max(weight_conservative), 2)) %>%
  mutate(id = 1:n()) %>% 
  pivot_longer(-id, names_to = "Label", values_to = "Value") %>% 
  select(Value, Label) %>% 
  add_row(Label = "No Weight Assigned", Value = NA) %>% 
  knitr::kable(format = "pipe")

weights_conservative <- censoc_numident_v2.1 %>% 
  filter(!is.na(weight_conservative)) %>% 
  group_by(death_age, dyear) %>% 
  summarize(weight_conservative = mean(weight_conservative))

## plot mortality sex ratio Lexis surface
weights_lexis_conservative <- weights_conservative %>% 
  ggplot() +
  geom_raster(aes(x = dyear, y = death_age,
                  fill = weight_conservative)) +
  ## Lexis grid
  geom_hline(yintercept = seq(65, 100, 10),
             alpha = 0.2, lty = "dotted") +
  geom_vline(xintercept = seq(1985, 2005, 10),
             alpha = 0.2, lty = "dotted") +
  geom_abline(intercept = seq(-100, 100, 10)-1910,
              alpha = 0.2, lty = "dotted") +
  scale_fill_viridis_c(option = "magma") +
  scale_x_continuous("Year", expand = c(0.02, 0),
                     breaks = seq(1988, 2005, 5)) +
  scale_y_continuous("Age", expand = c(0, 0),
                     breaks = seq(65, 100, 10)) +
  guides(fill = guide_legend(reverse = TRUE)) +
  # coord
  coord_equal() +
  # theme
  theme_void() +
  theme(
    axis.text = element_text(colour = "black"),
    axis.text.y = element_text(size=10),
    axis.text.x = element_text(size = 10, angle = 45, hjust = .5), 
    plot.title = element_text(size=10, vjust = 2),
    legend.text = element_text(size = 10), 
    axis.title=element_text(size=10,face="bold")
  ) + 
  labs(X = "Year",
       Y = "Age",
       title = "Average weight by age at death and year of death") 
```

\vspace{50pt}
```{r echo = F, message=FALSE, warning=FALSE}
weights_conservative_tabulated
```

\vspace{50pt}
```{r echo = F, message=FALSE, warning=FALSE, fig.align="center"}
weights_lexis_conservative
```

\newpage
  
**\huge IPUMS 1940 Census Variable**
\normalsize


\vspace{12pt}

The variables below are from the IPUMS-USA 1940 1% census sample. We recommend looking at the terrific documentation on the IPUMS-USA website: [https://usa.ipums.org/usa/index.shtml.](https://usa.ipums.org/usa/index.shtml) 

| Variable     | Label                                        |
|:-------------|:---------------------------------------------|
| pernum |Person number in sample weight|
|perwt |IPUMS person weight[^2] |
|   age |Age in 1940|
|   mbpl |Mother's place of birth|
|   fbpl |Father's place of birth|
|   educd |Educational attainment (detailed)|
|   empstatd |Employment status (detailed)|
|   hispan |Hispanic/Spanish/Latino origin|
|   inconwg |Had non-wage/salary income over $50|
|   marst |Marital status|
|   nativity |Foreign birthplace or parentage|
|   occ |Occupation|
|   occscore |Occupational income score|
|   ownership |Ownership of dwelling (tenure)|
|   race |Race|
|   rent |Montly contract rent|
|   serial |Household serial number|
|   statefip |State of residence 1940|
|   urban |Urban/rural status|


[^2]: The `perweight` accounts for the 1940 sampling procedure to construct the 1% sample, and thus is only available in the 1940 1% sample. For analysis, we recommend using both the IPUMS `perweight` and the CenSoc `weight.` A final weight can be constructed by multiplying the two weights together















