---
title: "Assessing match quality from BUNMD/Numident"
author: "Maria Osborne" 
---

Summary: Assessing match quality in the CenSoc-Numident file. Uses BUNMD as population denominator rather than
1940 census.

# Read in data
```{r}
# published files
library(plyr)
library(data.table)
library(cowplot)
library(ggsci)

pub_numident <- fread('/home/ipums/ugury-ipums/censoc-abe-implementation/data/final-v2-matched-datasets/censoc_numident_v2.csv')
full_bunmd <- list.files(path="/home/ipums/ugury-ipums/censoc-abe-implementation/data/cleaned-bunmd-by-bpl", full.names = TRUE) %>% 
  lapply(read.csv) %>% 
  rbind.fill() 

library(tidyverse)
library(dplyr)
# matched bunmd/census with multiple ID files (this has more rows than the published numident)
matched_bunmd <- fread('/home/ipums/ugury-ipums/censoc-abe-implementation/data/matched-census-bunmd/matched_census_bunmd.csv')
```


# make sure these are the correct data files
```{r}
# look at the number of rows
nrow(matched_bunmd)
nrow(pub_numident)

# make sure that the numident is a subset of those in the matched bunmd
matched_ids <- matched_bunmd$id_A
num_ids <- pub_numident$HISTID

common_nodes <- intersect(matched_ids, num_ids)
length(common_nodes)

setequal(common_nodes, num_ids)

# let's briefly check the id's that DON'T make it into the numident
# why are they matched but not published? (N ~ 2 million)
excluded_nodes <- setdiff(matched_ids, num_ids)
excluded_set <- matched_bunmd %>% filter(id_A %in% excluded_nodes)
head(excluded_set)
# the VAST majority have death dates outside out time window, and the rest have birth year > 1940. OK!
```


# Let's create the sets that we want to use for this analysis
```{r}
# matched sets
bun_numident_std <- matched_bunmd %>% filter(id_A %in% num_ids)
cons_ids <- (pub_numident %>% filter(link_abe_exact_conservative == 1))$HISTID
bun_numident_cons <- matched_bunmd %>% filter(id_A %in% cons_ids)

# make sure these are the correct sample sizes
nrow(bun_numident_std)
nrow(bun_numident_cons)

# data sets by sex
male_numident_std <- bun_numident_std %>% filter(sex_B==1)
female_numident_std <- bun_numident_std %>% filter(sex_B==2)
male_numident_cons <- bun_numident_cons %>% filter(sex_B==1)
female_numident_cons <- bun_numident_cons %>% filter(sex_B==2)

# get ssns for matched nodes
std_ssns <- (matched_bunmd %>% filter(id_A %in% num_ids))$id_B
cons_ssns <- (matched_bunmd %>% filter(id_A %in% cons_ids))$id_B
  
# full bunmd with flags for matches
full_bunmd <- full_bunmd %>% mutate(num_std_flag = ssn %in% std_ssns,
                                          num_cons_flag = ssn %in% cons_ssns)
```


# pare down the bunmd 
```{r}
# most important: people born after 1940, or who do not die in the death coverage window
bunmd_date_filtered <- full_bunmd %>% filter(byear < 1940 | (byear==1940 & bmonth <= 3)) %>% 
  filter(dyear %in% 1988:2005) # N= 32.5 million


# full case information only
bunmd_restricted <- bunmd_date_filtered %>% filter(bpl!=99999 & !is.na(race_first)) # N= 22.3 million
# bpl availible (not complete case)
bunmd_bpl <- bunmd_date_filtered %>% filter(bpl != 99999)


# looking at the proportion native born
table((bunmd_restricted$bpl < 10000)) # about 90% of the people in the full case sample are native born
table((bunmd_date_filtered$bpl < 10000)) # about 70% of the people in the date-restricted sample are native born
table((bunmd_bpl$bpl < 10000)) # 90% native born for those with birthplace info
```



# Male match rates
```{r}


# raw match rates
library(dplyr)
raw_male_standard <- bunmd_date_filtered %>%
  filter(sex==1, byear %in% 1900:1935) %>% 
  group_by(byear) %>% 
  dplyr::summarize(match_rate = round(mean(num_std_flag), 3)) %>% 
  mutate(match_type = "standard")
raw_male_conservative <- bunmd_date_filtered %>%
  filter(sex==1, byear %in% 1900:1935) %>% 
  group_by(byear) %>% 
  dplyr::summarize(match_rate = round(mean(num_cons_flag), 3)) %>% 
  mutate(match_type = "conservative")

# plot raw match rates
raw_bunmd_match_rate_male <- raw_male_standard %>% 
  bind_rows(raw_male_conservative) %>% 
  filter(byear %in% c(1900:1935)) %>% 
  ggplot(aes(x = byear, y = match_rate, color = match_type, shape = match_type)) + 
  geom_line(size=1) + 
  geom_point(size = 2.5) + 
  theme_cowplot() + 
  ggsci::scale_color_lancet() + 
  labs(x = "Birth Cohort",
       y = "Match Rate",
       title = "Raw Match Rate") + 
  ylim(0, .5) + 
  theme(legend.position = "bottom", legend.title = element_blank()) 


# bpl-availble match rates
bpl_male_standard <- bunmd_bpl %>%
  filter(sex==1, byear %in% 1900:1935) %>% 
  group_by(byear) %>% 
  dplyr::summarize(match_rate = round(mean(num_std_flag), 3)) %>% 
  mutate(match_type = "standard")
bpl_male_conservative <- bunmd_bpl %>%
  filter(sex==1, byear %in% 1900:1935) %>% 
  group_by(byear) %>% 
  dplyr::summarize(match_rate = round(mean(num_cons_flag), 3)) %>% 
  mutate(match_type = "conservative")

# plot bpl-availible match rates
bpl_bunmd_match_rate_male <- bpl_male_standard %>% 
  bind_rows(bpl_male_conservative) %>% 
  filter(byear %in% c(1900:1935)) %>% 
  ggplot(aes(x = byear, y = match_rate, color = match_type, shape = match_type)) + 
  geom_line(size=1) + 
  geom_point(size = 2.5) + 
  theme_cowplot() + 
  ggsci::scale_color_lancet() + 
  labs(x = "Birth Cohort",
       y = "Match Rate",
       title = "Birthplace-Available Match Rate") + 
  ylim(0, .5) + 
  theme(legend.position = "bottom", legend.title = element_blank()) 

# combined plots
bunmd_match_rate_combined_male <- cowplot::plot_grid(raw_bunmd_match_rate_male, bpl_bunmd_match_rate_male, labels = "auto")

bunmd_match_rate_combined_male 

ggsave(bunmd_match_rate_combined_male,
        filename = "/home/ipums/mariaosborne-ipums/censoc/bunmd_men_plot.pdf",
        height = 4, width = 10)

```


# Female match rates
```{r}
# raw
raw_female_standard <- bunmd_date_filtered %>%
  filter(sex==2, byear %in% 1900:1935) %>% 
  group_by(byear) %>% 
  dplyr::summarize(match_rate = round(mean(num_std_flag), 3)) %>% 
  mutate(match_type = "standard")
raw_female_conservative <- bunmd_date_filtered %>%
  filter(sex==2, byear %in% 1900:1935) %>% 
  group_by(byear) %>% 
  dplyr::summarize(match_rate = round(mean(num_cons_flag), 3)) %>% 
  mutate(match_type = "conservative")

# raw rates plot
raw_bunmd_match_rate_female <- raw_female_standard %>% 
  bind_rows(raw_female_conservative) %>% 
  filter(byear %in% c(1900:1935)) %>% 
  ggplot(aes(x = byear, y = match_rate, color = match_type, shape=match_type)) + 
  geom_line(size=1) + 
  geom_point(size=2.5) + 
  theme_cowplot() + 
  ggsci::scale_color_lancet() + 
  labs(x = "Birth Cohort",
       y = "Match Rate",
       title = "Raw Match Rate") + 
  ylim(0, .5) + 
  theme(legend.position = "bottom", legend.title = element_blank()) 


# bpl-availible
bpl_female_standard <- bunmd_bpl %>%
  filter(sex==2, byear %in% 1900:1935) %>% 
  group_by(byear) %>% 
  dplyr::summarize(match_rate = round(mean(num_std_flag), 3)) %>% 
  mutate(match_type = "standard")
bpl_female_conservative <- bunmd_bpl %>%
  filter(sex==2, byear %in% 1900:1935) %>% 
  group_by(byear) %>% 
  dplyr::summarize(match_rate = round(mean(num_cons_flag), 3)) %>% 
  mutate(match_type = "conservative")

# bpl-availible plot
bpl_bunmd_match_rate_female <- bpl_female_standard %>% 
  bind_rows(bpl_female_conservative) %>% 
  filter(byear %in% c(1900:1935)) %>% 
  ggplot(aes(x = byear, y = match_rate, color = match_type, shape=match_type)) + 
  geom_line(size=1) + 
  geom_point(size=2) + 
  theme_cowplot() + 
  ggsci::scale_color_lancet() + 
  labs(x = "Birth Cohort",
       y = "Match Rate",
       title = "Birthplace-Available Match Rate") + 
  ylim(0, .5) + 
  theme(legend.position = "bottom", legend.title = element_blank()) 

#combine plots
bunmd_match_rate_combined_female <- cowplot::plot_grid(raw_bunmd_match_rate_female, bpl_bunmd_match_rate_female, labels = "auto")

bunmd_match_rate_combined_female 

 ggsave(bunmd_match_rate_combined_female,
        filename = "/home/ipums/mariaosborne-ipums/censoc/bunmd_women_plot.pdf",
        height = 4, width = 10)
```


# Compare race in matched/unmatched samples (single sex)
set set == 1 or 2
```{r}
unmatched_characteristics <- bunmd_bpl %>% 
  filter(sex==1, byear %in% 1915:1940) %>% 
  group_by(byear) %>% 
  filter(!num_std_flag & !num_cons_flag) %>% 
  dplyr::summarize(p_black = mean(race_first==2, na.rm=T),
            p_white = mean(race_first==1, na.rm=T)) %>% 
  pivot_longer(cols = starts_with("p_"), names_to = "type",
               names_prefix = "p_", values_to = "prop") %>% 
  mutate(category = "Unmatched")

## Calculate sample proportion (standard)
matched_characteristics_standard <- bunmd_bpl %>% 
  filter(sex==1, byear %in% 1915:1940) %>% 
  group_by(byear) %>% 
  filter(num_std_flag) %>% 
  dplyr::summarize(p_black = mean(race_first==2, na.rm=T),
            p_white = mean(race_first==1, na.rm=T)) %>% 
  pivot_longer(cols = starts_with("p_"), names_to = "type",
               names_prefix = "p_", values_to = "prop") %>% 
  mutate(category = "Matched (Standard)")

## Calculate sample proportion (conservative)
matched_characteristics_conservative <- bunmd_bpl %>% 
  filter(sex==1, byear %in% 1915:1940) %>% 
  group_by(byear) %>% 
  filter(num_cons_flag) %>% 
  dplyr::summarize(p_black = mean(race_first==2, na.rm=T),
            p_white = mean(race_first==1, na.rm=T)) %>% 
  pivot_longer(cols = starts_with("p_"), names_to = "type",
               names_prefix = "p_", values_to = "prop") %>% 
  mutate(category = "Matched (Conservative)")

## Rename vars for facets and set factor levels  
matched_characteristics_combined <- unmatched_characteristics %>% 
  bind_rows(matched_characteristics_standard) %>% 
  bind_rows(matched_characteristics_conservative) %>% 
  mutate(type = as.factor(case_when(
    type == "black" ~ "Race: Black",
    type == "white" ~ "Race: White",
  ))) %>% 
  mutate(type = factor(type, levels=c('Race: Black','Race: White')))

matched_characteristics_plot_standard <- matched_characteristics_combined %>% 
   ggplot(aes(x =byear, y = prop, color = category, linetype= category)) + 
  geom_line(size = 1.2) + 
  theme_cowplot() + 
  ggsci::scale_color_lancet() + 
  theme(legend.position = "bottom", legend.title = element_blank()) + 
  facet_wrap(~type) + 
  background_grid() + 
  labs(title = "BUNMD: Comparison of race in matched and unmatched sets (men)",
       y = "Proportion",
       x = "Birth Year") +
  theme(legend.key.width=unit(1.5, "cm"))

matched_characteristics_plot_standard
ggsave(matched_characteristics_plot_standard,
       filename = "/home/ipums/mariaosborne-ipums/censoc/bunmd_race_comp_men.pdf",
       height = 7, width = 10)

```


# look at how often race is missing by cohort
(this is why we only include byear > 1915 in above plots)
```{r}
#missing race data?
bunmd_bpl %>% 
  filter(sex==2, byear %in% 1900:1940, !num_std_flag, !num_cons_flag) %>%
  group_by(byear, is.na(race_first)) %>% tally() %>% mutate(p=n/sum(n)) %>% arrange(byear)

bunmd_bpl %>% 
  filter(sex==2, byear %in% 1900:1940, num_std_flag) %>%
  group_by(byear, is.na(race_first)) %>% tally() %>% mutate(p=n/sum(n)) %>% arrange(byear)

bunmd_bpl %>% 
  filter(sex==2, byear %in% 1900:1940, num_cons_flag) %>%
  group_by(byear, is.na(race_first)) %>% tally() %>% mutate(p=n/sum(n)) %>% arrange(byear)
```



# Tabulations of geographic regions of birth and race 
(for a single sex)
```{r}
bunmd_bpl <- bunmd_bpl %>%
  mutate(race_recode= case_when(race_first == 1 ~ "White",
                                race_first ==2 ~ "Black",
                                race_first %in% 3:6 ~ "Other",
                                TRUE ~ 'Missing')) %>% 
  mutate(region_of_birth = case_when(bpl %in% c(900,2300,2500,3300,4400,5000) ~ 'New England',
                                     bpl %in% c(3400,3600,4200) ~ 'Middle Atlantic',
                                     bpl %in% c(1700,1800,2600,3900,5500) ~ 'East North Central',
                                     bpl %in% c(1900,2000,2700,2900,3100,3800,4600) ~ 'West North Central',
                                     bpl %in% c(1000,1100,1200,1300,2400,3700,4500,5100,5400) ~ 'South Atlantic',
                                     bpl %in% c(100,2100,2800,4700) ~ 'East South Central',
                                     bpl %in% c(500,2200,4000,4010,4800) ~ 'West South Central',
                                     bpl %in% c(400,800,1600,3000,3200,3500,4900,5600) ~ 'Mountain',
                                     bpl %in% c(200,600,1500,4100,5300) ~ 'Pacific',
                                     bpl == 99999 ~ 'Missing',
                                     TRUE ~ 'Foreign Born'))

bunmd_bpl_case <- bunmd_bpl %>% filter(byear %in% 1915:1940, sex==2)
bun_characteristics <- bunmd_bpl_case %>% 
  select(ssn, race_recode, region_of_birth) %>%
  pivot_longer(-ssn) %>% 
  group_by(name, value) %>%
  tally() %>%            
  mutate(prop = round(100*prop.table(n), 1)) %>% 
  rename(n_gen = n, prop_gen = prop)
numident_characteristics_standard <- bunmd_bpl_case %>% 
  filter(num_std_flag) %>% 
  select(ssn, race_recode, region_of_birth) %>%
  pivot_longer(-ssn) %>% 
  group_by(name, value) %>%
  tally() %>%            
  mutate(prop = round(100*prop.table(n), 1)) %>% 
  rename(n_gen_standard = n, prop_standard = prop)
numident_characteristics_conservative <- bunmd_bpl_case %>% 
  filter(num_cons_flag) %>% 
  select(ssn, race_recode, region_of_birth) %>%
  pivot_longer(-ssn) %>% 
  group_by(name, value) %>%
  tally() %>%            
  mutate(prop = round(100*prop.table(n), 1)) %>% 
  rename(n_gen_conservative = n, prop_conservative = prop)
combined_characteristics <- bun_characteristics %>% 
  inner_join(numident_characteristics_standard, by = c("name", "value")) %>% 
  inner_join(numident_characteristics_conservative, by = c("name", "value")) %>% 
  mutate(name = as.factor(name), value = as.factor(value)) %>% 
  mutate(name = factor(name, levels = c( "race_recode", "region_of_birth"))) %>% 
  arrange(name, value) 
```


# create table
```{r}
table_s3 <- gt(data = combined_characteristics) %>% 
  tab_spanner(
    label = "General Pop",
    columns = vars(
      n_gen, prop_gen)) %>% 
  tab_spanner(
    label = "Standard",
    columns = vars(
      n_gen_standard, prop_standard)) %>% 
  tab_spanner(
    label = "Conservative",
    columns = vars(
      n_gen_conservative, prop_conservative)) %>% 
  cols_label(
    "n_gen" = "No.",
    "prop_gen" = "%",
    "n_gen_standard" = "No.",
    "prop_standard" = "%",
    "n_gen_conservative"  = "No.",
    "prop_conservative" = "%",
    value = ""
  ) %>% 
  tab_style(
    style = list(
      cell_text(weight = "bold")),
    locations = cells_row_groups()
  ) %>% 
  opt_row_striping(row_striping = T) %>% 
  cols_align("left") 

table_s3 

table_s3  %>% gtsave("bunmd_table_women.tex", path ="/home/ipums/mariaosborne-ipums/censoc/")
```



