---
title: "1-read-in-data-enlistment"
output: html_document
---
#Adapted from code written by Ugur Yildirim

```{r}
library(plyr)  # rbind.fill
library(readr) # read_fwf, write_csv

```

#Set paths to raw data and outpath
```{r}
raw_data <- "/data/josh/CenSoc/censoc_data/enlistment-records-public/data/raw/ASNEF.FIN.DAT"

outpath <- "/data/josh/CenSoc/censoc_data/enlistment-records-public/code"

```

#Read everything as enlistment
```{r}
enlistment_col <- c("serial_number", "name", "residence_state", "residence_county", "place_of_enlistment", 
                          "date_of_enlistment_ddmmyy", "grade_alpha", "grade_code", "branch_alpha", "branch_code", 
                          "empty_field", "term_or_enlistment", "longevity", 
                          "source", "nativity", "year_of_birth", "race", "education", "civ_occupation", 
                          "marital_status", "height", "weight", "component", "card_number")


width_enlistment = c(8, 24, 2, 3, 4, 6, 4, 1, 3, 2, 
                      1, 1, 3, 
                      1, 2, 2, 1, 1, 3, 1, 2, 3, 1, 10)

data_enlistment <- read_fwf(raw_data, 
                            fwf_widths(width_enlistment, col_names = enlistment_col), 
                            col_types = cols(empty_field        = col_character(), 
                                             term_or_enlistment = col_character(), 
                                             longevity          = col_character(), 
                                             education          = col_character(), 
                                             marital_status     = col_character()),
                            trim_ws = F)
```

# Read data as if everything is E.R.C. Statistical (see p. 46 in 100.1ND_NC.pdf)
``` {r}
widths_erc = c(8, 24, 2, 3, 4, 6, 4, 1, 3, 2, 
               2, 4, 
               1, 2, 2, 1, 1, 3, 1, 2, 3, 1, 9)

col_names_erc <- c("serial_number", "name", "residence_state", "residence_county", "place_of_enlistment", 
                   "date_of_enlistment_ddmmyy", "grade_alpha", "grade_code", "branch_alpha", "branch_code", 
                   "educ_spec", "defer_date_mmyy", 
                   "source", "nativity", "year_of_birth", "race", "education", "civ_occupation", 
                   "marital_status", "height", "weight", "component", "card_number")

data_erc <- read_fwf(raw_data, 
                     fwf_widths(widths_erc, col_names = col_names_erc), 
                     col_types = cols(educ_spec      = col_character(), 
                                      race           = col_character(), 
                                      education      = col_character(), 
                                      marital_status = col_character()),
                     trim_ws = F)

data_erc <- subset(data_erc, serial_number != "ONE OR M") # 9,200,232 --> 9,039,840

```



# Separate Enlistment from E.R.C. Statistical

# Paragraph on p. 46 in 100.1ND_NC.pdf says:

# "Nevertheless, the format of this category of cards changes beginning in 
# column 58. Two new fields are introduced which cause the remaining fields 
# to be off one column when compared to the Enlistment cards. Fortunately, 
# for future interpretation purposes, the E.R.C. cards can be identified by a 
# 3-punch in column 80. Column 80 on the Enlistment cards is not punched."
``` {r}

data_enlistment$erc1 <- ifelse(substr(data_enlistment$card_number, 1, 1) == "3", 1, 0)
data_enlistment$erc2 <- ifelse(substr(data_enlistment$card_number, 1, 1) != " ", 1, 0)
sum(data_enlistment$erc1) / nrow(data_enlistment) # 0.03688627
sum(data_enlistment$erc2) / nrow(data_enlistment) # 0.09166788
enlistment_cards <- subset(data_enlistment, erc1 == 0)

data_erc$erc1 <- ifelse(data_erc$component == "3", 1, 0)
data_erc$erc2 <- ifelse(data_erc$component != " ", 1, 0)
sum(data_erc$erc1) / nrow(data_erc) # 0.03688627
sum(data_erc$erc2) / nrow(data_erc) # 0.09166788
erc_statistical_cards <- subset(data_erc, erc1 == 1)
```

# Merge records back together
```{r}
data_combined <- rbind.fill(enlistment_cards, erc_statistical_cards)
rm(data_enlistment, data_erc, enlistment_cards, erc_statistical_cards)

```

# Save processed dataset
```{r}
write_csv(data_combined, paste0(outpath, "/temp1.csv"))
rm(list = ls())
```

```