---
title: "5-clean-grade-enlistment"
output: html_document
---
#This rmd recodes grade_codes, branch_codes,source,term_or_enlistment, and component

#Load libraries
```{r}
library(dplyr)   
library(tidyverse) 
library(readr)  
library(data.table) 
```

#Set paths to data & csvs of NARA AAD codes
```{r}
path_to_temp_data   <-  "/global/scratch/p2p3/pl1_demography/censoc/workspace/enlistment_records/temp4.csv"
out_path <- "/global/scratch/p2p3/pl1_demography/censoc/workspace/enlistment_records"

grade_codes_path <- "/global/scratch/p2p3/pl1_demography/censoc/input_data/enlistment_records_public/code/awenlistmentcode/cl_2072.csv" 
branch_codes_path <- "/global/scratch/p2p3/pl1_demography/censoc/input_data/enlistment_records_public/code/awenlistmentcode/cl_2075.csv"
source_path <- "/global/scratch/p2p3/pl1_demography/censoc/input_data/enlistment_records_public/code/awenlistmentcode/cl_2077.csv"
branch_alpha_path <- "/global/scratch/p2p3/pl1_demography/censoc/input_data/enlistment_records_public/code/awenlistmentcode/cl_2074 (1).csv"
grade_alpha_path <- "/global/scratch/p2p3/pl1_demography/censoc/input_data/enlistment_records_public/code/awenlistmentcode/cl_2073.csv"
term_or_enlistment_path <- "/global/scratch/p2p3/pl1_demography/censoc/input_data/enlistment_records_public/code/awenlistmentcode/cl_2076.csv"
component_path <- "/global/scratch/p2p3/pl1_demography/censoc/input_data/enlistment_records_public/code/awenlistmentcode/cl_2078.csv"

```

#READ IN TEMP 4
```{r}

data_combined <- read_csv(path_to_temp_data, 
                          col_types = cols(
                            empty_field = col_character(),
                            term_or_enlistment = col_character(),
                            longevity = col_character(),
                            education = col_character(),
                            marital_status = col_character(),
                            component = col_character(),
                            educ_spec = col_character(),
                            defer_date_mmyy = col_character(),
                            mname_clean = col_character(),
                            sex = col_integer(),
                            sex_r = col_integer(),
                            fname_std = col_character()
                            )
                          )
```

#RECODING GRADE_CODE

#Count of grade codes
```{r}
data_combined %>% 
  dplyr::count(grade_code) 

```

#Create function to recode grade codes
```{r}
recode_grade_code_enlistment <- function(df){
  
  ## recode grade variable
  df <- df %>% 
    mutate(grade_code_recode = case_when(
      grade_code == "9" ~ 9, #Aviation Cadet
      grade_code == "6" ~ 6, #Corporal
      grade_code == "2" ~ 2, #First Sergeant
      grade_code == "1" ~ 1, #Master Sergeant
      grade_code == "8" ~ 8, #Private
      grade_code == "7" ~ 7, #Private First Class
      grade_code == "5" ~ 5, #Sergeant
      grade_code == "4" ~ 4, #Staff Sergeant
      grade_code == "3" ~ 3, #Technical Sergeant
      grade_code == "M" ~ 4, #Technician 3d Grade Equivalent to Grade 4: Staff Sergeant
      grade_code == "N" ~ 5, #Technician 4th Grade Equivalent to Code 5: Sergeant
      grade_code == "O" ~ 6, #Technician 5th Grade Equivalent to Code 6: Corporal
      TRUE ~ NA_real_     ## this recodes anything remaining to NA (this is better than blank)
      ))
  
  ## return data.frame   
  return(df)
}


```

#Apply recode_grade_code_enlistment function on DF
```{r}
grade_recode <- recode_grade_code_enlistment(df = data_combined) %>% 
  select(grade_code, grade_code_recode)

```

#RECODING BRANCH_CODES
#Count of branch codes
```{r}
data_combined %>% 
  dplyr::count(branch_code) 
```

#Create function to recode branch codes
```{r}
recode_branch_code_enlistment <- function(df){
  
  ## recode grade variable
  df <- df %>% 
    mutate(branch_code_recode = case_when(
      branch_code == "78" ~ 78, #Adjutant General's Department	
      branch_code == "20" ~ 20, #Air Corps	
      branch_code == "00" ~ 00, #Branch Immaterial - Warrant Officers, USA	
      branch_code == "93" ~ 93, #Cavalry
      branch_code == "96" ~ 96, #Chaplains
      branch_code == "80" ~ 80, #Chemical Warfare Service	
      branch_code == "40" ~ 40, #Coast Artillery Corps or Army Mine Planter Service	
      branch_code == "82" ~ 82, #Corps of Engineers	
      branch_code == "06" ~ 06, #Corps of Military Police	
      branch_code == "74" ~ 74, #Dental Corps - For Officers only	
      branch_code == "08" ~ 08, #Detached Enlisted Men's List	
      branch_code == "30" ~ 30, #Field Artillery	
      branch_code == "97" ~ 97, #Finance Department	
      branch_code == "01" ~ 01, #General Staff Corps	
      branch_code == "72" ~ 72, #Hospital Dietitians	
      branch_code == "09" ~ 09, #Inactive Reserve	
      branch_code == "10" ~ 10, #Infantry	
      branch_code == "98" ~ 98, #Inspector General's Department	
      branch_code == "99" ~ 99, #Judge Advocate General's Department	
      branch_code == "77" ~ 77, #Medical Administrative Corps - For Officers only	
      branch_code == "71" ~ 71, #Medical Corps - For Officers only	
      branch_code == "70" ~ 70, #Medical Department - For Enlisted Men of the Medical Department (includes Enlisted Men of the Medical Department on Duty with the Dental Corps) or Army Nurse Corps	
      branch_code == "05" ~ 05, #Military Intelligence	
      branch_code == "02" ~ 02, #No branch assignment	
      branch_code == "87" ~ 87, #Ordnance Department	
      branch_code == "79" ~ 79, #Pharmacy Corps	
      branch_code == "03" ~ 03, #Philippine Scouts	
      branch_code == "73" ~ 73, #Physical Therapy Aides	
      branch_code == "04" ~ 04, #Professor US Military Academy	
      branch_code == "50" ~ 50, #Quartermaster Corps	
       branch_code == "90" ~ 90, #Signal Corps	
      branch_code == "07" ~ 07, #Specialist Reserve	
      branch_code == "38" ~ 38, #Transportation Corps	
      branch_code == "75" ~ 75, #Veterinary Corps or Service - For Officers of the Veterinary Corps and for Enlisted Men of the Veterinary Service	
      branch_code == "18" ~ 18, #Women's Army Corps	
      TRUE ~ NA_real_     ## this recodes anything remaining to NA (this is better than blank)
      ))
  
  ## return data.frame   
  return(df)
}

```

#Apply recode_branch_code_enlistment function to df
```{r}
branch_recode <- recode_branch_code_enlistment(df = data_combined) %>% 
  select(branch_code, branch_code_recode)

```

#RECODING source

#Count of source

```{r}
data_combined %>% 
  dplyr::count(source) 
```

#Create function to recode source codes
```{r}
recode_source_enlistment <- function(df){
  
  ## recode source variable
  df <- df %>% 
    mutate(source_recode = case_when(
      source == "0" ~ 0, #Civil Life	
      source == "6" ~ 6, #Enlisted Man, Philippine Scout or recall to AD of an enlisted man who had been transferred to the ERC	
      source == "2" ~ 2, #Enlisted Man, Regular Army, after 3 months of Discharge	
      source == "1" ~ 1, #Enlisted Man, Regular Army, within 3 months of Discharge or former WAAC Auxiliary	
      source == "7" ~ 7, #National Guard	
      source == "5" ~ 5, #Enlisted Reserve or Medical Administrative Corps (MAC) Officer
      source == "4" ~ 4, #National Guard in Federal Service, after 3 months of Discharge
      source == "8" ~ 8, #Returned to Military Control, Surrendered	
      source == "9" ~ 9, #Returned to Military Control, Apprehended	
      source == "3" ~ 3, #National Guard in Federal Service, within 3 months of Discharge
      source == "&" ~ 12, #Selective Service after 3 months of Discharge	
      source == "-" ~ 11, #Selective Service within 3 months of Discharge	
      TRUE ~ NA_real_     ## this recodes anything remaining to NA (this is better than blank)
      ))
  
  ## return data.frame   
  return(df)
}

```

#Apply recode_source_enlistment function to df
```{r}
source_recode <- recode_source_enlistment(df = data_combined) %>% 
  select(source, source_recode)
```

#RECODE term_or_enlistment

#Count of term_or_enlistment
```{r}
data_combined %>% 
  dplyr::count(term_or_enlistment) 
```

#Create function to recode term_or_enlistment 
```{r}
recode_term_or_enlistment_enlistment <- function(df){
  
  ## recode grade variable
  df <- df %>% 
    mutate(term_or_enlistment_recode = case_when(
      term_or_enlistment == "1" ~ 1, #One year enlistment
      term_or_enlistment == "2" ~ 2, #Enlistment for Alaska
      term_or_enlistment == "3" ~ 3, #Enlistment for Hawaiian Department
      term_or_enlistment == "4" ~ 4, #Enlistment for assignment to another corps area
      term_or_enlistment == "5" ~ 5, #Enlistment for the duration of the War or..
      term_or_enlistment == "6" ~ 6, #Enlistment for the Panama Canal Department
      term_or_enlistment == "7" ~ 7, #Enlistment for the Philippine Department
      term_or_enlistment == "8" ~ 8, #Enlistment for the Puerto Rican Department
      term_or_enlistment == "9" ~ 9, #Three year enlistment
      TRUE ~ NA_real_     ## this recodes anything remaining to NA (this is better than blank)
      ))
  
  ## return data.frame   
  return(df)
}

```

#Apply recode_term_or_enlistment_enlistment function to df
```{r}
term_or_enlistment_recode <- recode_term_or_enlistment_enlistment(df = data_combined) %>% 
  select(term_or_enlistment, term_or_enlistment_recode)
```

#RECODE component
#Count of component

```{r}
data_combined %>% 
  dplyr::count(component) 
```
#Create function to recode component 
```{r}
recode_component_enlistment <- function(df){
  
  ## recode grade variable
  df <- df %>% 
    mutate(component_recode = case_when(
      component == "1" ~ 1, # Regular Army 
      component == "2" ~ 2, #	Regular Army Reserve
      component == "3" ~ 3, # Reserves - exclusive of Regular Army Reserve
      component == "4" ~ 4, # National Guard
      component == "5" ~ 5, # Philippine Scouts
      component == "6" ~ 6, # Army of the United States
      component == "7" ~ 7, # Selectees
      component == "9" ~ 9, # Women's Army Corps
      TRUE ~ NA_real_     ## this recodes anything remaining to NA (this is better than blank)
      )) %>%
    # Create separate variable that further recodes as either voluntary service or draftee
    mutate(enlistment_status = case_when(
      component_recode < 7 | component_recode == 9 ~ 0, # Voluntary service
      component_recode == 7 ~ 1, # Drafted by Selective Service
      TRUE ~ NA_real_
    ))
  
  ## return data.frame   
  return(df)
}

```
#Apply function to df
```{r}
component_recode <- recode_component_enlistment(df = data_combined) %>% 
  select(component, component_recode, enlistment_status)
```

#Add columns with recoded data to data_combined
```{r}
data_combined <- cbind(data_combined, source_recode$source_recode, grade_recode$grade_code_recode,term_or_enlistment_recode$term_or_enlistment_recode, branch_recode$branch_code_recode, component_recode$component_recode, component_recode$enlistment_status)
```


#Save dataset
```{r}
write_csv(data_combined, paste0(out_path, "/temp5.csv"))
rm(list = ls())
```

