---
title: "6-clean-socioeco-enlistment"
output: html_document
---

#This rmd recodes race, citizenship, education, and marital status

```{r}
library(dplyr)   # %>%, group_by, tally, print
library(ipumsr)  # read_ipums_ddi, ipums_val_labels
library(readr)   # read_csv, write_csv
library(stringr) # str_trim, str_split


```


```{r }
path_to_temp_data   <-  "/data/josh/CenSoc/censoc_data/enlistment-records-public/code/awenlistmentcode/temp5.csv"
out_path <- "/data/josh/CenSoc/censoc_data/enlistment-records-public/code/awenlistmentcode"
race_path <- "/data/josh/CenSoc/censoc_data/enlistment-records-public/code/awenlistmentcode/titlesnicknames/cl_2059.csv"
```

#READ IN TEMP 5
```{r}

data_combined <- read_csv(path_to_temp_data, 
                          col_types = cols(
                            empty_field = col_character(),
                            term_or_enlistment = col_character(),
                            longevity = col_character(),
                            education = col_character(),
                            marital_status = col_character(),
                            component = col_character(),
                            educ_spec = col_character(),
                            defer_date_mmyy = col_character(),
                            mname_clean = col_character(),
                            sex = col_integer(),
                            sex_r = col_integer(),
                            fname_std = col_character()
                            )
                          )
```

#Clean Race

#View race fields
```{r}
data_combined %>% 
  count(race) 
```

```{r}
recode_race_enlistment <- function(df){
  
  ## recode race variable using IPUM codes
  df <- df %>% 
    mutate(race_recode = case_when(
      race == "1" ~ 100, #white citizens/noncitizens
      race == "J" ~ 100, #white citizens/noncitizens
      race == "2" ~ 200, #Black/African American citizens/noncitizens
      race == "K" ~ 200, #Black/African American citizens/noncitizens
      race == "6" ~ 300, #American Indian citizens/noncitizens
      race == "O" ~ 300, #American Indian citizens/noncitizens
      race == "3" ~ 400, #Chinese citizen/noncitizen
      race == "L" ~ 400, #Chinese citizen/noncitizen
      race == "4" ~ 500, #Japanese citizen/noncitizen
      race == "M" ~ 500, #Japanese citizen/noncitizen
      race == "7" ~ 600, #filipino
      race == "P" ~ 600, #filipino citizen/noncitizen
      race == "5" ~ 630, #native hawaiian citizen/noncitizen
      race == "N" ~ 630, #native hawaiianr citizen/noncitizen
      race == "8" ~ 150, #Puerto Ricans
      race == "Q" ~ 150, # Puerto Ricans
      race == "9" ~ 991, #Other
      race == "R" ~ 991, #Other
      TRUE ~ NA_real_     ## this recodes anything remaining to NA (this is better than blank)
      ))
  
  ## return data.frame   
  return(df)
}
```

```{r}
race_recode <- recode_race_enlistment(df = data_combined) %>% 
  select(race, race_recode)
```

#recode citizenship

```{r}
recode_citizenship_enlistment <- function(df){
  
  ## recode citizenship variable using IPUM codes
  df <- df %>% 
    mutate(citizenship_recode = case_when(
      race == "1" ~ 1,
      race == "J" ~ 0, #white citizens/noncitizens
      race == "2" ~ 1,
      race == "K" ~ 0, #Black/African American citizens/noncitizens
      race == "6" ~ 1,
      race == "O" ~ 0, #American Indian citizens/noncitizens
      race == "3" ~ 1,
      race == "L" ~ 0, #Chinese citizen/noncitizen
      race == "4" ~ 1,
      race == "M" ~ 0, #Japanese citizen/noncitizen
      race == "7" ~ 1,
      race == "P" ~ 0,
      race == "5" ~ 1,
      race == "N" ~ 0, #Other Asian/Pacific Islander citizen/noncitizen
      race == "8" ~ 1,
      race == "Q" ~ 0,
      race == "9" ~ 1,
      race == "R" ~ 0, #Other + Puerto Ricans
      TRUE ~ NA_real_     ## this recodes anything remaining to NA (this is better than blank)
      ))
  
  ## return data.frame   
  return(df)
}
```

```{r}
citizenship_recode <- recode_citizenship_enlistment(df = data_combined) %>% 
  select(race, citizenship_recode)
```

## Recode education
```{r}
data_combined %>% 
  count(education) 
```

## write a function to recode education codes to better match ipums codes 

```{r}
recode_education_enlistment <- function(df){
  
  ## recode education variable
  df <- df %>% 
    mutate(education_recode = case_when(
      education == 0 ~ 00,
      education == 1 ~ 01,
      education == 2 ~ 02,
      education == 3 ~ 03,
      education == 4 ~ 04,
      education == 5 ~ 05,
      education == 6 ~ 06,
      education == 7 ~ 07,
      education == 8 ~ 08,
      education == 9 ~ 09,
      TRUE ~ NA_real_     ## this recodes anything remaining to NA (this is better than blank)
    ))
  
  ## return data.frame   
  return(df)
}
```

```{r}
education_recode <- recode_education_enlistment(df = data_combined) %>% 
  select(education, education_recode)
```

#Recode marital_status from NARA AAD TO IPUMS CODES
```{r}
data_combined %>% 
  count(marital_status) 
```
#Create a function to recode marital status to IPUMs codes
## no delineation of 1. married, spouse present & 2.married, spouse absent so all
## married are coded as 1
```{r}
recode_marital_status_enlistment <- function(df){
  
  ## recode grade variable using IPUM codes
  df <- df %>% 
    mutate(marital_status_recode = case_when(
      marital_status == "1" ~ 6, #single w dependents
      marital_status == "2" ~ 1, #married
      marital_status == "3" ~ 3, #separated w dependents
      marital_status == "4" ~ 4, #divorced w dependents
      marital_status == "5" ~ 5, #widowed w dependents
      marital_status == "6" ~ 6, #single
      marital_status == "7" ~ 3, #separated 
      marital_status == "8" ~ 4, #divorced 
      marital_status == "9" ~ 5, #widowed without dep

      TRUE ~ NA_real_     ## this recodes anything remaining to NA (this is better than blank)
      ))
  
  ## return data.frame   
  return(df)
}
```

#Recode marital status
```{r}
marital_status_recode <- recode_marital_status_enlistment(df = data_combined) %>% 
  select(marital_status, marital_status_recode)
```


#Add columns with recoded data to data_combined
```{r}
data_combined <- cbind(data_combined, race_recode$race_recode, citizenship_recode$citizenship_recode, education_recode$education_recode, marital_status_recode$marital_status_recode)
```

#Save dataset
```{r}
write_csv(data_combined, paste0(out_path, "/temp6.csv"))
rm(list = ls())
```


