---
title: "7-clean-weight-enlistment"
output: html_document
---

```{r}
library(dplyr)   # %>%, group_by, tally, print
library(ipumsr)  # read_ipums_ddi, ipums_val_labels
library(readr)   # read_csv, write_csv
library(stringr) # str_trim, str_split
library(ggplot2)
library(scales)
library(data.table)
```


```{r}
path_to_temp_data   <-   "//data/censoc/workspace/enlistment_records/temp6.csv"
out_path <-  "//data/censoc/workspace/enlistment_records"
```


```{r}
data_combined <- read_csv(path_to_temp_data, 
                          col_types = cols(
                            empty_field = col_character(),
                            term_or_enlistment = col_character(),
                            longevity = col_character(),
                            education = col_character(),
                            marital_status = col_character(),
                            component = col_character(),
                            educ_spec = col_character(),
                            defer_date_mmyy = col_character(),
                            mname_clean = col_character(),
                            sex = col_integer(),
                            sex_r = col_integer(),
                            fname_std = col_character()
                            )
                          )
```

#LOOKING AT ALL VARS FOR HT AND WEIGHT
USER NOTE FROM 100.1ND_NC.pdf PG 11.
TM 12-305 (1 May 1943) states "Height in inches (field use as desired) 74-75."
Direction for the Weight field reads, "Weight (AGCT [Army General Classification Test] will be
punched in this field) 76-78." TM 12-305 (1 November 1945) reads "Height in inches (not punched)"
and states "Weight (not punched). Neither set of instructions allows us to fix a date when, if ever, height
and weight were punched into the Enlistment cards. They do tell us that the fields might be used for other
purposes. Height and Weight fields were printed on Enlistment card (WO, AGO Form No. 317), but instructions
for their use or disuse changed during the war. Some versions of the Enlistment card and instructions
indicate that the same fields (columns 74-78) were used for Military Occupation. None of the
instructions clearly allows us to fix a date when, if ever, height and weight or Military Occupation were
punched into the Enlistment cards. For this reason, the data recorded in positions 74-78 are not displayed
in the Access to Archival Databases (AAD) implementation of the Electronic Army Serial Number
Merged File. The raw data electronic records that NARA preserves retains whatever values were
punched in these positions

# convert height and weight to integers
```{r}
data_combined$height <- as.integer(data_combined$height)
data_combined$weight <- as.integer(data_combined$weight)
data_combined <- as.data.table(data_combined)
```

# height distribution
```{r}
summary(data_combined$height) # 0 - 99
# all
hist(data_combined$height, breaks = seq(0,100,1))
#before 1943
hist(data_combined[date_enlistment_recode < "1943-01-01"]$height, breaks = seq(0,100,1))
#Jan - Feb 1943
hist(data_combined[date_enlistment_recode >= "1943-01-01" & date_enlistment_recode <= "1943-03-01"]$height,
     breaks = seq(0,100,1))
# March 1943 +
hist(data_combined[date_enlistment_recode >= "1943-03-01"]$height,
     breaks = seq(0,100,1)) # WAY different distribution
# missing date
hist(data_combined[is.na(date_enlistment_recode)]$height, breaks = seq(0,100,1)) # hard to tell
# dist by height for each month in 1943
data_combined %>% filter(lubridate::year(date_enlistment_recode) == 1943) %>% 
  group_by(lubridate::month(date_enlistment_recode)) %>% 
  ggplot(aes(x = height)) +
  geom_histogram(stat="count") +
  facet_wrap(~ lubridate::month(date_enlistment_recode))
```

# weight/AGCT stats:
```{r}
summary(data_combined$weight) # 0 - 999
# all (cut out 0's to make easier to see)
hist(data_combined[weight > 0]$weight, breaks = seq(0,1000,1)) #huge spike at 100, prob AGCT scores
#before 1943
hist(data_combined[weight > 0 & date_enlistment_recode < "1943-01-01"]$weight, breaks = seq(0,1000,1))
#Jan - Feb 1943
hist(data_combined[weight > 0 & date_enlistment_recode >= "1943-01-01" & date_enlistment_recode <= "1943-03-01"]$weight,
     breaks = seq(0,1000,1))
# March 1943 +
hist(data_combined[weight > 0 & date_enlistment_recode >= "1943-03-01"]$weight,
     breaks = seq(0,1000,1)) # def more characteristic of AGCT scores, but hard to separate the two.
# dist by weight for each month in 1943
data_combined %>% filter(weight > 0) %>% 
  filter(lubridate::year(date_enlistment_recode) == 1943) %>% 
  group_by(lubridate::month(date_enlistment_recode)) %>% 
  ggplot(aes(x = weight)) +
  geom_histogram(stat="count") +
  facet_wrap(~ lubridate::month(date_enlistment_recode))
```

HEIGHT AND WEIGHT OF SELECTIVE SERVICE REGISTRANTS PROCESSED FOR MILITARY
SERVICE DURING WORLD WAR II by Karpinos:
Height between 60 and 78 inches
Minimum weight for a 60 inch person is 105 pounds. (no clear maximum)


# Recode height:
Only retain for February 1943 and earlier
Remove values < 60 and > 78
```{r}
data_combined <- data_combined %>% 
  mutate(height_recode = case_when(date_enlistment_recode < "1943-03-01" ~ height,
                                   TRUE ~ NA_integer_))
data_combined <- data_combined %>% 
  mutate(height_recode = case_when(height_recode < 60 ~ NA_integer_,
                                   height_recode > 78 ~ NA_integer_,
                                   TRUE ~ height_recode))

summary(data_combined$height_recode)
summary(data_combined[date_enlistment_recode < "1943-03-01"]$height_recode)
summary(data_combined[date_enlistment_recode >= "1943-03-01"]$height_recode) # all NA's 
```


# Recode Weight:
Before 1943 = weight
set minimum to 105, maximum to 500

Period of early 1943: some sites did AGCT scores, hard to tell which ones
remove zeros and set max to 500 again
```{r}
data_combined <- data_combined %>% 
  mutate(weight_before_1943 = case_when(date_enlistment_recode < "1943-01-01" ~ weight,
                                        TRUE ~ NA_integer_))
data_combined <- data_combined %>% 
  mutate(weight_before_1943 = case_when(weight_before_1943 < 105 ~ NA_integer_,
                                   weight_before_1943 > 500 ~ NA_integer_,
                                   TRUE ~ weight_before_1943))

summary(data_combined$weight_before_1943)
hist(data_combined$weight_before_1943, breaks = seq(90,510,5))
```

```{r}
data_combined <- data_combined %>% 
  mutate(weight_or_AGCT = case_when(date_enlistment_recode >= "1943-01-01" ~ weight,
                                        TRUE ~ NA_integer_))

data_combined <- data_combined %>% 
  mutate(weight_or_AGCT = case_when(weight_or_AGCT == 0 ~ NA_integer_,
                                   weight_or_AGCT > 500 ~ NA_integer_,
                                   TRUE ~ weight_or_AGCT))

summary(data_combined$weight_or_AGCT)
hist(data_combined$weight_or_AGCT, breaks = seq(0,500,1))
```




```{r}
write_csv(data_combined, paste0(out_path, "/temp7.csv"))
rm(list = ls())
```




#visualize data variables 
#create weight histogram and overlay with height and remove 0s

```{r}
# 
# weight_hist <- data_combined %>% filter(weight < 250) %>% ggplot(mapping = aes(x = weight)) +                           # Modify title & axis labels
#   geom_histogram(stat="count") +
#   labs(title = "Weight/IQ",
#        x = "Weight/IQ",
#        y = "Count of Values") + stat_count() +  scale_y_continuous(labels = scales::comma)
# print(weight_hist)
# ```
# 
# #PLOT HEIGHT/IQ W XLIM
# 
# ```{r}
# height_hist <- data_combined %>% ggplot(mapping =  aes(x = height)) +                           # Modify title & axis labels geom_histogram(stat="count") +
#   labs(title = "Height",
#        x = "Height",
#        y = "Count of Values") + stat_count() +  scale_y_continuous(labels = scales::comma)
# 
# print(height_hist)
# ```
# #OVERLAY HEIGHT AND WEIGHT/IQ GRAPHS
# ```{r}
# set.seed(42)
# plot( weight_hist, col=rgb(0,0,1,1/4), xlim=c(0,10))  # first histogram
# plot( height_hist, col=rgb(1,0,0,1/4), xlim=c(0,10), add=TRUE)  # second
# 
# ```
# #clean height
# ```{r}
# data_combined %>% count(height)
# ```
# ```{r}
# recode_height_enlistment <- data_combined %>% function(df){
# df <- df %>% filter(height => 60 & height <= 100 )
# return(df)}
# ```
# 
# ```{r}
# height_recode <- recode_height_enlistment(df = data_combined) %>% 
#   select(height)
# ```
# 
# #Classify all people with a variable lower than 97 as an AGCT score and higher than 147 as weight 
# ```{r}
# recode_AGCT_wt <- function(df){
#   df <- df%>% mutate(AGCT_wt = case_when(
#     weight <= 97 ,
#     weight >= 147 ~ "AGCT Score",
#     TRUE ~ NA_character_
#   ))
#   return(df)
# }
# ```
# 
# ```{r}
# AGCT_wt_recode <- recode_AGCT_wt(df = data_combined) %>% 
#   select(weight, AGCT_wt)
# ```
# 
# 
# 
# 
# #Filter column by people with enlistment dates prior to March 1943– all those will be weights
# 
# #look at height and weight/AGCT vars
# ```{r}
# data_combined %>% 
#   count(date_of_enlistment_mm) 
# 
# data_combined %>% 
#   count(date_of_enlistment_yy) 
# colnames(data_combined)
# ```
# 
# 
# ```{r}
# recode_weight_pre_43 <- function(df){
# df <- df %>% 
#   mutate(wt_recode = if_else(date_of_enlistment_ddmmyy < "010343", 
#                                 weight, NA_character_))
#   
# 
# return(df)
# }
# ```
# 
# 
# ```{r}
# wt_recode1 <- recode_weight_pre_43(df = data_combined) %>% 
#   select(weight, wt_recode)
# ```
# 
# 
# #plot weight/AGCT with Date_of_enlistment_ddmmyy faceted by place_of_enlistment
# 
# 
# ```{r}
# data_combined %>%
#   count(place_of_enlistment, sort = TRUE) %>% 
#   mutate(freq = round(n / sum(n), 2))
# ```
# 
# #filter out reasonable heights
# ```{r}
# ht_recode_enlistment <-function(df){
#   df <- df %>% mutate(height = as.numeric(height)) %>% mutate(ht_recode = case_when(
#   height > 60 & height < 109 ~ height,
#         TRUE ~ NA_real_ 
# ))
# return(df)}
# ```
# 
# ```{r}
# height_recode1 <- ht_recode_enlistment(df = data_combined) %>% 
#   select(height, ht_recode)
# 
# ```
# 
# 
# ```{r}
# data_combined <- cbind(data_combined, height_recode1$ht_recode, wt_recode1$wt_recode)
# ```
# 
# ```{r}
# data_combined %>% count(wt_recode1$wt_recode)
```



