---
title: "8-clean-civ-occupation-enlistment"
output: html_document
date: "2023-01-19"
---

```{r}
library(dplyr)   # %>%, group_by, tally, print
library(ipumsr)  # read_ipums_ddi, ipums_val_labels
library(readr)   # read_csv, write_csv
library(stringr) # str_trim, str_split
```

# Read data
```{r}
path_to_temp_data   <-  "/data/censoc/workspace/enlistment_records/temp7.csv"
out_path <- "/data/censoc/workspace/enlistment_records/"
civ_path <- "/data/censoc/input_data/enlistment_records_public/code/awenlistmentcode/cl_3323.csv" # to update
```

#Read in temp7
```{r}

data_combined <- read_csv(path_to_temp_data, 
                          col_types = cols(
                            empty_field = col_character(),
                            term_or_enlistment = col_character(),
                            longevity = col_character(),
                            education = col_character(),
                            marital_status = col_character(),
                            component = col_character(),
                            educ_spec = col_character(),
                            defer_date_mmyy = col_character(),
                            mname_clean = col_character(),
                            sex = col_integer(),
                            sex_r = col_integer(),
                            fname_std = col_character()
                            )
                          )
```


# Recode state of residence
```{r}
data_combined %>% 
  count(residence_state) 
```

Original state of residence
notes: this recoding used state codes present in original NARA documentation. These codes appear to be unique and not correspond to FIPS or IPUMS coding, for example.
```{r}
recode_res_state_enlistment <- function(df){
  
  ## recode source variable
  df <- df %>% 
    mutate(state_recode = case_when(
      residence_state == "41" ~ 41, #ALABAMA	100
      residence_state == "D1" ~ 41, #ALABAMA*	
      residence_state == "M1" ~ 41, #ALABAMA**	
      residence_state == "02" ~ 81,#Alaska 200
      residence_state == "R0" ~ 81, #Alaska*
      residence_state == "N9" ~ 81, #Alaska**
      residence_state == "98" ~ 61, #ARIZONA 400
      residence_state == "I8" ~ 61, #ARIZONA*
      residence_state == "R8" ~ 61, #ARIZONA**
      residence_state == "87" ~ 42, #ARKANSAS 500
      residence_state == "H7" ~ 42, #ARKANSAS
      residence_state == "Q7" ~ 42, #ARKANSAS
            residence_state == "91" ~ 71, #CALIFORNIA 600
      residence_state == "I1" ~ 71, #CALIFORNIA*
      residence_state == "R1" ~ 71, #CALIFORNIA**
      residence_state == "70" ~ 62, #COLORADO 800
      residence_state == "G0" ~ 62, #COLORADO
      residence_state == "P0" ~ 62, #COLORADO
            residence_state == "11" ~ 01, #CONNECTICUT 900
      residence_state == "A1" ~ 01, #CONNECTICUT
      residence_state == "J1" ~ 01, #CONNECTICUT
      residence_state == "21" ~ 11, #DELAWARE 1000
      residence_state == "B1" ~ 11, #DELAWARE
      residence_state == "K1" ~ 11, #DELAWARE
            residence_state == "34" ~ 98, #DISTRICT OF COLUMBIA	 1100
      residence_state == "C4" ~ 98, #DISTRICT OF COLUMBIA	
      residence_state == "L4" ~ 98, #DISTRICT OF COLUMBIA	
      residence_state == "42" ~ 43, #FLORIDA 1200
      residence_state == "D2" ~ 43, #FLORIDA
      residence_state == "M2" ~ 43, #FLORIDA
            residence_state == "43" ~ 44, #GEORGIA 1300
      residence_state == "D3" ~ 44, #GEORGIA
      residence_state == "R8" ~ 44, #GEORGIA
      residence_state == "03" ~ 82, #HAWAII 1500
      residence_state == "92" ~ 63, #IDAHO 1600
      residence_state == "I2" ~ 63, #IDAHO
            residence_state == "R2" ~ 63, #IDAHO
      residence_state == "61" ~ 21, #ILLINOIS 1700
      residence_state == "F1" ~ 21, #ILLINOIS
      residence_state == "O1" ~ 21, #ILLINOIS
      residence_state == "51" ~ 22, #INDIANA	1800
      residence_state == "E1" ~ 22, #INDIANA	
            residence_state == "N1" ~ 22, #INDIANA**	
      residence_state == "72" ~ 31, #IOWA 1900
      residence_state == "G2" ~ 31, #IOWA
      residence_state == "P2" ~ 31, #IOWA
      residence_state == "73" ~ 32, #KANSAS 2000
      residence_state == "G3" ~ 32, #KANSAS*
            residence_state == "P3" ~ 32, #KANSAS**
      residence_state == "52" ~ 51, #KENTUCKY 2100
      residence_state == "E2" ~ 51, #KENTUCKY
      residence_state == "N2" ~ 51, #KENTUCKY
      residence_state == "88" ~ 45, #LOUISIANA 2200
      residence_state == "H8" ~ 45, #LOUISIANA
            residence_state == "Q8" ~ 45, #LOUISIANA
      residence_state == "12" ~ 02, #MAINE 2300
      residence_state == "A2" ~ 02, #MAINE
      residence_state == "J2" ~ 02, #MAINE
      residence_state == "31" ~ 52, #MARYLAND 2400
      residence_state == "C1" ~ 52, #MARYLAND
            residence_state == "L1" ~ 52, #MARYLAND
      residence_state == "13" ~ 03, #MASSACHUSETTS 2500
      residence_state == "A3" ~ 03, #MASSACHUSETTS
      residence_state == "J3" ~ 03, #MASSACHUSETTS
      residence_state == "62" ~ 23, #MICHIGAN 2600
      residence_state == "F2" ~ 23, #MICHIGAN
            residence_state == "O2" ~ 23, #MICHIGAN
      residence_state == "74" ~ 33, #MINNESOTA 2700
      residence_state == "G4" ~ 33, #MINNESOTA
      residence_state == "P4" ~ 33, #MINNESOTA
      residence_state == "45" ~ 46, #MISSISSIPPI 2800
            residence_state == "D5" ~ 46, #MISSISSIPPI
      residence_state == "M5" ~ 46, #MISSISSIPPI
      residence_state == "75" ~ 34, #MISSOURI 2900
      residence_state == "G5" ~ 34, #MISSOURI
      residence_state == "P5" ~ 34, #MISSOURI
      residence_state == "93" ~ 64, #MONTANA 3000
            residence_state == "I3" ~ 64, #MONTANA
      residence_state == "38" ~ 64, #MONTANA
      residence_state == "76" ~ 35, #NEBRASKA 3100
      residence_state == "G6" ~ 35, #NEBRASKA
      residence_state == "P6" ~ 35, #NEBRASKA
            residence_state == "94" ~ 65, #NEVADA 3200
      residence_state == "I4" ~ 65, #NEVADA
      residence_state == "R4" ~ 65, #NEVADA
      residence_state == "14" ~ 04, #NEW HAMPSHIRE	3300
      residence_state == "A4" ~ 04, #NEW HAMPSHIRE	
      residence_state == "J4" ~ 04, #NEW HAMPSHIRE	
            residence_state == "22" ~ 12, #NEW JERSEY 3400
      residence_state == "B2" ~ 12, #NEW JERSEY
      residence_state == "K2" ~ 12, #NEW JERSEY
      residence_state == "83" ~ 66, #NEW MEXICO	 3500
      residence_state == "H3" ~ 66, #NEW MEXICO	
            residence_state == "Q3" ~ 66, #NEW MEXICO	
      residence_state == "23" ~ 13,  #NEW YORK	3600
      residence_state == "B3" ~ 13, #NEW YORK	
      residence_state == "K3" ~ 13, #NEW YORK	
      residence_state == "46" ~ 47, #NORTH CAROLINA	3700
      residence_state == "D6" ~ 47, #NORTH CAROLINA	
            residence_state == "M6" ~ 47, #NORTH CAROLINA	
      residence_state == "77" ~ 36, #NORTH DAKOTA	3800
      residence_state == "G7" ~ 36, #NORTH DAKOTA	
      residence_state == "P7" ~ 36, #NORTH DAKOTA	
      residence_state == "53" ~ 24, #OHIO   3900
            residence_state == "E3" ~ 24, #OHIO
      residence_state == "N3" ~ 24, #OHIO
      residence_state == "84" ~ 53, #OKLAHOMA 4000
      residence_state == "H4" ~ 53, #OKLAHOMA
      residence_state == "Q4" ~ 53, #OKLAHOMA
      residence_state == "95" ~ 72, #OREGON  4100
            residence_state == "I5" ~ 72, #OREGON
      residence_state == "R5" ~ 72, #OREGON
      residence_state == "32" ~ 14, #PENNSYLVANIA  4200
      residence_state == "C2" ~ 14, #PENNSYLVANIA
      residence_state == "L2" ~ 14, #PENNSYLVANIA
            residence_state == "08" ~ 83, #PUERTO RICO  ****
      residence_state == "15" ~ 05, #RHODE ISLAND	4400
      residence_state == "A5" ~ 05, #RHODE ISLAND	
      residence_state == "J5" ~ 05, #RHODE ISLAND	
      residence_state == "47" ~ 48, #SOUTH CAROLINA	 4500
      residence_state == "D7" ~ 48, #SOUTH CAROLINA	
            residence_state == "M7" ~ 48, #SOUTH CAROLINA	
      residence_state == "78" ~ 37, #SOUTH DAKOTA	 4600
      residence_state == "G8" ~ 37, #SOUTH DAKOTA	
      residence_state == "P8" ~ 37, #SOUTH DAKOTA	
      residence_state == "48" ~ 54, #TENNESSEE	4700
            residence_state == "D8" ~ 54, #TENNESSEE	
      residence_state == "M8" ~ 54, #TENNESSEE	
      residence_state == "85" ~ 49, #TEXAS 4800
      residence_state == "H5" ~ 49, #TEXAS
      residence_state == "Q5" ~ 49, #TEXAS
      residence_state == "96" ~ 67, #UTAH 4900
            residence_state == "I6" ~ 67, #UTAH
      residence_state == "R6" ~ 67, #UTAH
      residence_state == "16" ~ 06, #VERMONT 5000
      residence_state == "A6" ~ 06, #VERMONT
      residence_state == "J6" ~ 06, #VERMONT
            residence_state == "33" ~ 4, #VIRGINIA 5100 error in original recoding
      residence_state == "C3" ~ 40, #VIRGINIA
      residence_state == "L3" ~ 40, #VIRGINIA
      residence_state == "97" ~ 40, #WASHINGTON 5300 note mistake in WA/VA classification here
      residence_state == "I7" ~ 73, #WASHINGTON
      residence_state == "R7" ~ 73, #WASHINGTON
            residence_state == "54" ~ 56, #WEST VIRGINIA	5400
      residence_state == "E4" ~ 56, #WEST VIRGINIA	
      residence_state == "N4" ~ 56, #WEST VIRGINIA	
      residence_state == "63" ~ 25, #WISCONSIN   5500
      residence_state == "F3" ~ 25, #WISCONSIN
            residence_state == "O3" ~ 25, #WISCONSIN
      residence_state == "79" ~ 8, #WYOMING 5600
      residence_state == "G9" ~ 9,#WYOMING 
      residence_state == "P9" ~ 3, #WYOMING
      residence_state == "&" ~ 12, #WYOMING #none with this code, can remove.
      TRUE ~ NA_real_     
      ))
  
  ## return data.frame   
  return(df)
}
```


Function for recoding residence states to IPUMS detailed bpl codes
note: some birthplaces are ambiguous, we assume the US bpl is correct.
```{r}
recode_res_state_enlistment_to_ipums <- function(df){
  df <- df %>% 
    mutate(state_recode = case_when(
      residence_state == "41" ~ 100, #ALABAMA	100
      residence_state == "D1" ~ 100, #ALABAMA*	
      residence_state == "M1" ~ 100, #ALABAMA**	
      residence_state == "02" ~ 200, #Alaska 200
      residence_state == "R0" ~ 200, #Alaska*
      residence_state == "N9" ~ 200, #Alaska**
      residence_state == "98" ~ 400, #ARIZONA 400
      residence_state == "I8" ~ 400, #ARIZONA*
      residence_state == "R8" ~ 400, #ARIZONA**
      residence_state == "87" ~ 500, #ARKANSAS 500
      residence_state == "H7" ~ 500, #ARKANSAS
      residence_state == "Q7" ~ 500, #ARKANSAS
      residence_state == "91" ~ 600, #CALIFORNIA 600
      residence_state == "I1" ~ 600, #CALIFORNIA*
      residence_state == "R1" ~ 600, #CALIFORNIA**
      residence_state == "70" ~ 800, #COLORADO 800
      residence_state == "G0" ~ 800, #COLORADO
      residence_state == "P0" ~ 800, #COLORADO
      residence_state == "11" ~ 900, #CONNECTICUT 900
      residence_state == "A1" ~ 900, #CONNECTICUT
      residence_state == "J1" ~ 900, #CONNECTICUT
      residence_state == "21" ~ 1000, #DELAWARE 1000
      residence_state == "B1" ~ 1000, #DELAWARE
      residence_state == "K1" ~ 1000, #DELAWARE
      residence_state == "34" ~ 1100, #DISTRICT OF COLUMBIA	 1100
      residence_state == "C4" ~ 1100, #DISTRICT OF COLUMBIA	
      residence_state == "L4" ~ 1100, #DISTRICT OF COLUMBIA	
      residence_state == "42" ~ 1200, #FLORIDA 1200
      residence_state == "D2" ~ 1200, #FLORIDA
      residence_state == "M2" ~ 1200, #FLORIDA
      residence_state == "43" ~ 1300, #GEORGIA 1300
      residence_state == "D3" ~ 1300, #GEORGIA
      residence_state == "R8" ~ 1300, #GEORGIA
      residence_state == "03" ~ 1500, #HAWAII 1500
      residence_state == "R2" ~ 1500, #HAWAII
         # R2 = Idaho or Hawaiian Islands. Most people w/ code R2 are born in HI,
         # so we choose HI as place of residence
      residence_state == "92" ~ 1600, #IDAHO 1600
      residence_state == "I2" ~ 1600, #IDAHO
      residence_state == "61" ~ 1700, #ILLINOIS 1700
      residence_state == "F1" ~ 1700, #ILLINOIS
      residence_state == "O1" ~ 1700, #ILLINOIS
      residence_state == "51" ~ 1800, #INDIANA	1800
      residence_state == "E1" ~ 1800, #INDIANA	
      residence_state == "N1" ~ 1800, #INDIANA**	
      residence_state == "72" ~ 1900, #IOWA 1900
      residence_state == "G2" ~ 1900, #IOWA
      residence_state == "P2" ~ 1900, #IOWA
      residence_state == "73" ~ 2000, #KANSAS 2000
      residence_state == "G3" ~ 2000, #KANSAS*
            residence_state == "P3" ~ 2000, #KANSAS**
      residence_state == "52" ~ 2100, #KENTUCKY 2100
      residence_state == "E2" ~ 2100, #KENTUCKY
      residence_state == "N2" ~ 2100, #KENTUCKY
      residence_state == "88" ~ 2200, #LOUISIANA 2200
      residence_state == "H8" ~ 2200, #LOUISIANA
            residence_state == "Q8" ~ 2200, #LOUISIANA
      residence_state == "12" ~ 2300, #MAINE 2300
      residence_state == "A2" ~ 2300, #MAINE
      residence_state == "J2" ~ 2300, #MAINE
      residence_state == "31" ~ 2400, #MARYLAND 2400
      residence_state == "C1" ~ 2400, #MARYLAND
            residence_state == "L1" ~ 2400, #MARYLAND
      residence_state == "13" ~ 2500, #MASSACHUSETTS 2500
      residence_state == "A3" ~ 2500, #MASSACHUSETTS
      residence_state == "J3" ~ 2500, #MASSACHUSETTS
      residence_state == "62" ~ 2600, #MICHIGAN 2600
      residence_state == "F2" ~ 2600, #MICHIGAN
            residence_state == "O2" ~ 2600, #MICHIGAN
      residence_state == "74" ~ 2700, #MINNESOTA 2700
      residence_state == "G4" ~ 2700, #MINNESOTA
      residence_state == "P4" ~ 2700, #MINNESOTA
      residence_state == "45" ~ 2800, #MISSISSIPPI 2800
            residence_state == "D5" ~ 2800, #MISSISSIPPI
      residence_state == "M5" ~ 2800, #MISSISSIPPI
      residence_state == "75" ~ 2900, #MISSOURI 2900
      residence_state == "G5" ~ 2900, #MISSOURI
      residence_state == "P5" ~ 2900, #MISSOURI
      residence_state == "93" ~ 3000, #MONTANA 3000
            residence_state == "I3" ~ 3000, #MONTANA
      residence_state == "38" ~ 3000, #MONTANA
      residence_state == "76" ~ 3100, #NEBRASKA 3100
      residence_state == "G6" ~ 3100, #NEBRASKA
      residence_state == "P6" ~ 3100, #NEBRASKA
            residence_state == "94" ~ 3200, #NEVADA 3200
      residence_state == "I4" ~ 3200, #NEVADA
      residence_state == "R4" ~ 3200, #NEVADA
      residence_state == "14" ~ 3300, #NEW HAMPSHIRE	3300
      residence_state == "A4" ~ 3300, #NEW HAMPSHIRE	
      residence_state == "J4" ~ 3300, #NEW HAMPSHIRE	
            residence_state == "22" ~ 3400, #NEW JERSEY 3400
      residence_state == "B2" ~ 3400, #NEW JERSEY
      residence_state == "K2" ~ 3400, #NEW JERSEY
      residence_state == "83" ~ 3500, #NEW MEXICO	 3500
      residence_state == "H3" ~ 3500, #NEW MEXICO	
            residence_state == "Q3" ~ 3500, #NEW MEXICO	
      residence_state == "23" ~ 3600,  #NEW YORK	3600
      residence_state == "B3" ~ 3600, #NEW YORK	
      residence_state == "K3" ~ 3600, #NEW YORK	
      residence_state == "46" ~ 3700, #NORTH CAROLINA	3700
      residence_state == "D6" ~ 3700, #NORTH CAROLINA	
            residence_state == "M6" ~ 3700, #NORTH CAROLINA	
      residence_state == "77" ~ 3800, #NORTH DAKOTA	3800
      residence_state == "G7" ~ 3800, #NORTH DAKOTA	
      residence_state == "P7" ~ 3800, #NORTH DAKOTA	
      residence_state == "53" ~ 3900, #OHIO   3900
            residence_state == "E3" ~ 3900, #OHIO
      residence_state == "N3" ~ 3900, #OHIO
      residence_state == "84" ~ 4000, #OKLAHOMA 4000
      residence_state == "H4" ~ 4000, #OKLAHOMA
      residence_state == "Q4" ~ 4000, #OKLAHOMA
      residence_state == "95" ~ 4100, #OREGON  4100
            residence_state == "I5" ~ 4100, #OREGON
      residence_state == "R5" ~ 4100, #OREGON
      residence_state == "32" ~ 4200, #PENNSYLVANIA  4200
      residence_state == "C2" ~ 4200, #PENNSYLVANIA
      residence_state == "L2" ~ 4200, #PENNSYLVANIA
            residence_state == "08" ~ 11000, #PUERTO RICO  **** 11000
      residence_state == "15" ~ 4400, #RHODE ISLAND	4400
      residence_state == "A5" ~ 4400, #RHODE ISLAND	
      residence_state == "J5" ~ 4400, #RHODE ISLAND	
      residence_state == "47" ~ 4500, #SOUTH CAROLINA	 4500
      residence_state == "D7" ~ 4500, #SOUTH CAROLINA	
            residence_state == "M7" ~ 4500, #SOUTH CAROLINA	
      residence_state == "78" ~ 4600, #SOUTH DAKOTA	 4600
      residence_state == "G8" ~ 4600, #SOUTH DAKOTA	
      residence_state == "P8" ~ 4600, #SOUTH DAKOTA	
      residence_state == "48" ~ 4700, #TENNESSEE	4700
            residence_state == "D8" ~ 4700, #TENNESSEE	
      residence_state == "M8" ~ 4700, #TENNESSEE	
      residence_state == "85" ~ 4800, #TEXAS 4800
      residence_state == "H5" ~ 4800, #TEXAS
      residence_state == "Q5" ~ 4800, #TEXAS
      residence_state == "96" ~ 4900, #UTAH 4900
            residence_state == "I6" ~ 4900, #UTAH
      residence_state == "R6" ~ 4900, #UTAH
      residence_state == "16" ~ 5000, #VERMONT 5000
      residence_state == "A6" ~ 5000, #VERMONT
      residence_state == "J6" ~ 5000, #VERMONT
            residence_state == "33" ~ 5100, #VIRGINIA 5100
      residence_state == "C3" ~ 5100, #VIRGINIA
      residence_state == "L3" ~ 5100, #VIRGINIA
      residence_state == "97" ~ 5300, #WASHINGTON 5300
      residence_state == "I7" ~ 5300, #WASHINGTON
      residence_state == "R7" ~ 5300, #WASHINGTON
            residence_state == "54" ~ 5400, #WEST VIRGINIA	5400
      residence_state == "E4" ~ 5400, #WEST VIRGINIA	
      residence_state == "N4" ~ 5400, #WEST VIRGINIA	
      residence_state == "63" ~ 5500, #WISCONSIN   5500
      residence_state == "F3" ~ 5500, #WISCONSIN
            residence_state == "O3" ~ 5500, #WISCONSIN
      residence_state == "79" ~ 5600, #WYOMING 5600
      residence_state == "G9" ~ 5600, #WYOMING 
      residence_state == "P9" ~ 5600, #WYOMING
      TRUE ~ NA_real_     
      ))
  ## return data.frame   
  return(df)
}

```


Function for adding residence state FIPS codes in addition to the IPUMS recodes.
```{r}
res_state_fips_code <- function(df) {
  df <- recode_res_state_enlistment_to_ipums(df) %>% 
  select(residence_state, state_recode) %>%
    mutate(residence_state_fips = case_when(
      state_recode == 100 ~ 01, #Alabama
      state_recode == 200 ~ 02, #Alaska
      state_recode == 400 ~ 04, #Arizona
      state_recode == 500 ~ 05, #Arkansas
      state_recode == 600 ~ 06, #California
      state_recode == 800 ~ 08, #Colorado
      state_recode == 900 ~ 09, #Connecticut
      state_recode == 1000 ~ 10, #Delaware
      state_recode == 1100 ~ 11, #District of Columbia
      state_recode == 1200 ~ 12, #Florida
      state_recode == 1300 ~ 13, #Georgia
      state_recode == 1500 ~ 15, #Hawaii
      state_recode == 1600 ~ 16, #Idaho
      state_recode == 1700 ~ 17, #Illinois
      state_recode == 1800 ~ 18, #Indiana
      state_recode == 1900 ~ 19, #Iowa
      state_recode == 2000 ~ 20, #Kansas
      state_recode == 2100 ~ 21, #Kentucky
      state_recode == 2200 ~ 22, #Louisiana
      state_recode == 2300 ~ 23, #Maine
      state_recode == 2400 ~ 24, #Maryland
      state_recode == 2500 ~ 25, #Massachusetts
      state_recode == 2600 ~ 26, #Michigan
      state_recode == 2700 ~ 27, #Minnesota
      state_recode == 2800 ~ 28, #Mississippi
      state_recode == 2900 ~ 29, #Missouri
      state_recode == 3000 ~ 30, #Montana
      state_recode == 3100 ~ 31, #Nebraska
      state_recode == 3200 ~ 32, #Nevada
      state_recode == 3300 ~ 33, #New Hampshire
      state_recode == 3400 ~ 34, #New Jersey
      state_recode == 3500 ~ 35, #New Mexico
      state_recode == 3600 ~ 36, #New York
      state_recode == 3700 ~ 37, #North Carolina
      state_recode == 3800 ~ 38, #North Dakota
      state_recode == 3900 ~ 39, #Ohio
      state_recode == 4000 ~ 40, #Oklahoma
      state_recode == 4100 ~ 41, #Oregon
      state_recode == 4200 ~ 42, #Pennsylvania
      state_recode == 4400 ~ 44, #Rhode Island
      state_recode == 4500 ~ 45, #South Carolina
      state_recode == 4600 ~ 46, #South Dakota
      state_recode == 4700 ~ 47, #Tennessee
      state_recode == 4800 ~ 48, #Texas
      state_recode == 4900 ~ 49, #Utah
      state_recode == 5000 ~ 50, #Vermont
      state_recode == 5100 ~ 51, #Virginia
      state_recode == 5300 ~ 53, #Washington
      state_recode == 5400 ~ 54, #West Virginia
      state_recode == 5500 ~ 55, #Wisconsin
      state_recode == 5600 ~ 56, #Wyoming
      state_recode == 11000 ~ 72, #Puerto Rico
      TRUE ~ NA_real_
    ))
  
  return(df)
}

```


recode states and add state FIPS codes
```{r}
res_state_recode <- res_state_fips_code(df = data_combined) %>% 
  select(residence_state, state_recode, residence_state_fips)
```





#view civ_occupation fields
```{r}
data_combined %>% 
  count(civ_occupation) 
```


#Load csv with NARA AAD codes
```{r}
civ_occupation_codes <- read_csv(civ_path)
```


#Recode civ_occupation by removing bad values
```{r}
# is civ occ currently a string or integer?
str(data_combined$civ_occupation)  #character

civ_occupation_codes <- civ_occupation_codes %>% rename(civ_occupation = Code) %>% 
  rename(civ_occ_meaning = Meaning) %>% 
  select(-c("Comments"))

data_combined <- left_join(data_combined, civ_occupation_codes, by = "civ_occupation")

# make NA if the code does not correspond to anything
data_combined <- data_combined %>% 
  mutate(civ_occupation_recode = ifelse(!is.na(civ_occ_meaning), civ_occupation, NA_character_))

# Add codes for students (992), not in full code list
# (https://aad.archives.gov/aad/popup-column-detail.jsp?c_id=24986&dt=893&cl_id=3323)
data_combined <- data_combined %>% 
  mutate(civ_occupation_recode = ifelse(civ_occupation == "992", "992", civ_occupation_recode))

# clean up
data_combined$civ_occ_meaning <- NULL
```


#add civ_occupation_recode, res_state_recode, and residence_state_fips to data_combined
```{r}
data_combined <- cbind(data_combined, 
                       res_state_recode$state_recode,
                       res_state_recode$residence_state_fips)

```


Remove county info for anyone w/o valid state and Hawaii (1500), Alaska (200), Puerto Rico (11000)
```{r}
data_combined <- data_combined %>% mutate(residence_county =
                                            case_when(`res_state_recode$state_recode` == 1500 ~ NA_character_,
                                                      `res_state_recode$state_recode` == 200 ~ NA_character_,
                                                      `res_state_recode$state_recode` == 11000 ~ NA_character_,
                                                      is.na(`res_state_recode$state_recode`) ~ NA_character_,
                                                      TRUE ~ residence_county))
```



#save dataset as temp8
```{r}
write_csv(data_combined, paste0(out_path, "/temp8.csv"))
```

clean up
```{r}
rm(list = ls())
```
